---
title: "dataprep"
author: "Stephanie Ma Lucero"
date: "2024-04-02"
output: html_document
---


<-- NA

--> 1_cluster.Rmd

- Set working directory
```{r}
knitr::opts_chunk$set(echo = TRUE)

setwd("~/Library/CloudStorage/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022/chaparraldegradation_2022")

```

# load libraries
```{r}

library(tidyverse) # data wrangling - includes ggplot2, dplyr, tidyr, readr, purr, tibble, stringr, forcats
library(dplyr)
library(ggplot2) # data visualization
library(vegan) #community ecology package - includes, MASS, cluster, mgcv
library(readxl) # for .xls and .xlsx sheets
library(multcompView) #Visualizations of Paired Comparisons, functions: TukeyHSD, dist{stats}, simint, simtest, csimint, csimtest{multcomp}, friedmanmc, kruskalmc{pgirmess}


# library(here) # similar to set working directory
# library(lubridate) # dates and times
# library(gridExtra)
# library(readr)
# library(ggthemes)
# library(plotly)
# library(ggpubr) # customize ggplot2 for publication
#library(broom)
#library(AICcmodavg)
#library(janitor)
#library(multcomp) # functions: multcompLetters2, multcompLetters3, multcompLetters4 using aov object
#library(DHARMa) #The ‘DHARMa’ package uses a simulation-based approach to create readily interpretable scaled (quantile) residuals for fitted (generalized) linear mixed models.
#citation("DHARMa")
# library(ggpmisc)
# library(segmented)
# library(strucchange)
# library(ggpattern)
# library(lme4)
# library(lmerTest) # load again or update if you're not getting p-values from environmental models
# library(sjPlot)

```


# TRANSECTS

-  percent cover x species 

These are the data from Kit
df: data_pc
```{r bringing in the csv you sent me}

###################          USE THESE DATA FOR CLUSTER ANALYSIS! (july 22, 2025) - SML


# load transect (n = 40) data

data_pc <- read.csv("data/twinspan_percentcover.csv")

rownames(data_pc) <- data_pc[,1]
data_pc[,1] <- NULL
# dim(data_pc)

# 52 species (in columns)
# 40 transects (in rows)
# 
# NOTE (july 2025):
# loading the spp list from the raw 2019 and 2024 transect data results in 73 species entries.
# differences include: perennial dead species, ground cover, unknown 1 forb
# 
# OPEN datasheet in EXCEL to see differences in the species lists:
# load_diffintransectspp <- read_xlsx("data/scrappaper_july2025.xlsx", 
#                                     sheet = "spp_transect",   
#                                     col_types = "text")

```

This is the code to find transect species from the raw 2019 and 2024 data
```{r, eval = FALSE}

# What this code does: load point intercept transect data from csv
############################################### load file 2019 transect data

file_transect_2019_load <- read_csv("data/deg_percentcover_byspecies_bylifeform_2024.csv") # warnings are from Notes 
  
file_transect_2019_raw <- file_transect_2019_load %>% 
  dplyr::select(site, year, transect, distance, vertical, species, lifeform, status, lifecycle, family) %>% 
  mutate(vertical = case_when(vertical == "ground" ~ "0", # oldname ~ newname
                              TRUE ~ vertical)) %>% 
  mutate(
       site = as.character(site),
       transect = as.character(transect),
       year = as.character(year),
       distance = as.numeric(distance),
       vertical = as.numeric(vertical),
       species = as.character(species),
        lifeform = as.character(lifeform),
       status = as.character(status),
       lifecycle = as.character(lifecycle),
       family = as.character(family)
  ) %>% 
  drop_na()

  # ground cover other = moss, wood

sort(unique(file_transect_2019_raw$site))  
file_transect_2019_raw %>%  
  group_by(site, transect)
# -- years 2017, 2019
# -- Site 1: transects 1-14
# -- Site 2: transects 1-18
# -- INT1, INT2, INT3, INT4, INT5

# Finding species == "Bromus spp." to rename them by species
# transect2019_bromus <- file_transect_2019_raw %>%
#   filter(grepl("Bromus", species))
# 
# transect2019_bromusspp <- file_transect_2019_raw %>% 
#   filter(species == "Bromus spp.")


transect2019 <- file_transect_2019_raw %>% 
  dplyr::select(c(site, transect, year, distance, vertical, species)) %>% # select columns to full join with 2024 data
  # filter(year %in% c(2017, 2019))  %>% #filter data to 2019 data
  
   mutate(species = ifelse(# a condition, a value if the condition is true, and a value if the condition is false
                              site == 2 & transect == 5 & distance %in% c(2, 6, 9) & species == "Bromus spp.", "Bromus hordeaceus", species)) %>% 
    mutate(species = ifelse(site == 2 & transect == 5 & distance %in% c(10, 28) & species == "Bromus spp.", "Bromus rubens", species)) %>% 
    mutate(species = ifelse(site == 2 & transect == 6 & distance %in% c(23, 35, 37) & species == "Bromus spp.", "Bromus hordeaceus", species)) %>% 
    mutate(species = ifelse(site == "INT1" & transect == 1 & distance %in% c(24) & species == "Bromus spp.", "Bromus diandrus", species))


# View the data (optional)
#View(transect2019)
#sort(unique(transect2019$transect))
#sort(unique(transect2019$site))





############################################### load file 2024 transect data

file_transect_2024_load <- read.csv("data/Transects_2024 - transect.csv", header = TRUE, na.strings=c("","NA")) %>% 
  dplyr::select(site, transect, year, distance, vertical, species, max_height_cm) %>% 
   mutate(vertical = case_when(vertical == "ground" ~ "0", # oldname ~ newname
                              TRUE ~ vertical)) %>% 
  filter(transect  >= 19) %>% 
  mutate(
       site = as.character(site),
       transect = as.character(transect),
       year = as.character(year),
       distance = as.numeric(distance),
       vertical = as.numeric(vertical),
       species = as.character(species),
       max_height_cm = as.numeric(max_height_cm)
  ) %>% 
   filter(species != "N/A")
# -- years 2024
# -- Site 2: transects 19-21

transect2024 <- file_transect_2024_load %>%  
  dplyr::select(year, site, transect, distance, vertical, species) %>% 
   mutate(species = case_when(species == "Bromus spp." ~ "Bromus rubens", # oldname ~ newname
                              TRUE ~ species))
  
# View the data (optional)
#view(transect2024)
#sort(unique(transect2024$transect))
#sort(unique(transect2024$site))


############################################### join transect data

abcover_fulljoin <- full_join(transect2019, transect2024, by = intersect(names(transect2019), names(transect2024))) %>%  # full join
  as.tibble() # convert dataframe to tibble
# View(abcover_fulljoin)
# sort(unique(abcover_fulljoin$transect))
# sort(unique(abcover_fulljoin$site))


### correct species names
abcover_full <- abcover_fulljoin %>% 
   mutate(site_rep = paste(site, transect, sep = "_")) %>% # combine site and rep to one column
   mutate(site_rep_dist = paste(site_rep, distance, sep = "_")) %>%  # combine site_rep and distance to one column 
   mutate(species = case_when(species == "BG bare ground" ~ "bare_ground", # oldname ~ newname
                              species == "ground" ~ "bare_ground", 
                              species == "native_litter" ~ "litter",
                              species == "nonnative_thatch" ~ "thatch",
                              species == "Bromus diandrus BRDI" ~ "Bromus diandrus",
                              species == "Bromus rubens BRRU" ~ "Bromus rubens", 
                              species == "Bromus hordeaceus BRHO" ~ "Bromus hordeaceus", 
                              species == "Rapistrum rugosum - round pods" ~ "Rapistrum rugosum", 
                              species == "HIIN Hirschfeldia incana - long pods" ~ "Hirschfeldia incana", 
                              species == "Hirschfeldia incana - long pods" ~ "Hirschfeldia incana", 
                              species == "Salvia lucophylla" ~ "Salvia leucophylla",
                              species == "Malacothamnus fasciculatus " ~ "Malacothamnus fasciculatus",
                              species == "dead HEAR" ~ "Heteromeles arbutifolia dead",
                              species == "dead MAFA" ~ "Malacothamnus fasciculatus dead",
                              species == "dead RHIL" ~ "Rhamnus ilicifolia dead",
                              species == "dead ARCA" ~ "Artemisia californica dead",
                              species == "dead SALE" ~ "Salvia leucophylla dead",
                              species == "ACGL Acmispon glaber" ~ "Acmispon glaber",
                              species == "dead Acmispon glaber" ~ "Acmispon glaber dead",
                              species == "Acmispon maritimus - round, sevfl" ~ "Acmispon maritimus",
                              species == "Acmispon strigosus - oval, sevfl" ~ "Acmispon strigosus",
                              species == "Sambucus nigra" ~ "Sambucus mexicana",
                              species == "Blitum californicum" ~ "Chenopodium californicum",
                              species == "Cryptantha sp." ~ "Cryptantha spp.",
                                species == "Festuca myros" ~ "Festuca myuros",
                              TRUE ~ species)) %>% 
   mutate(
      year = as.character(year),
      site = as.character(site),
      transect = as.character(transect),
      distance = as.character(distance),
      vertical = as.character(vertical),
      species = as.character(species)
      )

glimpse(abcover_full) # look at new columns
sort(unique(abcover_full$species)) # species list; with: bare ground, wood, moss, native_thatch, etc.
length(unique(abcover_full$species)) # n = 73 unique species names
length(unique(abcover_full$species)) - 9 # n = 64 unique species names
  # bare ground
  # litter
  # no_species
  # moss
  # native_thatch
  # rock
  # thatch
  # wood
  # unknown 1 forb

### omit ground cover data
abcover_omitground <- abcover_full %>% 
  filter(vertical != "ground", 
         vertical != "ground cover") %>% 
  filter(species != "no_species", 
         species != "thatch") # %>% 
  # mutate(lifeform_status = paste(lifeform, status, sep = "_")) %>% 
  # mutate(lifeform_status_leaftype = paste(lifeform_status, leaftype, sep = "_"))


```

### lifeformleaftype df

```{r, percent cover - load data}

################  percent cover - belt transects, point intercept transects
abcover_full_2 <- read.csv("deg_percentcover_byspecies_bylifeform.csv", header = TRUE, na.strings=c("","NA")) %>% 
  mutate(standtype = fct_relevel(standtype, 
            "intact", "matrix", "degraded")) %>% 
  dplyr::select(!species_code) %>% 
    mutate(hillslope = case_when(hillslope == "Intact_1" ~ "INT1", 
                                hillslope == "Intact_2" ~ "INT2",
                                hillslope == "Intact_3" ~ "INT3", 
          TRUE ~ hillslope))

```


### - species characteristics


```{r}

################ species characteristic list

sppchar_load <- read.csv("processed/specieslist.csv") 

sppchar <- sppchar_load %>%  # species and their characteristics :) Created in 1.1_transect_continuous.Rmd
  dplyr::select(!c(X)) %>%
   mutate(species = str_replace(species, " ", "_")) %>%
  mutate(species = gsub("\u00A0", " ", species, fixed = TRUE)) %>%
  mutate(species = str_replace(species, " ", "_")) %>% 
    mutate(species = str_replace(species, "_", "."))

length(unique(sppchar$species))

############### leaf traits
shrub_leaftype <- read.csv("data/shrub_leaftype.csv") %>%
  dplyr::select("species", "leaftype") %>%
   mutate(
      species = as.character(species),
      leaftype = as.character(leaftype)
      ) %>%
  mutate(
    leaftype = if_else(grepl("dead", species, ignore.case = TRUE), "dead_shrub", leaftype)
  )

```

```{r}
### -- code to update list of species characteristics
ab_species <- sort(unique(abcover_full$species))
species_species <- sort(unique(specieslist$species))


```

```{r}
glimpse(abcover_full)
str(abcover_full)

#abcover_full$standtype <- factor(abcover_full$standtype, levels = c("intact", "matrix", "degraded"))

# convert dataframe to tibble
abcover_full_v <- as_tibble(abcover_full_2) 
class(abcover_full_v)

# combine site and rep to one column
abcover_full_v$site_rep <- paste(abcover_full_v$site, abcover_full_v$rep, sep="_") 

# combine site_rep and distance to one column
abcover_full_v$site_rep_dist <- paste(abcover_full_v$site_rep, abcover_full_v$distance, sep="_") 

# combine standtype and native/nonnative status to one column
abcover_full_v$stand_status <- paste(abcover_full_v$standtype, abcover_full_v$status, sep="_") 

abcover_full_v <- abcover_full_v %>% 
  mutate(site_rep = fct_relevel
         (site_rep, 
                                "INT1_1", "INT2_2", "INT3_3", "ENH1_1", "ENH1_2",    "ENH1_3"  ,   "ENH2_4"  ,   "ENH2_5"    , "ENH2_6"  , "DEG1_1"   ,  "DEG1_2"  ,   "DEG1_3"     ,"DEG2_4" ,    "DEG2_5"    , "DEG2_6" ))

glimpse(abcover_full_v) # look at new columns

sort(unique(abcover_full_v$species))

abcover_full_v %>%  
  na.omit() %>%  
  group_by(standtype, rep) %>% 
  summarize(data_points = length(rep))

# omit "ground" as a status since interested in Native and Non-native species 
abcover_full_v_omitground <- filter(abcover_full_v, status !="ground")
glimpse(abcover_full_v_omitground)

shrubs <- abcover_full_v_omitground %>%
  filter(lifeform == "shrub") %>%
  left_join(shrub_leaftype, by = "species") %>%
  dplyr::select(!c(lifeform)) %>%
    rename(lifeform = leaftype ) # newname = oldname

herbsgrassesfern <-  abcover_full_v_omitground %>% 
  filter(lifeform =="herb" | lifeform == "grass" | lifeform == "fern") 
  
lifeformleaftype <- rbind(herbsgrassesfern, shrubs) %>%  
  group_by(site_rep_dist) %>%
   mutate(lifeform_status = paste(lifeform, status, sep = "_")) %>% 
  group_by(standtype, site, rep, site_rep, lifeform, status, lifeform_status) %>% 
  dplyr::summarize(
    count_per_transect = n_distinct(site_rep_dist), 
    percentcover = n_distinct(site_rep_dist)/41) %>% # <-- calculate percent cover
  ungroup() %>% 
   mutate(standtype = fct_relevel(standtype,"intact", "matrix", "degraded")) %>% 
   mutate(site = fct_relevel(site, "INT1", "INT2", "INT3", "ENH1", "ENH2", "DEG1", "DEG2")) %>%
  mutate(site_rep = fct_relevel(site_rep, "INT1_1", "INT2_2", "INT3_3", "ENH1_1" ,    "ENH1_2"   ,  "ENH1_3" ,    "ENH2_4" ,    "ENH2_5"    ,"ENH2_6", "DEG1_1",     "DEG1_2"  ,   "DEG1_3"   ,  "DEG2_4" ,    "DEG2_5"   ,  "DEG2_6")) %>% 
  mutate(lifeform_status = fct_relevel(lifeform_status, "grass_nonnative", "herb_nonnative",
                                  "fern_native",     "grass_native", "herb_native", "nonsclero_native", "sclerophyllous_native")
  )
```

### - Total Veg Cover (100%)
```{r allvegetationcover, include = TRUE}
abcover_vegcover <- abcover_full_v %>% # total above veg
  filter(vertical == "1") %>% 
  group_by(standtype, site, rep) %>% 
  dplyr::summarize(count_per_transect = length(vertical), 
    percentcover_veg = length(vertical)/41, 
    percentcover_ground = (1-percentcover_veg)
    )
print(abcover_vegcover)

abcover_vegcover_long <- gather(abcover_vegcover, covertype, percentcover, percentcover_veg:percentcover_ground, factor_key = TRUE) 
abcover_vegcover_long
```



### - spp richness

# CANOPY height

# SHRUB density

# ENVIRON variables

```{r}

environ_load <- read.csv("data/Extracted_Elevation_Slope_Aspect_Hillshade_20m.csv", header = TRUE, na.strings=c("","NA")) 





```



- hillshade: higher numbers = more sunny. Lower numbers = more north facing

```{r load data}

################### load environmental data

extracted_load <- read.csv("~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022/chaparraldegradation_2022/Extracted_Elevation_Slope_Aspect_Hillshade_20m.csv", header = TRUE, na.strings=c("","NA")) %>% 
  dplyr::select(system_ind, aspect, elevation, slope, site, lat, long, cluster) %>% 
  mutate(southwestness = cos(225-aspect))
# sort(unique(extracted_load$system_ind))

twi_firecount_solarrad_load <- read.csv("~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022/chaparraldegradation_2022/Extracted_Elevation_Slope_Aspect_Hillshade_20m_20241014_8pm.csv", header = TRUE, na.strings = c("", "NA")) %>% 
  dplyr::select(system_ind, twi, summer_solarradiation, winter_solarradiation, equinox_solarradiation, fire_event_count) %>% 
  mutate(total_solrad = summer_solarradiation + winter_solarradiation + equinox_solarradiation + equinox_solarradiation) 
# sort(unique(twi_firecount_solarrad_load$system_ind))

extracted_join <- extracted_load %>% 
  left_join(twi_firecount_solarrad_load, by = "system_ind")
# sort(unique(extracted_join$system_ind))

extracted <- extracted_join%>% 
    rename(site_rep = system_ind) %>% 
 filter(site_rep != "1.03") %>% 
      mutate(site_rep = case_when(site_rep == "1.1" ~ "1.10", # old ~ new
                                   site_rep == "2.1" ~ "2.10",
                                   site_rep == "2.2" ~ "2.20",
                                  site_rep == "INT1" ~ "INT1_1",
                                    site_rep == "INT2" ~ "INT2_1",
                                    site_rep == "INT3" ~ "INT3_1",
                                    site_rep == "INT4" ~ "INT4_1",
                                    site_rep == "INT5" ~ "INT5_1",
                                   TRUE ~ site_rep)) %>% 
     rownames_to_column() %>% 
     mutate(
      southwestness = as.numeric(southwestness),
      elevation = as.integer(elevation), # meters
      slope = as.numeric(slope), # degrees
      site_rep = as.character(site_rep), 
      site = as.character(site),
      cluster = as.character(cluster),
      fire_event_count = as.integer(fire_event_count),
       rowname = as.character(rowname)
      ) %>% 
  rename(plant_cluster = cluster)


extracted$site_rep <- gsub("\\.", "_", extracted$site_rep)

# length(extracted$site_rep)    # n = 39


############# load plant clusters (n = 5)

# plant_cluster_list <- extracted_prep %>% 
#   dplyr::select(site_rep, plant_cluster)

 
####################### load Species Percent Cover data 
### copied/pasted from 1.3_NMDS_percentcover.Rmd; ~ line 100
abcover_charateristics_load <- read.csv("processed/ab_specieslistwithcharateristics_08102024.csv", header = TRUE, na.strings=c("","NA")) %>% 
  dplyr::select(!c(X)) 

sale_load <- abcover_charateristics_load %>% 
  filter(species == c("Salvia leucophylla", "Salvia leucophylla dead")) %>% 
  mutate(species = case_when(
    species == "Salvia leucophylla dead" ~ "Salvia leucophylla", 
    TRUE ~ species)) %>% 
   mutate(site_rep = case_when(site_rep == "1_1" ~ "1_01", # old name ~ new name
                        site_rep == "1_2" ~ "1_02",
                        site_rep == "1_3" ~ "1_03",
                        site_rep == "1_4" ~ "1_04",
                        site_rep == "1_5" ~ "1_05",
                        site_rep == "1_6" ~ "1_06",
                        site_rep == "1_7" ~ "1_07",
                        site_rep == "1_8" ~ "1_08",
                        site_rep == "1_9" ~ "1_09",
                        site_rep == "2_1" ~ "2_01",
                         site_rep == "2_2" ~ "2_02", 
                         site_rep == "2_3" ~ "2_03",
                         site_rep == "2_4" ~ "2_04",
                         site_rep == "2_5" ~ "2_05",
                         site_rep == "2_6" ~ "2_06",
                         site_rep == "2_7" ~ "2_07",
                         site_rep == "2_8" ~ "2_08",
                         site_rep == "2_9" ~ "2_09",
   TRUE ~ site_rep)) %>% 
   group_by(site_rep) %>% 
  dplyr::summarize(
    SALE_count_per_transect = sum(n_distinct(site_rep_dist)), 
    percentcover = (SALE_count_per_transect/41) *100
    ) %>% 
  dplyr::select(!c(SALE_count_per_transect)) 
  
```






# SOIL CORES


# load data
```{r load deg_seedbank csv}

########### load seedbank data ---> raw data

seedbank_load_raw <- read.csv("deg_seedbank_202407.csv", header = TRUE, na.strings=c("","NA")) 

seedbank_load <- seedbank_load_raw %>% 
  dplyr::select(c("stand", "site", "soilcore", "depth_name", "treatment", "rep", "species_diversity", "genus_abundance", "lifeform", "status", "lifescycle", "family", "count")) 
# 2024 december 18 - SML note - I added Quercus spp. (ilfeform == "tree") to the .csv.   


seedbank_load$count <- as.numeric(seedbank_load$count)
str(seedbank_load)
         
# print(seedbank_load)
# unique(seedbank_load$stand)
# unique(seedbank_load$site)
# unique(seedbank_load$soilcore)
# unique(seedbank_load$depth_name)
# unique(seedbank_load$treatment)
# unique(seedbank_load$status)
# sort(unique(seedbank_load$species_diversity))
# sort(unique(seedbank_load$genus_abundance))
# sort(unique(seedbank_load$family))
# sort(unique(seedbank_load$lifeform))
# sort(unique(seedbank_load$lifescycle))




######## load germination by depth and treatment --> calculated n per gram sieved soil

## Trifolium spp. isn't in the germination density list, but it IS in the germination species list.

germination_m2_raw <- read_xlsx("processed/old_processed/germination_m2.xlsx")
average_cover <- read.csv("nat_nn_cover.csv", header = TRUE, na.strings=c("","NA"))

germination_depthtreat <- germination_m2_raw %>%
  mutate(site = fct_relevel(site, 
            "INT1", "INT2", "INT3","ENH1", "ENH2", "DEG1", "DEG2")) %>% 
  mutate(treatment = fct_relevel(treatment, 
                                 "control", "charate", "oven", "ovenchar")) %>% 
  left_join(average_cover, by = 'site') %>% 
  dplyr::select(stand, site, depth_name, subsample_g, treatment, species_diversity, genus_abundance, lifeform, status, lifescycle, family, count, stand_core, stand_core_depth, stand_core_depth_treat, n_per_g, native_shrub, nonnative_grass) %>% 
  mutate(lifescycle = case_when(
     lifescycle == "biannual" ~ "annual", # old name ~ new name
       TRUE ~ as.character(lifescycle))) %>% # Keep other observations unchanged
  mutate(species_diversity = case_when(
    species_diversity == "Ribes spp." ~ "Ribes malvaceum",
    TRUE ~ as.character(species_diversity)
  ))

glimpse(germination_depthtreat)

Trifolspp <- germination_depthtreat %>% 
  filter(str_detect(species_diversity, "Trifolium"))
sort(unique(Trifolspp$sp))


########### load species characteristics list

abcover_charateristics_load <- read.csv("processed/ab_specieslistwithcharateristics_08102024.csv", header = TRUE, na.strings=c("","NA")) %>% 
  dplyr::select(!c(X)) %>% 
  dplyr::select(species, lifeform, status, lifecycle, family)
spp_charateristics <- abcover_charateristics_load %>% 
  distinct(species, .keep_all = TRUE) %>% 
  mutate(species = case_when(species == "Acmispon glaber" ~ "Acmispon spp.", # oldname ~ newname
                             species == "Bromus rubens" ~ "Bromus spp.",
                             species == "Claytonia perfoliata" ~ "Claytonia spp.",
                             species == "Festuca myros" ~ "Festuca myuros",
                              TRUE ~ species)) %>% 
    add_row(species = "Erigeron canadensis",
          lifeform = "forb",
          status = "native",
          lifecycle = "annual",
          family = "Asteraceae") %>% 
      add_row(species = "Eucrypta chrysanthemifolia",
          lifeform = "forb",
          status = "native",
          lifecycle = "annual",
          family = "Hydrophyllaceae") %>% 
         add_row(species = "Lactuca serriola",
          lifeform = "forb",
          status = "nonnative",
          lifecycle = "annual",
          family = "Asteraceae") %>% 
        add_row(species = "Nicotiana glauca",
          lifeform = "shrub",
          status = "nonnative",
          lifecycle = "perennial",
          family = "Solanacaeae") %>%
        add_row(species = "Pseudognaphalium californicum",
          lifeform = "forb",
          status = "native",
          lifecycle = "annual",
          family = "Asteraceae") %>% 
       add_row(species = "Ribes malvaceum",
          lifeform = "shrub",
          status = "native",
          lifecycle = "perennial",
          family = "Grossulariaceae") %>% 
  add_row(species = "Typha domingensis",
          lifeform = "forb",
          status = "native",
          lifecycle = "perennial",
          family = "Typhaceae")
      
sort(unique(abcover_charateristics_load$species))



Trifolspp1 <- abcover_charateristics_load %>% 
  filter(str_detect(species, "Trifolium"))
sort(unique(Trifolspp1$species))


##########

setdiff(germination_depthtreat$species_diversity, abcover_charateristics_load$species)
```


# data prep - seedbank raw

```{r preparing df, include = TRUE}


####################################    (1) Organize the seed bank data by converting the df into a tibble.

dim(seedbank_load)
# [1] 2126   13
# head(seedbank_load)
# tail(seedbank_load)

# convert dataframe to tibble
seedbank_load <- as_tibble(seedbank_load) 
# class(seedbank_load) # now a tibble







####################################    (2) Organize the seedbank data by reordering the classes: Intact, Matrix, Degraded.

seedbank_load_v <- seedbank_load %>%
  mutate(stand = fct_relevel(stand, 
            "intact", "enhanced", "degraded")) %>% 
  mutate(treatment = fct_relevel(treatment, 
                                 "control", "charate", "oven", "ovenchar")) %>% 
  mutate(family = case_when(
    family ==  "Asteraceae  " ~ "Asteraceae", # old name ~ new name
    family ==  "Cucurbitaceae  " ~ "Cucurbitaceae",
    family ==  "Euphorbiaceae  " ~ "Euphorbiaceae",
    family ==  "Fabaceae  " ~ "Fabaceae",
    family ==  "Juncaceae  " ~ "Juncaceae",
    family ==  "Liliaceae  " ~ "Liliaceae",
    family ==  "Namaceae  " ~ "Namaceae",
   #family ==  "Phrymaceae  "  ~ "Phrymaceae",
    family ==  "Poaceae  " ~ "Poaceae",
    family == "Themidaceae  " ~ "Themidaceae", 
     TRUE ~ as.character(family)  # Keep other observations unchanged
  )) %>% 
   mutate(species_diversity = case_when(
     species_diversity == "Astragalus trichopodus " ~ "Astragalus trichopodus",
     species_diversity == "Erodium spp. "  ~ "Erodium spp.",
     species_diversity == "Festuca myuros " ~ "Festuca myuros",
     species_diversity == "Lactuca serriola " ~ "Lactuca serriola",
     species_diversity ==   "Lupinus bicolor "   ~ "Lupinus bicolor", 
     species_diversity ==  "Salvia leucophylla "   ~ "Salvia leucophylla", 
     species_diversity == "Thysanocarpus spp. " ~ "Thysanocarpus spp.",     
     species_diversity == "Ribes spp." ~ "Ribes malvaceum",     
     TRUE ~ as.character(species_diversity)  # Keep other observations unchanged
  )) %>% 
   mutate(lifescycle = case_when(
     lifescycle == "annual or biannual" ~ "annual", # old name ~ new name
       TRUE ~ as.character(lifescycle)  # Keep other observations unchanged
  )) %>% 
  mutate(genus_abundance = case_when(
    genus_abundance == "Lactuca serriola "  ~ "Lactuca serriola",
    genus_abundance == "Lupinus bicolor "   ~ "Lupinus bicolor", 
    genus_abundance == "Thysanocarpus spp. "  ~ "Thysanocarpus spp.",
    genus_abundance == "Unknown Phrymaceae  " ~ "Unknown Phrymaceae",
    TRUE ~ as.character(genus_abundance)  # Keep other observations unchanged
  )) 

sppchar <- seedbank_load_v %>%  
  dplyr::select(genus_abundance, lifeform, status, lifescycle, family) %>% 
  rename(species = genus_abundance)





####################################   (3) Create new columns for SITE_CORE, SITE_CORE_DEPTH, SITE_CORE_DEPTH_TREATMENT, SITE_TREATMENT, SITE_DEPTH for seed bank data. 


# combine SITE and CORE to one column
seedbank_load_v$stand_core <- paste(seedbank_load_v$site, seedbank_load_v$soilcore, sep="_") 

# combine SITE_CORE and DEPT to one column
seedbank_load_v$stand_core_depth <- paste(seedbank_load_v$stand_core, seedbank_load_v$depth_name, sep="_") 

# combine SITE_CORE_DEPTH and TREATMENT to one column
seedbank_load_v$stand_core_depth_treat <- paste(seedbank_load_v$stand_core_depth, seedbank_load_v$treatment, sep="_") 

# combine SITE_CORE_DEPTH and SPECIES to one column
seedbank_load_v$stand_core_depth_spp <- paste(seedbank_load_v$stand_core_depth, seedbank_load_v$species_diversity, sep="_") 

glimpse(seedbank_load_v)






####################################   (4) Omit "pre-treatment" as a treatment since I'm only interested in Control, Char, Oven, and CharOven treatments.  
seedbank_load_v_omitpretreatment <- seedbank_load_v %>% 
  filter(!str_detect(treatment, "pretreat")) 
# head(seedbank_load_v_omitpretreatment)

# list treatments 
# unique(seedbank_load_v_omitpretreatment$treatment) 

# list status 
# unique(seedbank_load_v_omitpretreatment$status)

# list stand_core --> 26 = 7 + 10 + 9
# unique(seedbank_load_v_omitpretreatment$stand_core) 
```




# data prep for ...
Data preparation for running NMDS and Cluster analysis
- data prep
```{r updated nmds data prep}
# species in columns
# sites or observations in rows

germination_depthtreat$site_core_depth_genus <- paste(germination_depthtreat$stand_core_depth, germination_depthtreat$genus_abundance, sep="_") 

########### running CLUSTER with MAX n per g per genus_abundance
seedbank_nperg_bysamplocation <- germination_depthtreat %>% 
  group_by(site_core_depth_genus) %>% 
  dplyr::summarise(max_n_per_g = max(n_per_g)) %>% # take max germ
  separate(col = site_core_depth_genus, into = c("site", "core", "depth", "genus"), sep = "_") %>%
   mutate(site_core = paste(site, core, sep = "_")) %>% 
   mutate(site_core_genus = paste(site_core, genus, sep = "_")) %>% 
  dplyr::select(depth, max_n_per_g, site_core_genus) %>% 
  pivot_wider(names_from = depth, values_from = max_n_per_g) %>% 
 mutate(across(c(deep, surface), ~replace_na(.x, 0))) %>% 
  rowwise() %>% 
  mutate(totalgerm_nperg_genus = mean(c(surface, deep), na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(check = (deep+surface)/2) %>% 
separate(col = site_core_genus, into = c("site", "core", "genus_abundance"), sep = "_") %>%
   mutate(stand_core = paste(site, core, sep = "_")) %>% 
  dplyr::select(site, stand_core, genus_abundance, totalgerm_nperg_genus)
   

########### running CLUSTER with SUM n per g
# seedbank_nperg_bysamplocation_old <- germination_depthtreat %>%
#   group_by(stand, site, stand_core, genus_abundance) %>% # lifeform, status, lifescycle, family
#   filter(count > 0) %>%
#   filter(!is.na(genus_abundance)) %>% # omit blank cells in species
#   dplyr::summarize(totalgerm_nperg_genus = sum(n_per_g))
# 
# head(seedbank_nperg_bysamplocation) # df before transpose - use for ellipses
# dim(seedbank_nperg_bysamplocation) #330   5
```

 - transpose data 
```{r updated transpose data by species for nmds, include=TRUE}

# species in columns down
# observations or sites in rows across

seedbank_nperg_bysamplocation_nmdsprep <- seedbank_nperg_bysamplocation %>% 
 tidyr::spread(genus_abundance, totalgerm_nperg_genus) %>% # transpose data so columns are species names
  replace(is.na(.), 0) # replace NA with zeros --> isn't compatible with character columns

# view(seedbank_nperg_bysamplocation_nmdsprep)

# write.csv(seedbank_nperg_bysamplocation_nmdsprep, "~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022b/processed/seedbank_NMDS_202401.csv")
```

 - replace NAs with zeros
```{r updated replace NAs with zeros}
seedbank_nperg_bysamplocation_nmdsprep_num <- seedbank_nperg_bysamplocation_nmdsprep %>% 
    ungroup() %>% 
  dplyr::select(!c('site', 'stand_core')) %>% 
  replace(is.na(.), 0)
  
# head(seedbank_nperg_bysamplocation_nmdsprep_num) # only numbers
dim(seedbank_nperg_bysamplocation_nmdsprep_num) # 26 55
```


### abground percent cover
```{r}
soil_cover_load <- read_xlsx("data/abcover_ofsoilcores_20250720.xlsx", sheet = "soil_cover_data") 

soil_cover_raw <- soil_cover_load %>% 
  dplyr::select(date, plot_number, ch2_covertype, cover_percent, soilID) %>% 
  dplyr::mutate(date = as.Date(date))
  
```