---
title: "1_cluster_analysis"
author: "Stephanie Lucero"
date: "2024-06-13"
output: html_document
editor_options: 
  chunk_output_type: console
---

<--- 0_dataprep.Rmd

--> 

What this code does:
Rmarkdown for Using cluster analysis to identify plant groupings among chaparral transects



load libraries
```{r}

library(cluster)
library(factoextra)
library(gridExtra)
library(gt)
library(ggrepel) # avoid overlaping labes in ggplot2

library(indicspecies)
library(plotly)


library(stringr)
library(kableExtra)
library(htmltools)
library(webshot2)

# library(MASS)
# library(dunn.test)
# library(caret)
# library(rpart)
# library(ggparty)
# library(stringr)


# Set seed for analyses
set.seed(3980)

```


# CLUSTER ANALYSIS

NOTE from Robert Fitch or Kit Swift: Assume your data is in a dataframe called `data` (only numeric variables)
df: data_pc <-- from 0_dataprep.Rmd (~ line 66)


### - run cluster analysis

- calculate the dissimilarity matrix
```{r clustering}

# NAME data set
df <- data_pc %>% 
  dplyr::select(!c(site, site_rep))

d <- vegdist(df, method = "bray")  # Bray-Curtis dissimilarity
# Calculate distance measure matrix, Bray is the best for community data
# Jaccard Index and Sorensen-Dice Coefficient treat percent cover data as binary (present/absent)
 
# Check which method produces best clustering
# average = Computes the average distance between all pairs of observations across two clusters.
# single = Uses the minimum distance between pairs of observations in two clusters.
# complete = Considers the maximum distance between pairs of observations in two clusters.
# ward = Minimizes the total within-cluster variance; merges clusters with the smallest increase in total variance.
# gaverage = Merges clusters based on the distance between their centroids (means of points in clusters).
m <- c( "average", "single", "complete", "ward", "gaverage")
names(m) <- c( "average", "single", "complete", "ward", "gaverage")

# function to compute coefficient
ac <- function(x) {
  agnes(d, method = x, diss = TRUE, par.method = -0.25)$ac}

# Best method?
map_dbl(m, ac)
# Best method is g average at 0.8563148 = Flexible Beta method


#########################
#### Find optimal number of clusters - k-means clustering

# Scree plot  
fviz_nbclust(df, FUN = hcut, diss = d, method = "wss") # Elbow method for kmeans

# Average Silhouette
# Measures how similar an object is to its own cluster (cohesion) compared to other clusters (separation). The Average Silhouette Width is the mean of all silhouette scores within a cluster or across all clusters.
# S ≈ 1: The point is well clustered
fviz_nbclust(df, FUN = hcut, diss = d, method = "silhouette") # Average silhouette for kmeans
# 4 groups has highest Avg. sillhouete width

# Gap Stat
# Compares the total within-cluster variation for different numbers of clusters with their expected values under a null reference distribution of the data (i.e., uniformly distributed points).
# The optimal number of clusters is the smallest k such that the Gap Statistic is maximized or the first local maximum occurs, indicating that adding more clusters does not significantly improve the clustering.
fviz_nbclust(df, FUN = hcut, diss = d, method = "gap_stat")
# Eight clusters is best, adding more clusteres doesn't improve clustering
# My conclusion is 5 clusters is optimum


# Above tests suggest at least two, but are also not conclusive...

# Define function to compute Mantel correlation tests for different cluster assignments
compute_mantel_correlation <- function(hclust_obj, distance_matrix) {

  # Initialize variables to store results
  mantel_correlation <- numeric(length = length(hclust_obj$order) - 1)
  
  # Iterate through different numbers of clusters
  for (k in 2:length(hclust_obj$order)) {
    cluster_assignments <- cutree(hclust_obj, k = k)
    cluster_distance <- as.matrix(dist(cluster_assignments))
    mantel_result <- mantel(cluster_distance, distance_matrix)
    mantel_correlation[k - 1] <- mantel_result$statistic
    # Print Mantel correlation for each number of clusters
    print(paste("Mantel Correlation (r) for", k, "clusters:", mantel_result$statistic))
  }
  
  return(mantel_correlation)
}

# Create Cluster and then run tests
Hclust.result <- agnes(d, method = "gaverage", par.method = -0.25)
mantel_results <- round(compute_mantel_correlation(Hclust.result, d), 3)
# Negative values of beta (e.g., -0.25) make the clustering process more similar to single linkage (minimum distance between clusters).
# Positive values of beta make the process more similar to complete linkage (maximum distance). beta = 0 corresponds to the standard average linkage (UPGMA).


# Find the optimal number of clusters based on the largest Mantel correlation, r
optimal_clusters <- which.max(mantel_results) + 1  # Add 1 to convert index to number of clusters
print(paste("Optimal Number of Clusters:", optimal_clusters))
# 4 groups = 0.608 
# 6 groups = 0.560
# 5 groups = 0.554
# 7 groups = 0.520




####
#### Use ISA (indicator species analysis) to determine optimal clusters

# Function to run multipatt() on k iterations of clusters and extract ISA values
run_multipatt_and_extract_ISA <- function(data, max_clusters) {
  
  # Initialize an empty dataframe to store ISA values
  isa_df <- data.frame(Species = colnames(data))
  
  # Iterate through different numbers of clusters
  for (k in 2:max_clusters) {
    cluster_assignments <- cutree(agnes(vegdist(data, method = "bray"), method = "gaverage", par.method = -0.25), k = k)
    
    # Convert cluster assignments to a factor
    cluster_factor <- as.factor(cluster_assignments)
    
    # Run multipatt() with the cluster factor
    multipatt_result <- multipatt(data, cluster = cluster_factor, duleg = FALSE)
    
    # Extract ISA values and p-values
    isa_values <- multipatt_result$sign$index
    p_values <- multipatt_result$sign$p.value
    
    # Filter by p-value <= 0.05
    significant_isa_values <- ifelse(p_values <= 0.05, isa_values, NA)
    
    # Extract ISA values for each species and add to the dataframe
    isa_df[[as.character(k)]] <- significant_isa_values
  }
  
  return(isa_df)
}


max_clusters <- 10  # Maximum number of clusters to iterate through

# Run multipatt() and extract ISA values for each species
isa_data <- run_multipatt_and_extract_ISA(df, max_clusters)

# Frequency distribution of number of significant ISA in each cluster step
frq.ISA <- isa_data %>%
  pivot_longer(cols = 2:10, names_to = "Cluster", values_to = "ISA_SCORE") %>%
  drop_na() %>% 
  group_by(Cluster) %>% 
  summarise(count = n()) %>% 
  arrange(as.numeric(Cluster))


#######
# Now calculate the sum of the differences in ISA scores from one step to the next
# When the indicator values of all species are decreasing, the clustering method does not explain anything more. 
# Take difference of species index values

isa.diff.data <- isa_data %>% 
  mutate(k2 = `2`-`3`) %>% 
  mutate(k3 = `3`-`4`) %>%
  mutate(k4 = `4`-`5`) %>%
  mutate(k5 = `5`-`6`) %>%
  mutate(k6 = `6`-`7`) %>%
  mutate(k7 = `7`-`8`) %>%
  mutate(k8 = `8`-`9`) %>%
  mutate(k9 = `9`-`10`) 

isa.diff.sum <- isa.diff.data %>%
  dplyr::select(Species, 11:18) %>% 
  pivot_longer(cols = 2:9, names_to = "Cluster", values_to = "ISA_SCORE") %>% 
  drop_na() %>%
  filter(ISA_SCORE <= 0) %>% 
  arrange(by = Cluster) %>% 
  group_by(Cluster) %>% 
  summarise(sum = sum(ISA_SCORE))


ggplot(isa.diff.sum,
       aes(x = Cluster,
           y = sum)) +
  geom_bar(stat = "identity") +
  labs(title = "", x = "Cluster", y = "Value") +
  theme_minimal()

# The differences from 2 to 3 groups (-8) and from 3 to 4 groups (-4) are small, 
# indicating that moving from 2 to 4 clusters doesn’t significantly improve the ISA scores. 
# This suggests that the structure in the data is relatively stable with these numbers of clusters, 
# and adding more groups is not greatly improving how well species are associated with their clusters.

# The difference from 4 to 5 groups is -90, which is a noticeable drop. This indicates that introducing a 5th cluster starts to capture more distinct ecological patterns, improving the ISA scores.

# Optimal Number of Clusters:
# 4 clusters may be suboptimal due to the small improvement seen moving to 4 clusters from 3 (-4).
# 5 clusters is likely the best choice. The drop to -90 suggests that the 5-cluster solution 
# is where distinct ecological patterns are maximized before significant over-segmentation occurs.
# Adding more than 5 clusters likely results in diminishing returns, as indicated by the sharp declines in ISA scores.


# Doing 5 clusters! 

plot(Hclust.result)
rect.hclust(Hclust.result, k = 5)

# Cut tree into 5 groups
sub_grp <- cutree(Hclust.result, k = 5)

# Number of members in each cluster
table(sub_grp)

# Calculate the distance of each sample to the centroid for each cluster
bd <- betadisper(d, sub_grp) # <-- this is the distance of samples to their group centroid
# plot(bd)
centroid_coords <- bd$centroids
centroid_distances <- round(vegdist(centroid_coords, method = "euclidean"), 2) # Calculate the distances between the centroids
#      1    2    3    4
# 2 0.62               
# 3 0.64 0.69          
# 4 0.60 0.41 0.54     
# 5 0.83 0.71 0.73 0.65


# Combine group labels into species matrix
temp.df <- data_pc %>%
  rownames_to_column()

my_groups <-  as.data.frame(sub_grp) %>% 
  rownames_to_column() %>% 
   mutate(sub_grp_name = recode(sub_grp,
                                 "1" = "NNG",
                                 "2" = "ssF", 
                                 "3" = "ssG",
                                 "4" = "ssFG",
                                "5" = "EVG"
                          )) 
cluster.subgrp.df <- left_join(temp.df, my_groups, by = "rowname") %>% 
   mutate(combined_labels = paste(sub_grp_name, site_rep, sep = "."))


# Plot the hierarchical clustering dendrogram with the combined labels
# plot(hcluster.abs.cover, labels = cluster.subgrp.df$combined_labels, main = "Hierarchical Clustering Dendrogram")
# rect.hclust(hcluster.abs.cover, k = 5)
hclust_obj <- as.hclust(Hclust.result)
cluster <- plot(
  hclust_obj,
  labels = cluster.subgrp.df$combined_labels,
  main = "Hierarchical Clustering Dendrogram",
  cex.main = 1.5,         # Increase title size
  cex = 1.5,  # Adjust this value as needed to enlarge text
   hang = -1               # Better spacing
)
rect.hclust(hclust_obj, k = 5, border = "red", )



```


### - indicator Species Analysis
```{r}

ISA.abs.cover <- multipatt(x = df, cluster = sub_grp, duleg = FALSE)

# A = The values are the **specificity** component of IV for that species in that combination of groups.
# B = The values are the **fidelity** component of IV for that species in that combination of groups
summary(ISA.abs.cover, indvalcomp=TRUE)
#  Multilevel pattern analysis
#  ---------------------------
# 
#  Association function: IndVal.g
#  Significance level (alpha): 0.05
# 
#  Total number of species: 52
#  Selected number of species: 11 
#  Number of species associated to 1 group: 5 
#  Number of species associated to 2 groups: 2 
#  Number of species associated to 3 groups: 2 
#  Number of species associated to 4 groups: 2 
# 
#  List of species associated to each combination: 
# 
#  Group 1  #sps.  1 
#                   A      B  stat p.value  
# Lupinus.spp. 0.8571 0.4000 0.586    0.04 *
# 
#  Group 4  #sps.  1 
#                    A      B  stat p.value   
# Festuca.myros 0.8444 0.6250 0.726    0.01 **
# 
#  Group 5  #sps.  3 
#                             A     B  stat p.value   
# Heteromeles.arbutifolia 1.000 1.000 1.000   0.005 **
# Rhus.ovata              0.863 0.600 0.720   0.010 **
# Rhamnus.ilicifolia      1.000 0.400 0.632   0.010 **
# 
#  Group 1+3  #sps.  1 
#                          A      B  stat p.value   
# Hirschfeldia.incana 0.8869 0.7500 0.816   0.005 **
# 
#  Group 2+4  #sps.  1 
#                           A      B  stat p.value   
# Centaurea.melitensis 0.9715 0.8667 0.918   0.005 **
# 
#  Group 1+3+4  #sps.  1 
#                    A      B  stat p.value   
# Avena.barbata 0.9453 0.9286 0.937   0.005 **
# 
#  Group 2+4+5  #sps.  1 
#                   A      B  stat p.value   
# Stipa.lepida 0.9958 0.6500 0.805   0.005 **
# 
#  Group 1+2+3+4  #sps.  1 
#                    A      B stat p.value   
# Bromus.rubens 0.9573 0.9429 0.95   0.005 **
# 
#  Group 1+3+4+5  #sps.  1 
#                      A      B  stat p.value  
# Bromus.diandrus 0.9836 0.7273 0.846    0.03 *
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1 
# > 

vegcluster_indicatorspp <- tibble(species = c(
                "Heteromeles.arbutifolia",
                "Rhus.ovata",
                "Rhamnus.ilicifolia",      
    
                "Lupinus.spp.", 
                
                "Stipa.lepida",
                
                "Avena.barbata",
                "Bromus.rubens",
                "Bromus.diandrus",
               "Festuca.myros",
               "Hirschfeldia.incana",
               "Centaurea.melitensis"
                ))
```

### - PCoA community composition
```{r}
##############
# Does community composition differ among sites?

# Check for homogeneity of dispersion among groups, required for MANOVA

# There is significant dispersion within groups, and groups are unbalanced --- cannot use MANOVA (adonis)
anova(betadisper(d, my_groups$sub_grp_name))
# A low p-value (e.g., < 0.05) indicates that the dispersion (variance) among groups is significantly different, meaning that the assumption of homogeneity of dispersion is violated.
# Response: Distances
#           Df  Sum Sq   Mean Sq F value  Pr(>F)  
# Groups     4 0.11019 0.0275474  3.0475 0.02957 *
# Residuals 35 0.31638 0.0090393      

mod <- betadisper(d, my_groups$sub_grp_name) 
# same output as bd above, but this one includes the group names
# betadisper() <-- calculates the distance of each sample to its group centroid in multivariate space.

TukeyHSD(mod) # <-- compares the average distances to centroid between all pairs of groups.
# Useful to test if any one group is significantly more or less variable (dispersed) than another.
boxplot(mod) # <-- plots distances to centroids by group.
# Visually compares the spread (variability) of each group in multivariate space.
plot(mod) # <-- ordination-style plot
# Helpful for visualizing whether some groups are more dispersed than others.

results <- mrpp(d, grouping = sub_grp, permutations = 999) # <-- mrpp = Multi-Response Permutation Procedure
# The key idea is simple: if groups differ, the mean within-group dissimilarity should be smaller
# than the mean dissimilarity among randomly selected groups of the same size.
results 
# Call:
# mrpp(dat = d, grouping = sub_grp, permutations = 999) 
# 
# Dissimilarity index: bray 
# Weights for groups:  n 
# 
# Class means and counts:
# 
#         1      2      3      4      5     
# delta = 0.2092 0.2408 0.1518 0.1582 0.1561 <- avg within-group dissimilarity
# n =     10     7      10     8      5     
# 
# Chance corrected within-group agreement A: 0.592 <-- A=effect size 
#       A = 1: all items within groups are identical.
#       A > 0: less heterogeneity within groups than expected
#       A = 0: heterogeneity within groups equals what is expected by chance.
# OBSERVED delta 0.1835 <-- the average within-group distance
# EXPECTED delta 0.4498 <-- the average distance expected by chance
# 
# Significance of delta: 0.001 <- p-value yay!
# Permutation: free
# Number of permutations: 999
```


### - Visualize nmds
- visualize community differences in NMDS

```{r}
# NMDS data prep
# Picking k=3 for small stress value, and interpretation
mds <- metaMDS(df, k = 3, trymax = 400, autotransform = FALSE)

mds$stress
# k = 2D: stress test = 0.1648784
# k = 3D: stress test = 0.1128948
# k = 4D: stress test = 0.0831275

stressplot(mds) # A Shepard plot compares the observed dissimilarities (distances in the original data) to the fitted distances (distances in the reduced NMDS space).
gof <- goodness(mds) # Higher values indicate points that are poorly represented in the reduced-dimensional space, contributing more to the stress
gof
# min(gof) = 0.010
# max(gof) = 0.031 <- relationships in your data are well represented!

plot(mds, display = "sites", type = "none")
points(mds, display = "sites", cex = 2* # 2x for visibility
         gof/mean(gof)) # standardized by the mean
# larger points = less well represented


# Plot NMDS
mds.df <- vegan::scores(mds, display = "sites") %>% 
  as.data.frame() %>% 
  rownames_to_column("Sites") %>% 
  rownames_to_column("rowname") %>% 
  full_join(my_groups, by = "rowname") %>%
  mutate(sub_grp = as.character(sub_grp))

# 
# ggplot(mds.df,
#        aes(x = NMDS1,
#            y = NMDS2,
#            color = sub_grp,
#            shape = sub_grp)) +
#   geom_point(size = 3, alpha = 0.8) +
#   stat_ellipse() + 
#   theme(axis.title = element_text(size = 20)) +
#   theme(axis.text = element_text(size = 16))
# 
# 
# ggplot(mds.df,
#        aes(x = NMDS2,
#            y = NMDS3,
#            color = sub_grp,
#            shape = sub_grp)) +
#   geom_point(size = 3, alpha = 0.8) +
#   stat_ellipse() + 
#   theme(axis.title = element_text(size = 20)) +
#   theme(axis.text = element_text(size = 16))


plot_ly(mds.df,
        x = ~NMDS1,
        y = ~NMDS2,
        z = ~NMDS3,
        type =  "scatter3d",
        color=sub_grp)



# vectors
genp <- envfit(mds, df, perm = 999) # generate p-values 
genp

pv <- genp$vectors$pvals %>% # extract p-values for each species
  as.data.frame() %>%
  rownames_to_column("species") %>%
  dplyr::rename("pvals" = ".") %>%
  left_join(sppchar, by = "species")
pv

# extract coordinates for species, only keep species with p-val = 0.05
spp <- genp %>% 
  scores(., display = "vectors") %>% # NMDS1, NMDS2 values
  as.data.frame() %>% 
  rownames_to_column("species") %>% 
  full_join(., pv, by = "species") # join pvalues and spp
  # filter(pvals <= 0.05)# <-- p-value 

# filter to keep the species that are indicator species (from cluster and indicator spp analysis)
matching_indicatorspp <- spp %>%
  filter(species %in% vegcluster_indicatorspp$species)


### plot
# Convert sub_grp to a factor with the desired order
mds.df$sub_grp <- factor(mds.df$sub_grp, 
                         levels = c("1", "3", "2", "4", "5"))  # Reorder as needed

#### native species
nmds_nat <- ggplot(mds.df,
       aes(x = NMDS1,
           y = NMDS2
           )) +
  geom_point(size = 3, alpha = 0.8, 
             aes(color = sub_grp, shape = sub_grp, fill = sub_grp)) +
   stat_ellipse(aes(color = sub_grp), size = 1, alpha = 0.5) +  # Apply color here. Alpha = transparency!
  theme(axis.title = element_text(size = 20)) +
  theme(axis.text = element_text(size = 16)) +
  geom_segment(data = matching_indicatorspp %>%  filter(status == "native"), 
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")),
               col = "black") +
  geom_text(data = matching_indicatorspp %>%  filter(status == "native"), 
            aes(label = species),  
             size = 5, 
            position = position_jitter(width = 0.0, height = 0.2)) + # <-- jitter
    # how the data points look
  scale_shape_manual(values = c("1" = 21, "2" = 21, "3" = 21, "4" = 21, "5" = 7),
                     labels = c("1" = "NNG", "2" = "ssF", "3" = "ssG", "4" = "ssFG", "5" = "EVG")) +
  scale_color_manual(values = c("1" = "black", "2" = "blue", "3" = "orange", "4" = "purple", "5" = "darkgreen"),
  # scale_color_manual(values = c("1" = "black", "2" = "black", "3" = "black", "4" = "black", "5" = "black"),
                     labels = c("1" = "NNG", "2" = "ssF", "3" = "ssG", "4" = "ssFG", "5" = "EVG")) +
  scale_fill_manual(values = c("1" = "black", "2" = "grey85", "3" = "grey60", "4" = "white", "5" = "black"),
                     labels = c("1" = "NNG", "2" = "ssF", "3" = "ssG", "4" = "ssFG", "5" = "EVG")) +
  theme_bw() +
   theme(
    axis.title = element_text(size = 20),
    axis.text = element_text(size = 16),
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 16)
  ) +
   guides(color = guide_legend(title = "Plant Group"),
          shape = guide_legend(title = "Plant Group"),
         fill = guide_legend(title = "Plant Group")) +
   # geom_label(aes(x = 1.5, y = 1, label = "NATIVE"), size = 6)
   geom_label(x = 1.75, y = 1, label = "NATIVE", size = 6)
 



#### non-native species
nmds_nn <- ggplot(mds.df,
       aes(x = NMDS1,
           y = NMDS2
           )) +
  geom_point(size = 3, alpha = 0.8, aes(color = sub_grp, shape = sub_grp, fill = sub_grp)) +
   stat_ellipse(aes(color = sub_grp), size = 1, alpha = 0.5) +  # Apply color here
  theme(axis.title = element_text(size = 20)) +
  theme(axis.text = element_text(size = 16)) +
  geom_segment(data = matching_indicatorspp %>%  filter(status == "nonnative"), 
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")),
               col = "black") +
  geom_text(data = matching_indicatorspp %>%  filter(status == "nonnative"), 
            aes(label = species),  
             size = 5, 
            position = position_jitter(width = 0.0, height = 0.2)) +
    # how the data points look
  scale_shape_manual(values = c("1" = 21, "2" = 21, "3" = 21, "4" = 21, "5" = 7),
                     labels = c("1" = "NNG", "2" = "ssF", "3" = "ssG", "4" = "ssFG", "5" = "EVG")) +
  scale_color_manual(values = c("1" = "black", "2" = "blue", "3" = "orange", "4" = "purple", "5" = "darkgreen"),
                     labels = c("1" = "NNG", "2" = "ssF", "3" = "ssG", "4" = "ssFG", "5" = "EVG")) +
  scale_fill_manual(values = c("1" = "black", "2" = "grey85", "3" = "grey60", "4" = "white", "5" = "black"),
                     labels = c("1" = "NNG", "2" = "ssF", "3" = "ssG", "4" = "ssFG", "5" = "EVG")) +
  theme_bw() +
    theme(
    axis.title = element_text(size = 20),
    axis.text = element_text(size = 16),
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 16)
  ) +
   guides(color = guide_legend(title = "Plant Group"),
          shape = guide_legend(title = "Plant Group"),
         fill = guide_legend(title = "Plant Group")) +
  geom_label(x = 1.75, y = 1, label = "NON-NATIVE", size = 6)


# Combine plots
grid.arrange(nmds_nat, 
             nmds_nn,
             nrow = 2, 
             ncol = 1, 
             widths = c(1), # Custom widths for each column
             heights = c(1, 1) # Custom heights for each row
             ) 

## ggsave("Ch1_Degradation_figures/nmds_above_vectors.png", nmdsabove_plot_site_rep, width = 8, height = 6)

```

# PERCENT x cluster

### - Table summary
```{r}

table_clustercover <- cluster.subgrp.df %>%
   # mutate(sub_grp = recode(sub_grp, 
   #                         `1` = "NNG", 
   #                         `2` = "ssF", 
   #                         `3` = "ssG", 
   #                         `4` = "ssFG", 
   #                         `5` = "EVG")) %>%
   pivot_longer(cols = 4:55, names_to = "species", values_to = "percent_cover") %>%
   group_by(sub_grp_name, species) %>%
   summarise(mean_percentcover = round(mean(percent_cover, na.rm = TRUE)*100, 2)) %>%
   pivot_wider(names_from = sub_grp_name, values_from = mean_percentcover) %>%
  mutate(species = case_when(
    species == "Festuca.myros" ~	"Festuca.myuros", 
    TRUE ~ species)) %>% 
  left_join(sppchar, by = "species") %>% 
   # mutate(species = ifelse(
   #    species %in% c("Artemisia.californica", 
   #                   "Heteromeles.arbutifolia", 
   #                   "Salvia.apiana", 
   #                   "Baccharis.pilularis",
   #                   "Ceanothus.oliganthus",
   #                   "Eriodictyon.crassifolium",
   #                   "Eriogonum.fasciculatum",
   #                   "Malacothamnus.fasciculatus",
   #                   "Rhamnus.ilicifolia",
   #                   "Rhus.ovata",
   #                   "Salvia.leucophylla",
   #                   "Sambucus.mexicana"
   #                   ),
   #    paste0("<b><i>", species, "<b><i>"), # Bold and italicize these species
   #    paste0("<i>", species, "<i>")              # Italicize all other species
   # )) %>% 
   mutate(species = str_replace_all(species, "\\.", " ")) %>%
   mutate(species = str_trim(species)) %>%
   mutate(species = case_when(
      species == "Cryptantha spp" ~ "Cryptantha spp.", # old name ~ new name
      species == "Erodium spp" ~ "Erodium spp.",
      species == "Lupinus spp" ~ "Lupinus spp.",
      TRUE ~ species
   )) %>% 
   relocate(status, lifecycle, lifeform, family, NNG, ssG, ssF, ssFG, EVG, .after = species)  %>%  # Move 'lifeform' after 'Species'
  rename(Species = species, 
         Status = status, 
         Lifecycle = lifecycle, 
         Lifeform = lifeform, 
         Family = family) %>% 
  mutate(Status = case_when(
    Status == "native" ~ "nat", 
    Status == "nonnative" ~ "nn", 
    TRUE ~ Status)
  ) %>% 
  mutate(Lifecycle = case_when(
    Lifecycle == "annual" ~ "ann", 
    Lifecycle == "perennial" ~ "per", 
    TRUE ~ Lifecycle)
  ) %>% 
  #  mutate(Lifeform = case_when(
  #   Lifeform == "forb" ~ "f", 
  #   Lifeform == "grass" ~ "g", 
  #    Lifeform == "shrub" ~ "s", 
  #   TRUE ~ Lifeform)
  # ) 
  mutate(Lifeform = ifelse(Species == "Heteromeles arbutifolia", "shrub (scl)", Lifeform)) %>%  # a condition, a value if the condition is true, and a value if the condition is false
  mutate(Lifeform = ifelse(Species == "Rhamnus ilicifolia", "shrub (scl)", Lifeform)) %>% 
  mutate(Lifeform = ifelse(Species == "Rhus ovata", "shrub (scl)", Lifeform)) 
# view(table_clustercover)


# Print as a gt table
table_cover <- table_clustercover %>%
  gt() %>%
  tab_header(
    title = "Mean Percent Cover by Community-type",
    #subtitle = "Mean ± Standard Deviation for Elevation, Hillshade, Slope, Southwestness, Fire Events, and Coordinates"
  ) %>%
  cols_align(
    align = "left", # Center alignment
    columns = "Species" # Specify the "cluster" column
  ) %>%
  tab_options(
    table.width = pct(80) # Make the table width 100%
  )
table_cover

# Export the table as a PDF
gtsave(table_cover, 
       "Ch1_Degradation_figures/table_spp_clustergroups_oct31.docx", # can also export as .png
       expand = 10)


# # print as a HTML table with KABLE
# table_cover <- table_clustercover %>%
#    kable(format = "html", escape = FALSE) %>%
#    kable_styling("striped", full_width = FALSE)
# 
# table_cover

```

```{r}
print("End of code :)")
```


# END 

Old code below


```{r, eval = FALSE}

# NAME data set
df <- data_pc %>% 
  dplyr::select(!c(site, site_rep))

d <- vegdist(df, method = "bray")  # Bray-Curtis dissimilarity

# Step 3: Perform hierarchical clustering
clustering <- agnes(d, method = "gaverage", par.method = -0.25)

# Step 4: Plot the dendrogram
# plot(clustering, labels = rownames(data), main = "Hierarchical Clustering Dendrogram")
plot(clustering, labels = data_pc$site_rep, main = "Hierarchical Clustering Dendrogram")


# Optionally, cut the tree to form clusters
cluster_2 <- cutree(clustering, k = 2)  # Cut the tree into 2 clusters
table(cluster_2)
data_pc$cluster_2 <- cluster_2

cluster_3 <- cutree(clustering, k = 3)  # Cut the tree into 3 clusters
table(cluster_3)
data_pc$cluster_3 <- cluster_3

cluster_4 <- cutree(clustering, k = 4)  # Cut the tree into 4 clusters
table(cluster_4)
data_pc$cluster_4 <- cluster_4

cluster_5 <- cutree(clustering, k = 5)  # Cut the tree into 5 clusters
table(cluster_5)
data_pc$cluster_5 <- cluster_5

cluster_6 <- cutree(clustering, k = 6)  # Cut the tree into 4 clusters
table(cluster_6)
data_pc$cluster_6 <- cluster_6

cluster_7 <- cutree(clustering, k = 7)  # Cut the tree into 4 clusters
table(cluster_7)
data_pc$cluster_7 <- cluster_7

cluster_8 <- cutree(clustering, k = 8)  # Cut the tree into 4 clusters
table(cluster_8)
data_pc$cluster_8 <- cluster_8

cluster_9 <- cutree(clustering, k = 9)  # Cut the tree into 4 clusters
table(cluster_9)
data_pc$cluster_9 <- cluster_9

cluster_10 <- cutree(clustering, k = 10)  # Cut the tree into 4 clusters
table(cluster_10)
data_pc$cluster_10 <- cluster_10
```

```{r, eval = FALSE}
plot(clustering, labels = data_pc$cluster_2, main = "Hierarchical Clustering Dendrogram") # splits are ranked by the height of their splits
plot(clustering, labels = data_pc$cluster_3, main = "Hierarchical Clustering Dendrogram")
plot(clustering, labels = data_pc$cluster_4, main = "Hierarchical Clustering Dendrogram")

plot(clustering, labels = data_pc$cluster_5, main = "Hierarchical Clustering Dendrogram") # these seem to be the major splits
plot(clustering, labels = data_pc$cluster_6, main = "Hierarchical Clustering Dendrogram") # this is the beginning of the smaller (less significant?) splits

plot(clustering, labels = data_pc$cluster_7, main = "Hierarchical Clustering Dendrogram") # HIIN is significant here
plot(clustering, labels = data_pc$cluster_8, main = "Hierarchical Clustering Dendrogram")
plot(clustering, labels = data_pc$cluster_9, main = "Hierarchical Clustering Dendrogram")
plot(clustering, labels = data_pc$cluster_10, main = "Hierarchical Clustering Dendrogram")
```

```{r indicator species, eval = FALSE}

# Assume `species_data` is your community data matrix (sites as rows, species as columns)
# Assume `clusters` is a vector defining the group membership for each site

# Run the Indicator Species Analysis
isa_results_2 <- multipatt(data, cluster_2, control = how(nperm=999))
# Summarize the results
summary(isa_results_2)
# View the significant indicator species
species_pvalues_2 <- isa_results_2$sign
significant_species_2 <- species_pvalues_2 %>% 
  filter(p.value <= 0.05)
significant_species_2

isa_results_3 <- multipatt(data, cluster_3, control = how(nperm=999))
summary(isa_results_3)
species_pvalues_3 <- isa_results_3$sign
significant_species_3 <- species_pvalues_3 %>% 
  filter(p.value <= 0.05)
significant_species_3

isa_results_4 <- multipatt(data, cluster_4, control = how(nperm=999))
summary(isa_results_4)
species_pvalues_4 <- isa_results_4$sign
significant_species_4 <- species_pvalues_4 %>% 
  filter(p.value <= 0.05)
significant_species_4

isa_results_5 <- multipatt(data, cluster_5, control = how(nperm=999)) # <- 5 clusters
summary(isa_results_5)
species_pvalues_5 <- isa_results_5$sign
significant_species_5 <- species_pvalues_5 %>% 
  filter(p.value <= 0.05)
significant_species_5

summary()isa_results <- multipatt(data, cluster_6, control = how(nperm=999)) # <- 6 clusters

isa_results <- multipatt(data, cluster_7, control = how(nperm=999))
isa_results <- multipatt(data, cluster_8, control = how(nperm=999))
isa_results <- multipatt(data, cluster_9, control = how(nperm=999))
isa_results <- multipatt(data, cluster_10, control = how(nperm=999))


```

```{r example, echo=FALSE, eval = FALSE}
# library(cluster)
# silhouette_5 <- silhouette(cluster_5, dist(data_pc))
# silhouette_6 <- silhouette(cluster_6, dist(data_pc))

# NOT RUN {
set.seed(123)

# Data preparation
# +++++++++++++++
data("iris")
head(iris)
# Remove species column (5) and scale the data
iris.scaled <- scale(iris[, -5])


# Optimal number of clusters in the data
# ++++++++++++++++++++++++++++++++++++++
# Examples are provided only for kmeans, but
# you can also use cluster::pam (for pam) or
#  hcut (for hierarchical clustering)
 
### Elbow method (look at the knee)
# Elbow method for kmeans
fviz_nbclust(iris.scaled, kmeans, method = "wss") +
geom_vline(xintercept = 4, linetype = 2)

# Average silhouette for kmeans
fviz_nbclust(iris.scaled, kmeans, method = "silhouette")

### Gap statistic
library(cluster)
set.seed(123)
# Compute gap statistic for kmeans
# we used B = 10 for demo. Recommended value is ~500
gap_stat <- clusGap(iris.scaled, FUN = kmeans, nstart = 25,
 K.max = 10, B = 10)
 print(gap_stat, method = "firstmax")
fviz_gap_stat(gap_stat)
 
# Gap statistic for hierarchical clustering
gap_stat <- clusGap(iris.scaled, FUN = hcut, K.max = 10, B = 10)
fviz_gap_stat(gap_stat)

 
# 
```

