---
title: "above ground species richness"
author: "Stephanie Ma Lucero"
date: "2024-02-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##### Rmarkdown for analyzing total above ground species richness 
- combining species lists from point-intercept transects and belt transects

# library
```{r install libraries, message=FALSE}

# Packages
library(tidyverse) # includes ggplot2, dplyr, tidyr, readr, purr, tibble, stringr, forcats
library(vegan)
library(here)
library(readxl) 

# library(ggpubr) # customize ggplot2 for publication
# library(broom)
# library(AICcmodavg)
# library(janitor)
# library(multcompView)
# library(plotly)
# if (!require(devtools)) {
#   install.packages("devtools")
# }
# devtools::install_github("gavinsimpson/ggvegan")
# library(ggpmisc)
```


```{r }

# load colors

# library(calecopal) 
 
# Stand colors: INTACT, MATRIX, DEGRADED 
# cal_palette(name = "chaparral3", n = 3, type = "continuous") 
# deg <- "#D3E3CA"
# mat <- "#92A587"
# int <- "#2F3525"
# standcolors <- c(int, mat, deg)
# 
# # Lifeform colors: shrub, grass, forb
# cal_palette(name = "chaparral3", n = 3, type = "continuous") 
# shrub <- "#D3E3CA"
# grass <- "#92A587"
# forb <- "#2F3525"
# lifeformcolors <- c(int, mat, deg)
# 
# # Status colors: native, exotic 
# cal_palette(name = "chaparral1", n = 6, type = "continuous") 
# native <- "#AEBFA8"
# exotic <- "#F19B34"
# statuscolors <- c(native, exotic)
```



## data



```{r load belt intercept transect data}

################## load belt transect data

# 2019 belt transect data 
belt.spp_2019_load <- read.csv(here("deg_beltdensity_byspecies.csv")) 

belt.spp_2019 <- belt.spp_2019_load %>% 
  dplyr::select(site, rep, species) %>% 
  mutate(site_rep = paste(site, rep, sep = ".")) %>% 
  mutate(species = case_when(species == "Bromus madritensis" ~ "Bromus rubens", # old name ~ new name
                             species == "Eucrypta\xcachrysanthemifolia" ~ "Eucrypta chrysanthemifolia",
                             TRUE ~ species)) %>%
  mutate(species = str_replace(species, " ", "_"))


### 2024 belt transect data
belt.spp_2024_load <- read.csv(here("belt_spprich_2024.csv"))  

belt.spp_2024 <- belt.spp_2024_load %>% 
  dplyr::select(site, rep, species) %>% 
  mutate(site_rep = paste(site, rep, sep = ".")) %>% 
  filter(species != "moss") %>%
  mutate(species = str_replace(species, " ", "_")) %>% 
  mutate(species = case_when(species == "Toxicodendron_pubescens" ~ "Toxicodendron_diversilobum", # old name ~ new name
                             species == "Sambucus_nigra" ~ "Sambucus_mexicana",
                             TRUE ~ species))

belt.spprich <- bind_rows(belt.spp_2019, belt.spp_2024) 
# sort(unique(belt.spprich$species))


# List of seven unknown species: chaparraldegradation2022 > chaparraldegradation2022 > unknown_spp.xlsx


### join species characteristics
sppchar <- read.csv("processed/specieslist.csv") %>% # species and their characteristics :) Created in 1.1_transect_continuous.Rmd
  dplyr::select(!c(X, notes, ejepson)) %>%
  mutate(species = str_replace(species, " ", "_")) %>%
  mutate(species = gsub("\u00A0", " ", species, fixed = TRUE)) %>%
  mutate(species = str_replace(species, " ", "_")) %>% 
  mutate(family = case_when(family == "Agavaceae  " ~ "Agavaceae", # old name ~ new name
                             family == "Boraginaceae " ~ "Boraginaceae",
                              family == "Chenopodiaceae  " ~ "Chenopodiaceae",
                              family == "Fagaceae  " ~ "Fagaceae",
                            family == "Onagraceae  " ~ "Onagraceae",
                            family == "Ranunculaceae  " ~ "Ranunculaceae",
                            family == "Urticaceae  " ~ "Urticaceae",
                             TRUE ~ family))

#These are the species that are in belt.spprich that are NOT in the species characteristics list
a <- sort(unique(belt.spprich$species))
b <- sort(unique(sppchar$species))

speciesmissingfrom.sppchar <- setdiff(a, b) # <- KIT - can you help resolve this issue? I added the missing species and their characteristics to "processed/specieslist.csv" but they're not joining
speciesmissingfrom.sppchar 

speciesmissingfrom.belt <- setdiff(b, a)
speciesmissingfrom.belt

# list of seven unknown species
# unknown.spp <- read_excel("unknown_spp.xlsx")

belt <- belt.spprich %>% 
  left_join(sppchar, by = "species") %>% 
   mutate(site_rep_species = paste(site_rep, species, sep = "_")) 



################## load point intercept transect data

abcover_full_load <- read.csv("deg_percentcover_byspecies_bylifeform.csv", header = TRUE, na.strings=c("","NA")) # point-intercept data

file_transect_2019_load <- read_excel("deg_percentcover_byspecies_bylifeform_2024.xlsx", sheet = "2019_data")

ab_sppr_raw <- abcover_full_load %>% 
  dplyr::select(!c(X, distance, vertical)) %>% 
  filter(status != "ground") %>% # unselect ground
   mutate(site = case_when(
    site == "Intact_1" ~ "Intact1",
      site == "Intact_2" ~ "Intact2",
      site == "Intact_3" ~ "Intact3",
    TRUE ~ as.character(site)  # Keep other observations unchanged
  )) %>% 
  mutate(site_rep = paste(site, rep, sep = "_")) %>%
   mutate(site_rep = case_when(
    site_rep == "DEG2_4" ~ "DEG2_1", # old name ~ new name
      site_rep == "DEG2_5" ~ "DEG2_2",
      site_rep == "DEG2_6" ~ "DEG2_3",
     site_rep == "ENH2_4" ~ "ENH2_1",
      site_rep == "ENH2_5" ~ "ENH2_2",
     site_rep == "ENH2_6" ~ "ENH2_3",
    TRUE ~ as.character(site_rep))) %>%   # Keep other observations unchanged
   filter(!is.na(site)) %>% # omit blank cells in species 
    as_tibble() %>% 
  mutate(species = case_when(
    species == "Avena barbata dead" ~ "Avena barbata", # old name ~ new name
    species == "Bromus diandrus dead" ~ "Bromus diandrus", 
     species == "Bromus madritensis" ~ "Bromus rubens", 
     species == "Bromus rubens dead" ~ "Bromus rubens", 
       species == "Bromus spp. dead" ~ "Bromus spp.", 
     species == "Centaurea melitensis dead" ~ "Centaurea melitensis", 
    species == "Elymus condensatus dead" ~ "Elymus condensatus", 
   species == "Heteromeles arbutifolia dead" ~ "Heteromeles arbutifolia",
   species == "Hirschfeldia incana dead" ~ "Hirschfeldia incana",
    species == "Malacothamnus fasciculatus dead" ~ "Malacothamnus fasciculatus", 
   species == "Salvia leucophylla dead" ~ "Salvia leucophylla",
   species == "Stipa lepida dead" ~ "Stipa lepida",
   TRUE ~ as.character(species)
  )) %>% 
   mutate(site_rep_species = paste(site_rep, species, sep = "_")) 

sort(unique(ab_sppr_raw$species))
length(unique(ab_sppr_raw$site_rep_species))
# 107

ptintcp_list <- list(unique(ab_sppr_raw$site_rep_species))


################################ shrub density


# native shrub density 
shrubdensitylife_full <- read_csv("~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022/chaparraldegradation_2022/deg_beltdensity_bylifestage.csv", 
    col_types = cols(
      StandType = col_character(), 
      Site = col_character(),
      rep = col_integer(),
      ScientificName = col_character(), 
      dead = col_integer(),
      immature = col_integer(),
      mature = col_integer(), 
      resprout = col_integer(), 
      seedling = col_integer()
      )
     ) %>% 
  rename()
  


shrubdensityspp_full <- read_csv("~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022/chaparraldegradation_2022/deg_beltdensity_byspecies.csv")
# shrubdensityalive_full <- read_csv("~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/deg_beltdensity_seperatealive.csv")
```



### RICHNESS: spp, genera, family

```{r}

################ species richness


# sort(unique(belt.spprich$species))

belt.spprich_total <- belt %>% 
  filter(species != "Erodium_spp.") %>% 
  filter(species != "Galium_spp.") %>% 
  filter(species != "Lupinus_spp.") %>% 
    filter(species != "Phacelia_spp.") %>% 
    filter(species != "Trifolium_spp.")
sort(unique(belt.spprich_total$species))

length(sort(unique(belt.spprich_total$species)))
# 87 species  

   
# erodium <- belt.spp_2019 %>%  filter(grepl("Erodium", species))
# lupin <- belt.spp_2019 %>%  filter(grepl("Lupinus", species))
# 
# belt.spp_2019 %>% 
#   filter(!(species == "Lupinus spp." & site_rep == "1.1")) %>% 
#   filter(!(species == "Lupinus spp." & site_rep == "1.2"))



################ genera richness

belt.genrich <- belt %>% 
  separate(col = species, into = c("Genus", "Species"), sep = "_") %>%
  dplyr::select("Genus") 
length(sort(unique(belt.genrich$Genus))) - 4 # minus 4 unknowns
# 67 

################ family richness

belt_famrich <- belt %>% 
   dplyr::select("family") 
# sort(unique(belt_famrich$family))
length(sort(unique(belt_famrich$family))) - 1 # duplicate Chenopodiaceae
# 36

```


```{r duplicate sppr, include=FALSE}

# total spp richness using ab_sppr_raw and belt dfs - same calculation as above (n = 87)

ab_sppr_raw2 <- ab_sppr_raw %>% 
  dplyr::select(site, rep, species, site_rep, lifeform, status, lifecycle, family, site_rep_species)


belt2 <- belt %>% 
   mutate(species = str_replace(species, "_", " "))
  sort(unique(belt2$species))
  
ab_totalsppr <- dplyr::union(ab_sppr_raw2, belt2) 
length(unique(ab_totalsppr$species)) -6 # Bromus spp., Erodium spp.,  "Galium spp."  , Lupinus spp., Phacelia spp, Trifolium spp.
# 87

```

```{r}
# difference in n species in point intercept transect vs. belt transect
trans_n <- length(unique(ab_sppr_raw %>% 
  dplyr::select(species) %>% 
  pull(species))) - 2 # Bromus spp, Erodium spp
trans_n
# 31

# z <- ab_sppr_raw %>% 
#   dplyr::select(species)
# sort(unique(z$species))

belt_n <- length(unique(belt %>% 
  dplyr::select(species) %>% 
  pull(species))) - 5 # erodium sop., galium spp, lupinus spp., phacelia spp., trifolium spp
belt_n
# 87

# y <- belt %>%
#   dplyr::select(species)
# sort(unique(y$species))

87-31
# 56

round(31/87*100, 2)
# [1] 35.63
```


# NATIVE lifeforms
```{r}

####### shrub
length(sort(unique(
  belt %>% 
  filter(status == "native") %>% 
  filter(lifeform == "shrub") %>% 
    pull(species))))
# 14


####### grass
length(sort(unique(
  belt %>% 
  filter(status == "native") %>% 
  filter(lifeform == "grass") %>% 
    pull(species))))
#2


####### forb
length(sort(unique(
  belt %>% 
  filter(status == "native") %>% 
  filter(lifeform == "forb") %>% 
    pull(species)))) - 4 # remove four "spp." -->  "Galium_spp."     "Lupinus_spp."    "Phacelia_spp."  "Trifolium_spp." 
# 48

# and one fern 
```

# NN lifeforms
```{r}

####### grass
length(sort(unique(
  belt %>% 
  filter(status == "nonnative") %>% 
  filter(lifeform == "grass") %>% 
    pull(species))))
# 5

####### forb
length(sort(unique(
  belt %>% 
  filter(status == "nonnative") %>% 
  filter(lifeform == "forb") %>% 
    pull(species))))
# 12


```


# sampling effort

```{r}

# Calculate species accumulation
species_accum <- specaccum(species_data)

# Plot the species accumulation curve
plot(species_accum, xlab = "Sampling Sites", ylab = "Species Richness", main = "Species Accumulation Curve")

# Add confidence intervals
plot(species_accum, ci.type="polygon", col="blue", ci.col="lightblue", xlab = "Sampling Sites", ylab = "Species Richness", main = "Species Accumulation Curve with CI")

# Add species richness estimator (Chao, Jackknife, etc.)
chao <- estimateR(belt.spp_2019)
chao

```




END


overall native richness
- sppr
```{r native species richness, include=FALSE}
native_spp <- ab_totalsppr %>% 
   filter(status == "native") # %>% 
  # filter(species! = "Lupinus spp.") # unselect spp.
  # 

sort(unique(native_spp$species)) # 39 species
length(unique(native_spp$species)) -1 # for Lupinus spp. ;  38 species
```

- family richness
```{r transect native family richness}
sort(unique(native_spp$family))
length(unique(native_spp$family))  # 24 families
```

overall non-native richness
- sppr
```{r nnnative species richness}
nn_sppr <- ab_totalsppr %>% 
   filter(status == "nonnative") 

sort(unique(nn_sppr$species))
length(unique(nn_sppr$species))-2 # for Bromus spp. and "Erodium spp." ; 10 species
```

- family richness
```{r transect native family richness}
sort(unique(nn_sppr$family))
length(unique(nn_sppr$family)) # 4 families
```
 stats - categorical by hillslope (n = 15)
- data prep
```{r}
ab_totalsppr_hillslope <- ab_totalsppr %>%  # n = 15
  group_by(standtype, site, site_rep) %>% 
  summarize(sppr = n_distinct(site_rep_species)) 
  

sppr_bystatus <- ab_totalsppr %>% # n = 33
  group_by(standtype, site, site_rep, status) %>% 
  summarize(sppr = n_distinct(site_rep_species))
```

- normality 
```{r}
# data vis
length(ab_totalsppr_hillslope$sppr) # n = 15
hist(ab_totalsppr_hillslope$sppr) 

min(ab_totalsppr_hillslope$sppr) # 10 species
max(ab_totalsppr_hillslope$sppr) # 22 species

shapiro.test(ab_totalsppr_hillslope$sppr)
# Shapiro-Wilk normality test
# 
# data:  ab_totalsppr_hillslope$sppr
# W = 0.91175, p-value = 0.1441 <-- normally distributed
```

 - anova - hillslope
```{r}
summary(aov(sppr ~ site, ab_totalsppr_hillslope))
#  Df Sum Sq Mean Sq F value   Pr(>F)    
# site         6 223.60   37.27   31.94 3.62e-05 *** <-- sig
# Residuals    8   9.33    1.17                      

TukeyHSD(aov(sppr ~ site, ab_totalsppr_hillslope))
# Fit: aov(formula = sppr ~ site, data = ab_totalsppr_hillslope)
# 
# $site
#               diff          lwr       upr     p adj
# SD2-SD1   2.000000  -3.83171698  7.831717 0.8311488
# SD3-SD1  -3.000000  -8.83171698  2.831717 0.4961437
# MX1-SD1  -1.000000  -5.76157698  3.761577 0.9780860
# MX2-SD1  -9.666667 -14.42824364 -4.905090 0.0006579 <-- MX2-SD1
# GD1-SD1  -8.000000 -12.76157698 -3.238423 0.0023942 <--GD1-SD1  0.0023942
# GD2-SD1  -6.333333 -11.09491031 -1.571756 0.0104732 <--GD2-SD1 

# SD3-SD2  -5.000000 -10.83171698  0.831717 0.1015938
# MX1-SD2  -3.000000  -7.76157698  1.761577 0.3046200
# MX2-SD2 -11.666667 -16.42824364 -6.905090 0.0001711 <--MX2-SD2
# GD1-SD2 -10.000000 -14.76157698 -5.238423 0.0005182 <--GD1-SD2 0.0005182
# GD2-SD2  -8.333333 -13.09491031 -3.571756 0.0018235 <--GD2-SD2

# MX1-SD3   2.000000  -2.76157698  6.761577 0.6853253
# MX2-SD3  -6.666667 -11.42824364 -1.905090 0.0076753 <--MX2-SD3
# GD1-SD3  -5.000000  -9.76157698 -0.238423 0.0390970 <--GD1-SD3 0.0390970
# GD2-SD3  -3.333333  -8.09491031  1.428244 0.2198423

# MX2-MX1  -8.666667 -12.03361004 -5.299723 0.0001192 <-- MX2-MX1
# GD1-MX1  -7.000000 -10.36694337 -3.633057 0.0005566 <--GD1-MX1
# GD2-MX1  -5.333333  -8.70027670 -1.966390 0.0035214 <--GD2-MX1

# GD1-MX2   1.666667  -1.70027670  5.033610 0.5337581
# GD2-MX2   3.333333  -0.03361004  6.700277 0.0525268 <-- 0.053

# GD2-GD1   1.666667  -1.70027670  5.033610 0.5337581

# SUMMARY
# GD1 is sig diff from INT hillslopes
# GD2 is sig diff from INT1 and 2 and MX1 but not INT3

# MX2 is sig diff from INT and MX1 hillslopes




```

```{r}
ggplot(ab_totalsppr_hillslope) +
  stat_summary(
     mapping = aes(x = site, y = sppr),
    fun.min = min,
    fun.max = max,
    fun = median
  )+
  ylim(0,25)

ggplot(data = ab_totalsppr_hillslope, 
       mapping = aes(x = site, y = sppr)) + 
  geom_boxplot() +
  ylim(0,25)

boxplot(sppr ~ site, data = ab_totalsppr_hillslope)
```


```{r}
sppr_lm <- ggplot(data = ab_totalsppr_hillslope, 
       aes(x = site, y = sppr)) +
  geom_point(aes(color = site, shape = site), size = 3, alpha = 0.8) +
  scale_color_manual(values = c( "lightblue", "cornflowerblue", "blue","green", "darkgreen", "orange", "brown")) + #"gray0",
   scale_shape_manual(values = c(17, 17, 17, 15, 15, 16,16)) + 
   ylim(0, 25) +
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", color = "black") +  # lm, loess, glm
  labs(x = "Hillslope", 
       y = "Species richness (#)", 
       title = "Species richness by hillslope") +
  
   stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() 
sppr_lm

sppr_bystatus_lm <- ggplot(data = sppr_bystatus, 
       aes(x = site, y = sppr)) +
    geom_point(aes(color = site, shape = status), size = 3, alpha = 0.8, stroke = 2) +
   scale_fill_manual(values = c("gray", "white", "black")) +  # Fill colors for each status
  scale_color_manual(values = c( "lightblue", "cornflowerblue", "blue","green", "darkgreen", "orange", "brown")) + #"gray0",
 #  scale_shape_manual(values = c(17, 17, 17, 15, 15, 16,16)) + 
   ylim(0, 25) +
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", color = "black") +  # lm, loess, glm
  labs(x = "Hillslope", 
       y = "Species richness (#)", 
       title = "Native and Non-native species richness by hillslope") +
   stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() 
sppr_bystatus_lm


sppr_bystatus$site_rep <- factor(sppr_bystatus$site_rep, levels=c("Intact1_1", "Intact2_2", "Intact3_3", "ENH1_1", "ENH1_2", "ENH1_3", "ENH2_1", "ENH2_2", "ENH2_3", "DEG1_1", "DEG1_2", "DEG1_3", "DEG2_1", "DEG2_2", "DEG2_3"))

ggplot(data = sppr_bystatus, 
       aes(x = site_rep, y = sppr)) +
    geom_point(aes(color = site, shape = status), size = 3, alpha = 0.8) +
  scale_color_manual(values = c( "lightblue", "cornflowerblue", "blue","green", "darkgreen", "orange", "brown")) + #"gray0",
  # scale_shape_manual(values = c(17, 17, 17, 15, 15, 16,16)) + 
   ylim(0, 25) +
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", color = "black") +  # lm, loess, glm
  labs(x = "Hillslope", 
       y = "Species richness (#)", 
       title = "Native and Non-native species richness by transect") +
   stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) # turn x-axis labels 90 degrees

```

 - anova - status x hillslope

- normality - native
```{r native anova sppr}
# data vis
length(sppr_bystatus$sppr) # n = 33 = 15 x 2 + 3 unknowns

hist((sppr_bystatus %>% 
       filter(status == "native"))$
       sppr) 

min((sppr_bystatus %>% 
       filter(status == "native"))$
       sppr) # 3 native species
max((sppr_bystatus %>% 
       filter(status == "native"))$
       sppr) # 18 native species

shapiro.test((sppr_bystatus %>% 
       filter(status == "native"))$
       sppr)
# Shapiro-Wilk normality test
# 
# data:  (sppr_bystatus %>% filter(status == "native"))$sppr
# W = 0.96634, p-value = 0.8007 <-- normaly distributed


summary(aov(sppr ~ site, sppr_bystatus %>% 
       filter(status == "native")))
# Df Sum Sq Mean Sq F value   Pr(>F)    
# site         6  264.9   44.16   35.32 2.47e-05 ***
# Residuals    8   10.0    1.25   

TukeyHSD(aov(sppr ~ site, sppr_bystatus %>% 
       filter(status == "native")))
# $site
#                diff         lwr        upr     p adj
# SD2-SD1   2.0000000  -4.0364006  8.0364006 0.8502944 <-- SDs are all similar 
# SD3-SD1  -3.0000000  -9.0364006  3.0364006 0.5298787
# SD3-SD2  -5.0000000 -11.0364006  1.0364006 0.1174050
# 
# MX1-SD1  -2.3333333  -7.2620338  2.5953671 0.5767649
# MX1-SD2  -4.3333333  -9.2620338  0.5953671 0.0911244
# MX1-SD3   0.6666667  -4.2620338  5.5953671 0.9977271.      SD3 and MX1 are the same
# 
# MX2-SD1 -11.6666667 -16.5953671 -6.7379662 0.0002199 <-- SD1 greater than MX2, GD1 and 2
# GD1-SD1  -8.3333333 -13.2620338 -3.4046329 0.0022958 <--
# GD2-SD1  -7.0000000 -11.9287004 -2.0712996 0.0070286 <--
# 


# 
# MX2-SD2 -13.6666667 -18.5953671 -8.7379662 0.0000686 <-- SD2 greater than MX2, GD1 and 2
# GD1-SD2 -10.3333333 -15.2620338 -5.4046329 0.0005245 <--
# GD2-SD2  -9.0000000 -13.9287004 -4.0712996 0.0013684 <--
# 

# MX2-SD3  -8.6666667 -13.5953671 -3.7379662 0.0017665 <-- SD3 greater than MX2 and GD1 BUT NOT diff than GD2
# GD1-SD3  -5.3333333 -10.2620338 -0.4046329 0.0334505 <--
# GD2-SD3  -4.0000000  -8.9287004  0.9287004 0.1275715 
# 
# MX2-MX1  -9.3333333 -12.8184508 -5.8482158 0.0000889 <-- MX1 greater than MX2 and GD1 and 2
# GD1-MX1  -6.0000000  -9.4851175 -2.5148825 0.0020356 <--
# GD2-MX1  -4.6666667  -8.1517842 -1.1815492 0.0100614 <--
# 
# GD1-MX2   3.3333333  -0.1517842  6.8184508 0.0620313 <-- 0.06
# GD2-MX2   4.6666667   1.1815492  8.1517842 0.0100614 <-- MX2 less than GD2
# GD2-GD1   1.3333333  -2.1517842  4.8184508 0.7593849

boxplot(sppr ~ site, data = sppr_bystatus %>% 
       filter(status == "native"))
```

- normality - nn
```{r nn anova sppr}
hist((sppr_bystatus %>% 
       filter(status == "nonnative"))$
       sppr) 
min((sppr_bystatus %>% 
       filter(status == "nonnative"))$
       sppr) # 4 nn species
max((sppr_bystatus %>% 
       filter(status == "nonnative"))$
       sppr) # 6 nn species

shapiro.test((sppr_bystatus %>% 
       filter(status == "nonnative"))$
       sppr)
# Shapiro-Wilk normality test
# 
# data:  (sppr_bystatus %>% filter(status == "nonnative"))$sppr
# W = 0.78295, p-value = 0.00224 <-- not normally distributed
```




```{r}


# Renaming observations
ab_diversity_load <- abcover_full_raw %>%
  #dplyr::select(!c(species == "bare_ground")) 
   mutate(species = case_when(
    species == "Avena barbata dead" ~ "Avena barbata", # old name ~ new name
        species == "Bromus diandrus dead" ~ "Bromus diandrus", 
           species == "Bromus rubens dead" ~ "Bromus rubens", 
       species == "Bromus spp." ~ "Bromus rubens", 
    species == "Bromus spp. dead" ~ "Bromus rubens",
          species == "Bromus madritensis" ~ "Bromus rubens",
     species == "Centaurea melitensis dead" ~ "Centaurea melitensis",
    species == "Erodium cicutarium" ~ "Erodium spp.", 
 #   species == "Elymus condensatus dead" ~ "Elymus condensatus",
 #   species == "Heteromeles arbutifolia dead" ~ "Heteromeles arbutifolia",
    species == "Hirschfeldia incana dead" ~ "Hirschfeldia incana",
 #   species == "Malacothamnus fasciculatus dead" ~ "Malacothamnus fasciculatus",
      #     species == "Salvia leucophylla dead" ~ "Salvia leucophylla",
     # species == "Stipa lepida dead" ~ "Stipa lepida",
    TRUE ~ as.character(species)  # Keep other observations unchanged
  )) %>% 
  mutate(site = case_when(
    site == "Intact_1" ~ "Intact1",
      site == "Intact_2" ~ "Intact2",
      site == "Intact_3" ~ "Intact3",
    TRUE ~ as.character(site)  # Keep other observations unchanged
  )) %>% 
  mutate(standtype = fct_relevel(standtype, 
            "intact", "matrix", "degraded")) %>% 
   mutate(site = fct_relevel(site, 
            "Intact1", "Intact2", "Intact3", "ENH1",    "ENH2", "DEG1",    "DEG2")) %>% 
  filter(species !="native_litter" ) %>% 
  filter(species != "nonnative_thatch") %>% 
  filter(species != "bare_ground") 

sort(unique(ab_diversity_load$species))
sort(unique(ab_diversity_load$site))
glimpse(ab_diversity_load)
#Rows: 1,450
#Columns: 10

# combine site and rep to one column
ab_diversity_load$site_rep <- paste(ab_diversity_load$site, ab_diversity_load$rep, sep="_") 
ab_diversity_load$site_rep_species <- paste(ab_diversity_load$site_rep, ab_diversity_load$species, sep="_") 
sort(unique(ab_diversity_load$site_rep)) 
#"DEG1_1"    "DEG1_2"    "DEG1_3"    "DEG2_4"    "DEG2_5"    "DEG2_6"    
#"ENH1_1"    "ENH1_2"    "ENH1_3"    "ENH2_4"    "ENH2_5"    "ENH2_6"    
#"Intact1_1" "Intact2_2" "Intact3_3"

glimpse(ab_diversity_load)  # species diversity from point-line intercept transect
# Rows: 1,445
#Columns: 10

ab_diversity <- ab_diversity_load %>% 
  dplyr::select(!c(distance, vertical))

#write.csv(ab_diversity, "processed/ab_diversity.csv")
```

 Belt transect data
 load data
```{r SHRUB species richness by belt transect, include = TRUE}
belttrans_byspecies <- read.csv("deg_beltdensity_byspecies.csv")
sort(unique(belttrans_byspecies$species))
sort(unique(belttrans_byspecies$site))
sort(unique(belttrans_byspecies$origin))
sort(unique(belttrans_byspecies$lifespan))

belttrans_sppr <- belttrans_byspecies %>%  
   mutate(species = case_when(
    species == "Bromus madritensis" ~ "Bromus rubens", # old name ~ new name
        species == "Erodium cicutarium" ~ "Erodium spp.",      
#    species == "Lupinus spp." ~ "Lupinus bicolor", # lupins are in different transects so leaving as three seperate entries: LUBI, LUSU, and Lupinus spp.
#     species == "Lupinus spp." ~ "Lupinus succulentus",
    TRUE ~ as.character(species)  # Keep other observations unchanged
  )) %>% 
      mutate(origin = case_when(
        origin == "non-native" ~ "nonnative", 
        TRUE ~ as.character(origin))) %>% 
  mutate(lifespan = case_when(
    lifespan == "Perennial" ~ "perennial", 
     TRUE ~ as.character(lifespan))) %>% 
    rename(rep = transectstand, #rename(new_name = old_name) 
           status = origin) %>% 
   filter(site != "ENH2") # why am I omitting ENH2? (feb 7, 2024) <-- STEPHANIE

belttrans_sppr$site <- factor(belttrans_sppr$site, levels=c("Intact1", "Intact2", "Intact3", "ENH1", "DEG1", "DEG2"))
unique(belttrans_sppr$site)
unique(belttrans_sppr$species)
  
# combine site and rep to one column
belttrans_sppr$site_rep <- paste(belttrans_sppr$site, belttrans_sppr$rep, sep="_") 
belttrans_sppr$site_rep_species <- paste(belttrans_sppr$site_rep, belttrans_sppr$species, sep="_") 

belttrans_sppr <- belttrans_sppr %>% 
  dplyr::select(!c("transectsite", "lifestage")) 
sort(unique(belttrans_sppr$site))

glimpse(belttrans_sppr)
sort(unique(belttrans_sppr$site_rep))
sort(unique(belttrans_sppr$site))
sort(unique(belttrans_sppr$species))



write.csv(belttrans_sppr, "processed/belttrans_sppr.csv")
```


```{r belt transect join percent cover, include=TRUE}
belttrans_sppr_native <- belttrans_sppr %>% # join above ground species diversity with native transect percent cover
  left_join(nativecover3, by = "site_rep") 
sort(unique(belttrans_sppr_native$site_rep))

#view(belttrans_sppr_native_nn)
glimpse(belttrans_sppr_native)
# Rows: 904
# Columns: 21 

#write.csv(belttrans_sppr_native_nn, "processed/belttrans_sppr_native_nn.csv") #<-- export to EXCEL to create belt transect sppr figures.

```

```{r join sppr from belt and point-intercept transect}
# nov. 6, 2023
# KIT - Can you create a df of the total species list (point intercept transect + belt transect species list) at INT 1-3, ENH 1-3, and DEG 1-6?

belt_sppr <- belttrans_sppr %>% #select only site_rep_species from belt trans data
  dplyr::select(c(site_rep_species)) 

pointintercept_sppr <- ab_diversity_native_nn %>% #select only site_rep_species from belt trans data
  dplyr::select(c(site_rep_species))

belt_point_join_sppr <- belt_sppr %>% 
  full_join(pointintercept_sppr, by = "site_rep_species")

belt_point_join_sppr %>% 
  summarize(total_sppr = n_distinct(site_rep_species))
# ahhh.... standtype, site_rep, status, etc. needs to be brougt in from somewhere.... 


```


 Join transect and belt sppr.
```{r species richness above ground, include=TRUE}

# created ab_diversity_2 in EXCEL

ab_diversity_2 <- read.csv("processed/ab_diversity_2.csv", header = TRUE, na.strings=c("","NA")) # combined point-intercept transect and belt transect species diversity
nativecover <- read.csv("processed/abcover_nativecover_nngrasscover.csv") # native percent cover from ch1_standdeg_canopy.Rmd
nncover <- read.csv("processed/abcover_nonnativecover_nngrasscover.csv") # non-native percent cover from ch1_standdeg_canopy.Rmd

ab_diversity_2 <- ab_diversity_2 %>% 
  mutate(site = case_when(
    site == "Intact_1" ~ "Intact1",
      site == "Intact_2" ~ "Intact2",
      site == "Intact_3" ~ "Intact3",
    TRUE ~ as.character(site)  # Keep other observations unchanged
  )) %>% 
  mutate(standtype = fct_relevel(standtype, 
            "intact", "matrix", "degraded")) %>% 
   mutate(site = fct_relevel(site, 
            "Intact1", "Intact2", "Intact3", "ENH1",    "ENH2", "DEG1",    "DEG2")) %>% 
   na.omit() 

view(ab_diversity_2)
glimpse(ab_diversity_2)
# Rows: 2,018
# Columns: 9

nativecover <- nativecover %>% 
  dplyr::select(!c(X, standtype.x, standtype.y ))
nncover <- nncover %>% 
  dplyr::select(c(site_rep, percentcover_nonnative))

ab_diversity_2_native <- ab_diversity_2 %>% # join above ground species diversity with native transect percent cover
  left_join(nativecover, by = "site_rep")
ab_diversity_2_native_nn <- ab_diversity_2_native %>% # join above ground species diversity with non-native transect percent cover
  left_join(nncover, by = "site_rep")

view(ab_diversity_2_native_nn)
glimpse(ab_diversity_2_native_nn)
# Rows: 2,018
# Columns: 20

```

```{r species richness by stand type}

ab_diversity_2_native_nn %>% 
  filter(status =="native") %>% 
distinct(species) %>%
  pull(species)
#  [1] "Acmispon glaber"            "Acmispon maritimus"        
#  [3] "Apiastrum angustifolium"    "Artemisia californica"     
#  [5] "Baccharis pilularis"        "Calochortus clavatus"      
#  [7] "Cenothus oliganthus"        "Claytonia perfoliata"      
#  [9] "Dipterostemon capitatus"    "Elymus condensatus"        
# [11] "Eriodictyon crassifolium"   "Eriogonum fasciculatum"    
# [13] "Galium porrigens"           "Heteromeles arbutifolia"   
# [15] "Lupinus spp."               "Malacothamnus fasciculatus"
# [17] "Marah macrocarpa"           "Pellaea andromedifolia"    
# [19] "Rhamnus ilicifolia"         "Rhus ovata"                
# [21] "Salvia apiana"              "Salvia leucophylla"        
# [23] "Sambucus mexicana"          "Solanum xanti"             
# [25] "Stephanomeria virgata"      "Stipa lepida"  

ab_diversity_2_native_nn %>% 
  filter(status =="native") %>% 
distinct(family) %>%
  pull(family)
#  [1] "Fabaceae"             "Apiaceae"             "Asteraceae"          
#  [4] "Liliaceae"               "Montiaceae"          
#  [6] "Themidaceae"          "Poaceae"              "Namaceae"            
#  [9] "Rubiaceae"            "Rosaceae"            
# [11] "Malvaceae"            "Cucurbitaceae"        "Pteridaceae"         
# [14] "Rhamnaceae"           "Anacardiaceae"        "Lamiaceae"           
#  [17] "Solanacaeae" 

ab_diversity_2_native_nn %>% 
  filter(status =="nonnative") %>% 
distinct(species) %>%
  pull(species)
# [1] "Avena barbata"               "Bromus diandrus"            
#  [3] "Bromus hordeaceus"           "Bromus madritensis"         
#  [5] "Bromus rubens"               "Carduus pycnocephalus"      
#  [7] "Centaurea melitensis"        "Erodium spp."               
#  [9] "Hirschfeldia incana"         "Pseudognaphalium luteoalbum"

ab_diversity_2_native_nn %>% 
  filter(status =="nonnative") %>% 
distinct(family) %>%
  pull(family
    )
# [1] "Poaceae"      "Asteraceae"   "Geraniaceae"  "Brassicaceae"

ab_diversity_2_native_nn <- ab_diversity_2_native_nn %>% 
  dplyr::filter(species != "moss") %>% 
   dplyr::filter(species != "nonnative_thatch") %>% 
dplyr::filter(species != "native_litter") 

# combine site and site_rep_species to one column
ab_diversity_2_native_nn$site_rep_species <- paste(ab_diversity_2_native_nn$site_rep, ab_diversity_2_native_nn$species, sep="_") 

ab_diversity_2_native_nn %>% 
  distinct(species) %>%
  pull(species)


ab_div_bysiterep <- ab_diversity_2_native_nn %>% # number of unique species per transect
  group_by(standtype, site_rep) %>%  
  dplyr::summarize(num_species = length(unique(species)))
print(ab_div_bysiterep) 
```


```{r check - number of unique species per transect, include=TRUE}

ab_diversity_2_native_nn %>% 
   dplyr::filter(site_rep == "DEG1_1") %>% 
  distinct(species) %>%
  pull(species)
  
```

normal
```{r}
shapiro.test(ab_div_bysiterep$num_species)
# data:  ab_div_bysiterep$num_species
# W = 0.96712, p-value = 0.742 <-- normally distrubuted
```

figure regressions by Kit
```{r scatterplot species diversity by nn grass cover, include=TRUE} 

# KIT's code: this section

ab_diversity_2_native_nn_plot <- ab_diversity_2_native_nn %>%
  group_by(site_rep) %>%
  summarize(richness = n_distinct(species))

hold <- rep(c("DEG1_1", "DEG1_2", "DEG1_3", "DEG2_4", "DEG2_5", "DEG2_6", "ENH1_1", "ENH1_2", "ENH1_3", "ENH2_4", "ENH2_5", "ENH2_6", "Intact_1_1", "Intact_2_2", "Intact_3_3"), times = 3)

hold5 <- rep(c("DEG", "DEG", "DEG", "DEG", "DEG", "DEG", "ENH", "ENH", "ENH", "ENH", "ENH", "ENH", "Intact", "Intact", "Intact"), times = 3)

hold2 <- rep(c("total", "native", "nn"), each = 15)

hold3 <- rep(c(90,98,98,98,90,100,83,83,88,46,66,76,12,7,24), times = 3)

hold4 <- c(10,6,7,7,7,9,8,8,10,14,10,12,6,12,9,5,2,2,3,3,5,3,3,5,10,6,6,5,10,6,5,4,5,4,4,4,5,5,5,4,3,6,1,1,3)

abcover_diversity_plot <- data.frame(site_rep = hold, standtype = hold5, richness = hold4, nngrass = hold3, type = hold2)
  
# Stephanie isn't sure what this code is about... ?
# ab_diversity_2_native_nn_2 <- ab_diversity_2_native_nn %>%
#   left_join(ab_diversity_2_native_nn_plot, by = "site_rep")


ab_sppr_nngrass <- ggplot(data = abcover_diversity_plot, 
       aes(x = nngrass, y = richness, shape = type, color = type)) +
  geom_point(size = 3, alpha = 0.8) +
  scale_shape_manual(values = c(16, 17, 18)) +
  ylim(0, 20)+
  xlim(0,100) +
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", color = "black") +  # lm, loess, glm
  scale_color_manual(values = c(standcolors),
                     name = "Stand Type",
                     labels = c("intact" = "shrub dominated", 
                                "matrix" = "matrix", 
                                "degraded" = "grass dominated")) + #rename standtypes labels
  labs(x = "Non-native grass cover (%)", 
       y = "Species richness (#)", 
       title = "Species richness x non-native grass cover") +
  theme_bw() +  
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2")))

print(ab_sppr_nngrass)
```

stats
```{r species richness df from An code}
sppr <- specnumber(abcover_byspecies_nmdsprep_num)
sppr
# 11 15  7 15 13 16 11 13 13 10  8  9 12 10 12

abcover_byspecies_nmdsprep <- abcover_byspecies_nmdsprep %>% 
  mutate(standtype_ID = 1)

abcover_byspecies_nmdsprep$standtype_ID <- 1:15

sppr_df <- sppr %>% 
  enframe() %>% 
  full_join(abcover_byspecies_nmdsprep, by = c("name" = "standtype_ID")) %>% 
    rename(spp_rich = value)
view(sppr_df)
```

```{r stats spp richness }

shapiro.test(sppr_df$spp_rich)
# data:  sppr_df$spp_rich
# W = 0.97374, p-value = 0.9091 <-- normally distributed

summary(aov(spp_rich ~ standtype, data = sppr_df)) # by stand type

#             Df Sum Sq Mean Sq F value Pr(>F)  
# standtype    2  35.00  17.500   3.481 0.0643 .
# Residuals   12  60.33   5.028  

summary(aov(spp_rich ~ site, data = sppr_df)) # by hillslope (aka. site)
#             Df Sum Sq Mean Sq F value  Pr(>F)   
# site         6  83.33   13.89   9.259 0.00305 **
# Residuals    8  12.00    1.50 

TukeyHSD(aov(spp_rich ~ site, data = sppr_df))
histogram(sppr_df$spp_rich)

boxplot(spp_rich~site,data=sppr_df, main="Species richness by Hillslope",
   xlab="Hillslope", ylab="Species Richness (N)")
```

```{r species richness figure from An code}
#pal <- c("lightsalmon1", "gold1", "palegreen4")

plot_sppr <- ggplot(sppr_df, aes(x = site, y = spp_rich, fill = site)) +
  geom_boxplot() +
  #scale_fill_manual(spp_rich = pal) +
#  scale_x_discrete(labels = c("Intact \n (n = 9)", "Matrix \n (n = 6)", "Degraded \n (n = 6)")) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = "Hillslope",
       y = "Number of species per hillslope",
       title = "Above ground species richness")
plot_sppr

```





# USEFUL CODE
- names() = column titles
- head()
- tail()
- sort(unique(df$column))
- class() = tibble or data.frame?

f <-  belt %>% 
  filter(status == "native") %>% 
  filter(lifeform == "forb") %>% 
   filter(grepl("spp\\.", species))
sort(unique(f$species)) 

- Shannon Diversity
```{r shannon diversity}
# this doesn't make sense for % cover, but I could run it on shrub density. I could also run it on the seedbank data. 
shannondiv <- diversity(abcover_byspecies_nmdsprep_num)
head(shannondiv)

sppdiv_aov <- aov(shannondiv ~ standtype, data = abcover_byspecies_nmdsprep)
summary(sppdiv_aov)

#     Df Sum Sq Mean Sq F value Pr(>F)  
#standtype    2 0.2899 0.14493   3.924 0.0489 *
#Residuals   12 0.4433 0.03694        

```

# old code

```{r above ground - species richness - native richness, include = FALSE}

### SPECIES RICHNESS DATA IS IN CH1_DEGRADATION_ABCOVERNMDS.RMD FILE ###






# species richness, native richness
abrich_full <- read.csv("~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022b/deg_abrich.csv", header = TRUE, na.strings=c("","NA")) %>% 
  dplyr::select('trasecttype', 'standtype', 'site', 'rep', 'distance', 'vertical', 'species', 'lifeform', 'status', 'lifecycle', 'Family', 'lifestage') %>% 
  rename(transecttype = trasecttype, 
         family = Family)
glimpse(abrich_full)


```
