---
title: "transect_continuous"
author: "Stephanie Ma Lucero"
date: "2024-04-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Code to hang on to
- create a new third column by: mutate(site_rep = paste(site, transect, sep = "_")) %>% # create a new third column
- combine two columns into one column: unite("lifeform_total", 2:3, sep = "_") # consolidate columns
- line types
-- "solid": A solid line.
-- "dashed": A dashed line.
-- "dotted": A dotted line.
-- "dotdash": Alternating dots and dashes.
-- "longdash": Long dashes.

```{r libraries, message=FALSE}
library(here) # similar to set working directory
library(tidyverse) # data wrangling - includes ggplot2, dplyr, tidyr, readr, purr, tibble, stringr, forcats
library(readxl) 
```
# colors
```{r colors, include=TRUE}
devtools::install_github("an-bui/calecopal")

library(calecopal)
# all palettes
names(cal_palettes)

#chaparral1
#chaparral2
cal_palette(name = "chaparral3", n = 20, type = "continuous") 

#species colors
HEAR <- "#D3E3CA"
RHOV <- "#92A587"
RHIL <- "#2F3525"

MAFA <- "gray60"
SALE <- "gray40"
ARCA <- "gray20"

scl <- 16
css <- 17

colors_scl <- c(HEAR, RHOV, RHIL)
colors_css <- c(MAFA, SALE, ARCA)

# Lifeform colors: shrub, grass, forb
shrub <- "green4"
grass <- "yellow2"
forb <- "purple2"
fern <- "black"
colors_lifeform <- c(shrub, grass, forb, fern)

# Status colors: native, exotic 
native <- "black"
nonnative <- "maroon"
unknown <- "lightblue"
colors_status <- c(native, nonnative, unknown)

# "orange", "lightblue", "limegreen", "deeppink","mediumpurple2", "darkgreen"
```

#NOTES Transect data overview - Years collected: 2017, 2019, 2024 - Sites: 1 and 2 - Comparing to intact sites: INT1, INT2, INT3, INT4, INT5

# load data

```{r load point intercept transect data from csv, include=TRUE}
file_transect_2019_load <- read_excel("deg_percentcover_byspecies_bylifeform_2024.xlsx", sheet = "2019_data")
# -- years 2017, 2019
# -- Site 1: transects 1-6
# -- Site 2: transects 1-6
# -- INT1, INT2, INT3, INT4

# Finding species == "Bromus spp." to rename them by species
# transect2019_bromus <- file_transect_2019_load %>%
#   filter(grepl("Bromus", species))
# 
# transect2019_bromusspp <- file_transect_2019_load %>% 
#   filter(species == "Bromus spp.")

transect2019 <- file_transect_2019_load %>% 
  dplyr::select(c(site, transect, year, distance, vertical, species)) %>% # select columns to full join with 2024 data
  filter(year %in% c(2017, 2019))  %>% #filter data to 2019 data
   dplyr::select(year, site, transect, distance, vertical, species) %>% 
   mutate(species = ifelse(# a condition, a value if the condition is true, and a value if the condition is false
                              site == 2 & transect == 5 & distance %in% c(2, 6, 9) & species == "Bromus spp.", "Bromus hordeaceus", species)) %>% 
    mutate(species = ifelse(site == 2 & transect == 5 & distance %in% c(10, 28) & species == "Bromus spp.", "Bromus rubens", species)) %>% 
    mutate(species = ifelse(site == 2 & transect == 6 & distance %in% c(23, 35, 37) & species == "Bromus spp.", "Bromus hordeaceus", species)) %>% 
    mutate(species = ifelse(site == "INT1" & transect == 1 & distance %in% c(24) & species == "Bromus spp.", "Bromus diandrus", species))



  
# View the data (optional)
#View(transect2019)
#sort(unique(transect2019$transect))
#sort(unique(transect2019$site))

file_transect_2024_load <- read.csv("Transects_2024 - transect.csv", header = TRUE, na.strings=c("","NA"))
# -- years 2024
# -- Site 1: transects 7-14
# -- Site 2: transects 7-21
# -- INT5
transect2024 <- file_transect_2024_load %>%  
  dplyr::select(c(site, transect, year, distance, vertical, species)) %>% 
  filter(species != "N/A") %>% 
   mutate(species = case_when(species == "Bromus spp." ~ "Bromus rubens", # oldname ~ newname
                              TRUE ~ species))
  
# View the data (optional)
#view(transect2024)
#sort(unique(transect2024$transect))
#sort(unique(transect2024$site))
```

## full join transect data

```{r full join}
abcover_fulljoin <- full_join(transect2019, transect2024, by = intersect(names(transect2019), names(transect2024)))

# View the data (optional)
#View(abcover_full)
#sort(unique(abcover_full$transect))
#sort(unique(abcover_full$site))
```

## species list

```{r convert to tibble}
# convert dataframe to tibble
abcover_fulljoin_tib <- as_tibble(abcover_fulljoin) 
class(abcover_fulljoin_tib)

abcover_full <- abcover_fulljoin_tib %>% 
   mutate(site_rep = paste(site, transect, sep = "_")) %>% # combine site and rep to one column
   mutate(site_rep_dist = paste(site_rep, distance, sep = "_")) %>%  # combine site_rep and distance to one column 
   mutate(species = case_when(species == "BG bare ground" ~ "bare_ground", # oldname ~ newname
                              species == "ground" ~ "bare_ground", 
                              species == "native_litter" ~ "litter",
                              species == "nonnative_thatch" ~ "thatch",
                              species == "Bromus diandrus BRDI" ~ "Bromus diandrus",
                              species == "Bromus rubens BRRU" ~ "Bromus rubens", 
                              species == "Bromus hordeaceus BRHO" ~ "Bromus hordeaceus", 
                              species == "Rapistrum rugosum - round pods" ~ "Rapistrum rugosum", 
                              species == "HIIN Hirschfeldia incana - long pods" ~ "Hirschfeldia incana", 
                              species == "Hirschfeldia incana - long pods" ~ "Hirschfeldia incana", 
                              species == "Salvia lucophylla" ~ "Salvia leucophylla",
                              species == "Malacothamnus fasciculatus " ~ "Malacothamnus fasciculatus",
                              species == "dead HEAR" ~ "Heteromeles arbutifolia dead",
                              species == "dead MAFA" ~ "Malacothamnus fasciculatus dead",
                              species == "dead RHIL" ~ "Rhamnus ilicifolia dead",
                              species == "dead ARCA" ~ "Artemisia californica dead",
                              species == "dead SALE" ~ "Salvia leucophylla dead",
                              species == "ACGL Acmispon glaber" ~ "Acmispon glaber",
                              species == "dead Acmispon glaber" ~ "Acmispon glaber dead",
                              species == "Acmispon maritimus - round, sevfl" ~ "Acmispon maritimus",
                              species == "Acmispon strigosus - oval, sevfl" ~ "Acmispon strigosus",
                              species == "Sambucus nigra" ~ "Sambucus mexicana",
                              species == "Blitum californicum" ~ "Chenopodium californicum",
                              species == "Cryptantha sp." ~ "Cryptantha spp.",
                              TRUE ~ species)) %>% 
   mutate(
      year = as.character(year),
      site = as.character(site),
      transect = as.character(transect),
      distance = as.character(distance),
      vertical = as.character(vertical),
      species = as.character(species)
      )

glimpse(abcover_full) # look at new columns

sort(unique(abcover_full$species))
```

## species characteristics

```{r load species characteristics, include=TRUE}
specieslist <- file_transect_2019_load %>%
  dplyr::select(species, lifeform, status, lifecycle, family) %>%
   mutate(status = case_when(status == "Native" ~ "native", # oldname ~ newname
                                status == "Non-native" ~ "nonnative",
                                status == "thatch" ~ "ground",
                                status == "litter" ~ "ground",
                                status == "rock" ~ "ground",
                                TRUE ~ status)) %>%
   mutate(lifeform = case_when(lifeform == "Grass" ~ "grass",
                                lifeform == "Shrub" ~ "shrub",
                                lifeform == "Forb" ~ "forb",
                                lifeform == "thatch" ~ "ground",
                                lifeform == "litter" ~ "ground",
                                lifeform == "rock" ~ "ground",
                                TRUE ~ lifeform)) %>%
   mutate(lifecycle = case_when(lifecycle == "Annual" ~ "annual",
                                lifecycle == "Perennial" ~ "perennial",
                                lifecycle == "thatch" ~ "ground",
                                lifecycle == "litter" ~ "ground",
                                lifecycle == "rock" ~ "ground",
                                TRUE ~ lifecycle)) %>%
   mutate(family = case_when(family == "thatch" ~ "ground",
                             family == "litter" ~ "ground",
                             family == "rock" ~ "ground",
                                TRUE ~ family)) %>%
   mutate(species = case_when(species == "BG bare ground" ~ "bare_ground",
                                species == "ground" ~ "bare_ground",
                                species == "native_litter" ~ "litter",
                                species == "nonnative_thatch" ~ "thatch",
                                species == "Bromus diandrus BRDI" ~ "Bromus diandrus",
                                species == "Bromus rubens BRRU" ~ "Bromus rubens",
                                species == "Bromus hordeaceus BRHO" ~ "Bromus hordeaceus",
                                species == "Rapistrum rugosum - round pods" ~ "Rapistrum rugosum",
                                species == "HIIN Hirschfeldia incana - long pods" ~ "Hirschfeldia incana",
                                species == "Hirschfeldia incana - long pods" ~ "Hirschfeldia incana",
                                species == "Salvia lucophylla" ~ "Salvia leucophylla",
                                species == "Malacothamnus fasciculatus " ~ "Malacothamnus fasciculatus",
                                species == "dead HEAR" ~ "Heteromeles arbutifolia dead",
                                species == "dead MAFA" ~ "Malacothamnus fasciculatus dead",
                                species == "dead RHIL" ~ "Rhamnus ilicifolia dead",
                                species == "dead ARCA" ~ "Artemisia californica dead",
                                species == "dead SALE" ~ "Salvia leucophylla dead",
                                species == "ACGL Acmispon glaber" ~ "Acmispon glaber",
                                species == "Acmispon maritimus - round, sevfl" ~ "Acmispon maritimus",
                                species == "Acmispon strigosus - oval, sevfl" ~ "Acmispon strigosus",
                                species == "Mirabilis laevis var. crassifolia" ~ "Mirabilis laevis",
                                TRUE ~ species)) %>%
  add_row(species = "Cryptantha spp.",
          lifeform = "forb",
          status = "native",
          lifecycle = "annual",
          family = "Boraginaceae ") %>% 
    add_row(species = "Acmispon glaber dead",
          lifeform = "forb",
          status = "native",
          lifecycle = "annual",
          family = "Fabaceae ") %>% 
    add_row(species = "Uropappus lindleyi",
          lifeform = "forb",
          status = "native",
          lifecycle = "annual",
          family = "Asteraceae   ") %>% 
  distinct()

#View(specieslist)
sort(unique(specieslist$species))
```

-- code to update list of species characteristics
```{r playing with da species list}
ab_species <- sort(unique(abcover_full$species))
species_species <- sort(unique(specieslist$species))

#These are the species that are in abcover_full that are NOT in the species list
difab <- setdiff(ab_species, species_species)
difab 

#These are the species that are in the species list that are NOT in abcover_full
difsp <- setdiff(species_species, ab_species)
difsp

# # add species to the specieslist
# new.specieslist <- specieslist %>%
#   add_row(species = "Cryptantha spp.",
#           lifeform = "forb",
#           status = "native",
#           lifecycle = "annual",
#           family = "Boraginaceae ")
# #           
# # Rename species in the specieslist if needed
# new.specieslist <- specieslist %>%
#     mutate(species = case_when(species == "" ~ "",
#                                TRUE ~ species))
# 
# # Rename species in the data frame (abcover_full) if needed
# new.abcover_full <- abcover_full %>%
#     mutate(species = case_when(species == "Cryptantha sp." ~ "Cryptantha spp.",
#                               species == "dead Acmispon glaber" ~ "Acmispon glaber dead",
#                                TRUE ~ species))
# 
# new.ab_species <- sort(unique(new.abcover_full$species))
# new.species_species <- sort(unique(new.specieslist$species))
# 
# new.ab_species == new.species_species
# 
# #These are the species that are in abcover_full that are NOT in the species list
# new.difab <- setdiff(new.ab_species, new.species_species)
# new.difab
# 
# #These are the species that are in the species list that are NOT in abcover_full
# new.difsp <- setdiff(new.species_species, new.ab_species)
# new.difsp
# # 
# 
# # Run the code below when you are sure you added everything correctly 
# specieslist <- new.specieslist
# abcover_full <- new.abcover_full
# 
# write.csv(specieslist, file = "processed/specieslist.csv") # species and their characteristics :)

```

### left_join species characteristics
```{r left_join species characteristics}
abcover_charateristics <- abcover_full %>% 
  left_join(specieslist, by = "species")

#sort(unique(abcover_charateristics$species))
```

## leaf type
```{r load schlero leaf data, echo=FALSE, eval=F}
#shrub_leaftype <- read_excel("shrub_leaftype.xlsx") #original data

shrub_leaftype <- read.csv("shrub_leaftype.csv") %>% 
  dplyr::select("species", "leaftype") %>% 
   mutate(
      species = as.character(species),
      leaftype = as.character(leaftype)
      ) %>% 
  mutate(
    leaftype = if_else(grepl("dead", species, ignore.case = TRUE), "dead_shrub", leaftype)
  )

print(sort(shrub_leaftype$species))
```

-- code to update list of shrub leaf characteristics
```{r update schlero leaf data}
ab_shrubs <- abcover_charateristics %>% 
   filter(lifeform == "shrub") %>%
  pull(species) %>% 
  unique() %>% 
  sort()

shrub_shrubs <- sort(unique(shrub_leaftype$species))

#These are the shrubs that are in abcover_full that are NOT in the shrub species list
difshr <- setdiff(ab_shrubs, shrub_shrubs)
difshr 

# new.shrub_leaftype <- shrub_leaftype %>%
#   add_row(species = "", leaftype = "")
# 
# # Run the code below when you are sure you added everything correctly
# shrub_leaftype <- new.shrub_leaftype
# 
# write.csv(shrub_leaftype, file = "shrub_leaftype.csv")
```

### left_join leaftype
```{r}
abcover_charateristics_leaftype <- abcover_charateristics %>% 
  left_join(shrub_leaftype, by = "species")
```

# Data prep
```{r data check, include = TRUE}
abcover_charateristics_leaftype %>% 
  group_by(site, transect) %>%
  summarize(data_points = length(unique(distance))) # number of "hits" per transect

abcover_full %>% 
  group_by(site) %>%
  summarize(n_transects = length(unique(transect))) # number of transects per site
# sort(unique(abcover_full$site_rep))
```
- omit ground
```{r omit ground as a status}
abcover_omitground <- abcover_charateristics_leaftype %>% 
  filter(vertical != "ground", 
         vertical != "ground cover") %>% 
  filter(species != "no_species")

#sort(unique(abcover_omitground$status))
```

### subset data by status
```{r hist percent cover by site and status}
cover_bystatus <- abcover_omitground %>% 
  group_by(site, transect, site_rep, year, status) %>% 
  dplyr::summarize(
    count_per_transect = n_distinct(site_rep_dist),
    percentcover = n_distinct(site_rep_dist)/41 # <-- calculate percent cover by transect, lifeform, and status
    ) 

#view(cover_bystatus)
hist(cover_bystatus$percentcover)

# jitter - by site and status
ggplot(data = cover_bystatus, 
       aes(y = percentcover*100, x = site, color = status)) + 
  geom_jitter() +
  facet_wrap(~status)

```

### subset data by lifeform x status
```{r hist percent cover by lifeform and status}

cover_bylfstatus <- abcover_omitground %>% 
  group_by(site, transect, site_rep, year, status, lifeform) %>% 
  dplyr::summarize(
    count_per_transect = n_distinct(site_rep_dist),
    percentcover = n_distinct(site_rep_dist)/41 # <-- calculate percent cover by transect, lifeform, and status
    ) %>% 
  ungroup() %>% 
  unite("lifeform_status", 6:5, sep = "_") #combines characters in two columns to one column, withoutcreating a third column

#view(cover_bystatus_bylf)
hist(cover_bylfstatus$percentcover) 

# jitter - by site, lifeform, and status
cover1 <- ggplot(data = cover_bylfstatus, 
       aes(y = percentcover*100, x = site, color = lifeform_status)) + 
  geom_point() + 
  facet_wrap(~lifeform_status)

# ggplot(data = cover_bylfstatus %>% filter(site == "1" | site == "2"), 
#        aes(y = percentcover*100, x = site, color = lifeform_status)) + 
#   geom_point() + 
#   facet_wrap(~lifeform_status)

# ggplot(data = cover_bylfstatus %>% filter(site == "INT1" | site == "INT2" | site == "INT3" | site == "INT4" | site == "INT5"), 
#        aes(y = percentcover*100, x = site, color = lifeform_status)) + 
#   geom_point() + 
#   facet_wrap(~lifeform_status)

```

### subset data by leaftype
```{r}
cover_byleaftype <- abcover_omitground %>% 
  group_by(site, transect, site_rep, year, status, lifeform, leaftype) %>% 
  dplyr::summarize(
    count_per_transect = n_distinct(site_rep_dist),
    percentcover = n_distinct(site_rep_dist)/41
    ) %>% 
  ungroup() %>% 
  unite("lifeform_status_leaftype", 5:7, sep = "_") %>% 
  dplyr::select(!c(count_per_transect))
```

- Here I'm transposing the dataframes from long to wide  because I want to visualize native shrub cover by non-native grass cover.
```{r transpose}
cover_bylfstatus_wide <- cover_bylfstatus %>%
   dplyr::select(!c(count_per_transect)) %>%
   pivot_wider(names_from = lifeform_status, values_from = percentcover) %>%
   replace_na(replace = list(grass_native = 0, grass_nonnative = 0,
                             forb_native = 0, forb_nonnative = 0, forb_unknown = 0,
                             shrub_native = 0,
                             fern_native = 0,
                             unknown_unknown = 0))  # Replace NA with 0 for each value column

cover_byleaftype_wide <- cover_byleaftype %>% 
  pivot_wider(names_from = lifeform_status_leaftype, values_from = percentcover) %>%
  replace_na(replace = list(native_fern_NA = 0, 
                            unknown_unknown_NA = 0,
                            native_forb_NA = 0, nonnative_forb_NA = 0, unknown_forb_NA = 0,
                            native_shrub_dead_shrub = 0, native_shrub_nonsclero = 0, native_shrub_sclerophyllous = 0,
                            native_grass_NA = 0, nonnative_grass_NA = 0)) 
  
#view(cover_byleaftype_wide)
```
 
# REGRESSIONS
## SHRUBS
This is the spread of 
- non-native grass cover by native shrub cover... 
```{r regression nngrass x shrub}
ggplot(data = cover_bylfstatus_wide, 
       aes(y = grass_nonnative*100, x = shrub_native*100)) + 
   geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "blue", linetype = "solid") # se = confidence interval 

# Fit the linear model
model.s <- lm(grass_nonnative ~ shrub_native, data = cover_bylfstatus_wide)

# Print the summary of the model
model.s_summary <- summary(model.s)

# Extract the R^2 value
r_squared <- round(model.s_summary$r.squared, 3)
print(paste("Multiple R^2:", r_squared))
# adjusted_r_squared <- round(model.s_summary$adj.r.squared, 3)
# print(paste("Adjusted R^2:", adjusted_r_squared))

# Multiple R-squared = variance in the dependent variable that is predictable from the independent variables.  It indicates how well the independent variables explain the variation in the dependent variable. 

# Adjusted R-squared = adjusts for the number of predictors in the model.  It provides a more accurate measure when comparing models with a different number of predictors.
```

- native shrub cover by non-native grass cover ...
```{r regression shrub x nn grass}
cover2 <- ggplot(data = cover_bylfstatus_wide, 
       aes(y = shrub_native*100, x = grass_nonnative*100)) + 
   geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "blue", linetype = "solid")  +
  xlim(0, 100) +
  ylim(0, 100) +
  theme_bw()

model.g <- lm(shrub_native ~ grass_nonnative, data = cover_bylfstatus_wide)
model.g_summary <- summary(model.g)
r_squared <- round(model.g_summary$r.squared, 3)
print(paste("Multiple R^2:", r_squared))
```

- native shrub cover by leaftype by non-native grass cover ...
```{r regression shrub leaftype x nn grass}
sch1 <- ggplot(data = cover_byleaftype_wide) +
  
  geom_point(aes(x = nonnative_grass_NA * 100, y = native_shrub_nonsclero * 100, color = "Native Shrub Nonsclero", shape = site), size = 4, show.legend = TRUE) +
  geom_smooth(aes(x = nonnative_grass_NA * 100, y = native_shrub_nonsclero * 100, color = "Native Shrub Nonsclero"), 
              method = "lm", se = TRUE, linetype = "solid", show.legend = TRUE) +
  
  geom_point(aes(x = nonnative_grass_NA * 100, y = native_shrub_sclerophyllous * 100, color = "Native Shrub Sclero", shape = site), size = 4,  show.legend = TRUE) +
  geom_smooth(aes(x = nonnative_grass_NA * 100, y = native_shrub_sclerophyllous * 100, color = "Native Shrub Sclero"), 
              method = "lm", se = TRUE, linetype = "dashed", show.legend = TRUE) +
  labs(
    x = "Percent cover non-native grass (%)",
    y = "Percent cover native Shrub (%)",
    title = "Comparison of Native Shrub percent cover by Nonnative Grass percent cover",
    color = "Legend"
    ) +
  scale_color_manual(values = c("Native Shrub Nonsclero" = "blue", "Native Shrub Sclero" = "red"),
                     labels = c("Non-sclerophyllous", "Sclerophyllous")) +
  xlim(0, 100) +
  ylim(0, 100) +
  theme_bw()

#only six transects had schlero shrubs present: INT 1-5, Site1 transect 4
#cover_byleaftype_wide %>% filter(native_shrub_sclerophyllous > 0) %>% group_by(site_rep, year) %>%  summarize(native_shrub_sclerophyllous = native_shrub_sclerophyllous)

# model.nonscl <- lm(native_shrub_nonsclero ~ nonnative_grass_NA, data = cover_byleaftype_wide)
# model.nonscl_summary <- summary(model.nonscl)
# r_squared <- round(model.nonscl_summary$r.squared, 3)
# print(paste("Multiple R^2:", r_squared))
# 
# model.scl <- lm(native_shrub_sclerophyllous ~ nonnative_grass_NA, data = cover_byleaftype_wide)
# model.scl_summary <- summary(model.scl)
# r_squared <- round(model.scl_summary$r.squared, 3)
# print(paste("Multiple R^2:", r_squared))
```

## forbs
```{r regression forb x grass}
ggplot(data = cover_bylfstatus_wide) +
  geom_point(aes(x = grass_nonnative * 100, y = forb_native * 100, color = "forb_native"), show.legend = TRUE) +
  geom_smooth(aes(x = grass_nonnative * 100, y = forb_native * 100, color = "forb_native"), 
              method = "lm", se = TRUE, linetype = "solid", show.legend = TRUE) +
  geom_point(aes(x = grass_nonnative * 100, y = forb_nonnative * 100, color = "forb_nonnative"), show.legend = TRUE) +
  geom_smooth(aes(x = grass_nonnative * 100, y = forb_nonnative * 100, color = "forb_nonnative"), 
              method = "lm", se = TRUE, linetype = "dashed", show.legend = TRUE) +
  labs(
    x = "Percent cover non-native grass (%)",
    y = "Percent cover native forb (%)",
    title = "Comparison of Native forb percent cover by Nonnative Grass percent cover",
    color = "Legend"
    ) +
  scale_color_manual(values = c("forb_native" = "blue", "forb_nonnative" = "red"),
                     labels = c("Native ", "Non-native")) +
  xlim(0, 100) +
  ylim(0, 100) +
  theme_bw()
```

## grass
```{r regression grass x grass}
ggplot(data = cover_bylfstatus_wide) +
  geom_point(aes(x = grass_nonnative * 100, y = grass_native * 100, color = "grass_native"), show.legend = TRUE) +
  geom_smooth(aes(x = grass_nonnative * 100, y = grass_native * 100, color = "grass_native"), 
              method = "lm", se = TRUE, linetype = "solid", show.legend = TRUE) +
  labs(
    x = "Percent cover non-native grass (%)",
    y = "Percent cover native grass (%)",
    title = "Comparison of Native grass percent cover by Nonnative Grass percent cover",
    color = "Legend"
    ) +
  scale_color_manual(values = c("grass_native" = "blue"),
                     labels = c("Native ", "Non-native")) +
  xlim(0, 100) +
  ylim(0, 100) +
  theme_bw()

#14 transecrs have native grass present - mostly at Site 1 and at INT transects
#cover_bylfstatus_wide %>% filter(grass_native > 0) %>% group_by(site_rep, year) %>%  summarize(native_grass = grass_native)

```