---
title: "One file to rule all: environmental data for the 39 transects"
author: "Stephanie Ma Lucero"
date: "2024-04-23"
output: html_document
---

```{r libraries, message=FALSE}
library(here) # similar to set working directory
library(tidyverse) # data wrangling - includes ggplot2, dplyr, tidyr, readr, purr, tibble, stringr, forcats
library(rpart) # for classification and regression trees
```

#### load data

```{r load data}
extracted_load <- read.csv("~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022/chaparraldegradation_2022/Extracted_Elevation_Slope_Aspect_Hillshade_20m.csv", header = TRUE, na.strings=c("","NA")) 

# names(extracted_load)

extracted <- extracted_load%>% 
  dplyr::select(!c(system.index, ID, .geo)) %>% 
  rename(transect_ID = system_ind, 
         aspect_degrees = aspect, 
         elevation_m = elevation, 
         slope_degrees = slope
         ) %>% 
 filter(transect_ID != "1.03") %>% 
# “southwest-ness” is calculated using: cos (aspect - 225°)
      mutate(transect_ID = case_when(transect_ID == "1.1" ~ "1.10", # old ~ new
                                   transect_ID == "2.1" ~ "2.10",
                                   transect_ID == "2.2" ~ "2.20",
                                   TRUE ~ transect_ID)) %>% 
  
   mutate(
      aspect_degrees = as.numeric(aspect_degrees),
      elevation_m = as.integer(elevation_m),
      slope_degrees = as.numeric(slope_degrees),
      group_15 = as.character(group_15),
       group_3 = as.character(group_3),
      group_10 = as.character(group_10),
      group_6 = as.character(group_6),
      group_4 = as.character(group_4),
      group_2 = as.character(group_2),
      transect_ID = as.character(transect_ID), 
      fire_history = as.integer(fire_history), 
      twi = as.numeric(twi)
      )
  
# view(extracted)
# length(extracted$transect_ID)    # n = 39
```

# GROUPS by ALL 

_Notes on Classification Trees (from ChatGPT)_

_Purpose: Classification trees are used to classify an outcome into a discrete set of classes (categories). For example, classifying if an email is spam or not spam._

_How It Works:_
_- Splitting: The algorithm splits the dataset into subsets based on the value of input features. The goal is to create branches that result in the most homogenous subsets._
_- Decision Nodes and Leaves: Each internal node represents a "test" on an attribute (e.g., whether a feature is above or below a certain value), each branch represents the outcome of the test, and each leaf node represents a class label (decision taken after computing all attributes)._
_- Gini Impurity or Entropy: To decide the best splits, the algorithm uses metrics like Gini impurity or entropy to measure the homogeneity of the nodes._

_Example:_
_A classification tree for predicting whether a patient has a disease based on symptoms might start with a root node testing if the patient has a fever, and then branch into other symptoms like cough and sore throat._

Environmental variables include:
- aspect (degrees)
- slope (degrees)
- elevation (m)
- hillshade (0-255) - considers sun azimuth and altitude 
--- definition: hillshading computes __surface illumination as values from 0 to 255__ based on a given compass direction to the sun (azimuth) and a certain altitude above the horizon (altitude). Hillshades are often used to produce maps that are visually appealing.
- southwestness (-1 to 1)


```{r classification tree - group_15}

model_15 <- rpart(group_15 ~ 
                  #aspect_degrees + elevation_m + hillshade + slope_degrees + southwestness,  # tree can't handle interacting terms
                   elevation_m + slope_degrees + southwestness, 
                  data = extracted)
summary(model_15)
model_15 # data output

# Plotting the tree
par(mfrow = c(1,2), xpd = NA)
plot(model_15)
text(model_15, use.n = TRUE)
```

```{r classification tree - group_9}
model_9 <- rpart(group_9 ~ 
                   #aspect_degrees + elevation_m + hillshade + slope_degrees + southwestness,  # tree can't handle interacting terms
                     elevation_m + slope_degrees + southwestness, 
                  data = extracted)

summary(model_9) # data output

# Plotting the tree
par(mfrow = c(1,2), xpd = NA)
plot(model_9)
text(model_9, use.n = TRUE)
```

```{r classification tree - group_6}
model_6 <- rpart(group_6 ~ 
                   #aspect_degrees + elevation_m + hillshade + slope_degrees + southwestness,  # tree can't handle interacting terms
                   elevation_m + slope_degrees + southwestness, # southwestness and aspect_degrees result in the same tree
                  data = extracted)

model_6 # data output

# Plotting the tree
par(mfrow = c(1,2), xpd = NA)
plot(model_6)
text(model_6, use.n = TRUE)
```

```{r classification tree - group_4}
model_4 <- rpart(group_4 ~ 
                   #aspect_degrees + elevation_m + hillshade + slope_degrees + southwestness,  # tree can't handle interacting terms
                   elevation_m + slope_degrees + aspect_degrees, # doesn't run with southwestness
                 data = extracted)

model_4 # data output
summary(model_4)

# Plotting the tree
par(mfrow = c(1,2), xpd = NA)
plot(model_4)
text(model_4, use.n = TRUE)
```
```{r classification tree - group_2}
model_2 <- rpart(group_2 ~ aspect_degrees + elevation_m + hillshade + slope_degrees + southwestness,  # tree can't handle interacting terms
                  data = extracted)

model_2 # data output

# model_2 is not a tree, just a root - because there are already two groups so nothing predicts the split.

```

# GROUPS by ASPECT 

```{r classification tree - aspect}
model <- rpart(community_number ~ aspect_degrees, data = extracted)
# Plotting the tree
par(mfrow = c(1,2), xpd = NA)
plot(model)
text(model, use.n = TRUE)
```

```{r classification tree - elevation}
model <- rpart(community_number ~ elevation_m, data = extracted)
# Plotting the tree
par(mfrow = c(1,2), xpd = NA)
plot(model)
text(model, use.n = TRUE)
```

```{r classification tree - slope}
model <- rpart(community_number ~ slope_degrees, data = extracted)
# Plotting the tree
par(mfrow = c(1,2), xpd = NA)
plot(model)
text(model, use.n = TRUE)
```
