---
title: "One file to rule all: let's visualize the environmental data by community groups!"
author: "Stephanie Ma Lucero"
date: "2024-08-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libraries, message=FALSE}
library(here) # similar to set working directory
library(tidyverse) # data wrangling - includes ggplot2, dplyr, tidyr, readr, purr, tibble, stringr, forcats
library(ggplot2)
library(plotly)
```

```{r colors}
devtools::install_github("an-bui/calecopal")

library(calecopal)
# all palettes
# names(cal_palettes)

# Group_15 colors
cal_palette(name = "grassdry", n = 15, type = "continuous") 
c15 <- "#E1BC8D" 
c14 <- "#C6A076" 
c13 <- "#AB845F" 
c12 <- "#916849" 
c11 <- "#7E593A"
c10 <- "#725532" 
c9 <- "#66512A" 
c8 <- "#5B4E23" 
c7 <- "#504521" 
c6 <- "#453C1F"
c5 <- "#3A341D" 
c4 <- "#383524" 
c3 <- "#3E3F33" 
c2 <- "#454943" 
c1 <- "#4C5454"
colors_15 <- c(c1, c2, c3, c4, c5, c6, c7, c8, c9, c10, c11, c12, c13, c14, c15)

# group_3 colors
cal_palette(name = "grassdry", n = 5, type = "continuous") 
grass <- "#E1BC8D" 
css <- "#66512A" 
chap <- "#4C5454"

# group_4 colors
cal_palette(name = "grassdry", n = 5, type = "continuous") 
d <- "#E1BC8D" 
c <- "#765635" 
b <- "#5B4E23"
a <- "#4C5454"
colors_4 <- c(a, b, c, d)


cal_palette(name = "creek", n = 15, type = "continuous") 
"#EBDAC9" "#E0C9B6" "#D6B9A4" "#CEAF9A" "#CEBBB0"
"#CEC7C6" "#BFC1C3" "#9B9F9A" "#767C70" "#606B5A"
"#546450" "#475E46" "#3B5139" "#2F422B" "#23341E"

# # Status colors: native, exotic 
# cal_palette(name = "chaparral1", n = 6, type = "continuous") 
# native <- "#AEBFA8"
# exotic <- "#F19B34"
# statuscolors <- c(native, exotic)


```

#### load data

```{r load data}
extracted_load <- read.csv("~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022/chaparraldegradation_2022/Extracted_Elevation_Slope_Aspect_Hillshade_20m.csv", header = TRUE, na.strings=c("","NA")) 

extracted <- extracted_load%>% 
  dplyr::select(!c(system.index, ID, .geo, aspect)) %>% # omit aspect because INT transects cross from 21 to 350 degrees
  rename(transect_ID = system_ind, 
         elevation_m = elevation, 
         slope_degrees = slope
         ) %>% 
# “southwest-ness” is calculated using: cos (aspect - 225°)
      mutate(transect_ID = case_when(transect_ID == "1.1" ~ "1.10", # old ~ new
                                   transect_ID == "2.1" ~ "2.10",
                                   transect_ID == "2.2" ~ "2.20",
                                   TRUE ~ transect_ID)) %>% 
  
   mutate(
      southwestness = as.numeric(southwestness),
      elevation_m = as.integer(elevation_m),
      slope_degrees = as.numeric(slope_degrees),
      group_15 = as.character(group_15),
      group_10 = as.character(group_10),
      group_6 = as.character(group_6),
      group_4 = as.character(group_4),
      group_2 = as.character(group_2),
      transect_ID = as.character(transect_ID), 
      fire_history = as.integer(fire_history), 
      twi = as.numeric(twi), 
      lat = as.numeric(lat), 
      long = as.numeric(long)
   )   %>% 
   mutate(group_typeconvert = factor(
     group_typeconvert, levels = c("chaparral", "css", "grass")
    )
  ) %>% 
  mutate(group_2 = factor(
     group_2, levels = c("HEAR", "no HEAR")
    )
  ) %>% 
   mutate(group_4 = factor(
     group_4, levels = c("HEAR+CEOL", "HEAR", "SALE+MAFA+ARCA", "PHHU")
    )
  ) %>% 
   mutate(group_6 = factor(
     group_6, levels = c("HEAR+CEOL", "HEAR", 
                         "SALE+MAFA+ARCA+STLE+CEME", 
                         "SALE+MAFA+ARCA+HIIN",
                         "PHHU", "PHHU+RARU")
    )
  ) %>% 
     mutate(group_10 = factor(
     group_10, levels = c("HEAR+CEOL", "HEAR", 
                         "SALE+MAFA+ARCA+STLE+CEME+BRDI",
                         "SALE+MAFA+ARCA+STLE+CEME+Erodium_spp.",
                         "SALE+MAFA+ARCA+HIIN+BRHO+CEME",
                         "SALE+MAFA+ARCA+HIIN+Erodium_spp.",
                         "PHHU",
                         "PHHU+BRDI",
                         "PHHU+ARCA",
                         "PHHU+RARU")
    )
  ) %>% 
     mutate(group_15 = factor(
     group_15, levels = c("1", "2", 
                         "3",
                         "4",
                         "5",
                         "6",
                         "7",
                         "8",
                         "9",
                         "10", "11", "12", "13", "14", "15")
    )
  ) 
  
# view(extracted)
```

Environmental variables include:
- southwestness (-1 to 1)
- slope (degrees)
- elevation (m)
- hillshade (0-255) - considers sun azimuth and altitude 
--- definition: hillshading computes __surface illumination as values from 0 to 255__ based on a given compass direction to the sun (azimuth) and a certain altitude above the horizon (altitude). Hillshades are often used to produce maps that are visually appealing.


#SOUTHWESTNESS

```{r southwestness normality}
hist(extracted$southwestness) # southwestness
# -1  
# 1
ggqqplot(extracted$southwestness)
shapiro.test(extracted$southwestness)
# Shapiro-Wilk normality test
# data:  extracted$southwestness
# W = 0.88105, p-value = 0.0006598 <- not normally distributed
```

```{r southwestness mean}
extracted %>%
  group_by(group_4) %>% 
  summarise(
    southwestness_N = length(southwestness),
    southwestness_mean = mean(southwestness),
    southwestness_SE = round(((sd(southwestness))/sqrt(length((southwestness)))),2)
            )
```

```{r southwestness visuals}
# boxplot(southwestness ~ group_4,
#   data = extracted)

ggplot(data = extracted, 
       aes(y = southwestness, x = group_4)) + 
  # geom_point() +
  geom_boxplot() +
  theme_classic() +
  labs(x = "Community Groups (n = 4)", 
       y = "Southwestness (-1 = northeast, 1 = southwest)", 
       title = "Community Groups by Southwestness")
 
ggplot(data = extracted, 
       aes(y = southwestness, x = group_9)) + 
  geom_boxplot() +
  theme_classic() +
  labs(x = "Community Groups (n = 9)", 
       y = "Southwestness (-1 = northeast, 1 = southwest)", 
       title = "Community Groups by Southwestness") +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  ylim(-1, 1)

```


# SLOPE

```{r slope normality}
hist(extracted$slope_degrees) # slope

ggqqplot(extracted$slope_degrees)
shapiro.test(extracted$slope_degrees)
# data:  extracted$slope_degrees
# W = 0.96687, p-value = 0.2994 <-- normally distributed

```

```{r slope mean}
extracted %>%
  group_by(group_4) %>% 
  summarise(
    slope_N = length(slope_degrees),
    slope_mean = mean(slope_degrees),
    slope_SE = round(((sd(slope_degrees))/sqrt(length((slope_degrees)))),2)
            )

summary(aov(slope_degrees ~ group_4, data = extracted))
TukeyHSD((aov(slope_degrees ~ group_4, data = extracted)))
plot(TukeyHSD((aov(slope_degrees ~ group_4, data = extracted))))
```
```{r slope visuals}
boxplot(slope_degrees ~ group_4,
  data = extracted)

ggplot(data = extracted, 
       aes(y = slope_degrees, x = group_4)) + 
  # geom_point() +
  geom_boxplot() +
  theme_classic()

```

#ELEVATION
```{r}
hist(extracted$elevation) # elevation

ggplot(data = extracted, 
       aes(y = elevation_m, x = group_9)) + 
  # geom_point() +
  geom_boxplot() +
  theme_classic()


```

# lat x long x elevation
```{r 3D figure}
p <- plot_ly(data = extracted, 
       x = ~elevation_m, y = ~long, z = ~lat, 
       type = "scatter3d", 
       mode = 'markers', marker = list(size=5),
       color = ~group_typeconvert
       ) %>% 
   layout(scene = list(
     xaxis = list(title = 'Elevation (m)'),
     yaxis = list(title = 'Longitude (E-W)'),
     zaxis = list(title = 'Latitude (N-S)')
     )
     )



p3 <- plot_ly(data = extracted, 
       x = ~lat*-1, 
       y = ~long, 
       z = ~elevation_m, 
       type = "scatter3d", 
       mode = 'markers', marker = list(size=5),
       color = ~group_typeconvert, colors = (c("green", "purple", "orange"))
       ) %>% 
   layout(scene = list(
      xaxis = list(title = 'Latitude (-1) (S-N)'),
     yaxis = list(title = 'Longitude (W-E)'),
      zaxis = list(title = 'Elevation (m)')  
     )
     )

p15 <- plot_ly(data = extracted, 
       x = ~lat*-1, 
       y = ~long, 
       z = ~elevation_m, 
       type = "scatter3d", 
       mode = 'markers', marker = list(size=10),
       color = ~group_15, colors = (colors_15)
       ) %>% 
   layout(scene = list(
      xaxis = list(title = 'Latitude (-1) (S-N)'),
     yaxis = list(title = 'Longitude (W-E)'),
      zaxis = list(title = 'Elevation (m)')  
     )
     )

p4 <- plot_ly(data = extracted, 
       x = ~lat*-1, 
       y = ~long, 
       z = ~elevation_m, 
       type = "scatter3d", 
       mode = 'markers', marker = list(size=10),
       color = ~group_4, colors = (colors_4)
       ) %>% 
   layout(scene = list(
      xaxis = list(title = 'Latitude (-1) (S-N)'),
     yaxis = list(title = 'Longitude (W-E)'),
      zaxis = list(title = 'Elevation (m)')  
     )
     )

p_site1 <- plot_ly(data = (extracted %>% filter(site == "site_1")), 
       x = ~lat*-1, 
       y = ~long, 
       z = ~elevation_m, 
       type = "scatter3d", 
       mode = 'markers', marker = list(size=10),
       color = ~group_4, #group
       colors = colors_4 #color
       ) %>% 
   layout(scene = list(
      xaxis = list(title = 'Latitude (-1) (S-N)'),
     yaxis = list(title = 'Longitude (W-E)'),
      zaxis = list(title = 'Elevation (m)')  
     )
     )


p_site2 <- plot_ly(data = (extracted %>% filter(site == "site_2")), 
       x = ~lat*-1, 
       y = ~long, 
       z = ~elevation_m, 
       type = "scatter3d", 
       mode = 'markers', marker = list(size=10),
       color = ~group_4, # grouping 
       colors = colors_4 # color
       ) %>% 
   layout(scene = list(
      xaxis = list(title = 'Latitude (-1) (S-N)'),
     yaxis = list(title = 'Longitude (W-E)'),
      zaxis = list(title = 'Elevation (m)')  
     )
     )

```


```{r}
hist(extracted$hillshade) # hillshade

ggplot(data = extracted, 
       aes(y = hillshade, x = group_9)) + 
   # geom_point() +
  geom_boxplot() +
  theme_classic()
```