---
title: "One file to rule all: environmental data for the 39 transects"
author: "Stephanie Ma Lucero"
date: "2024-04-23"
output: html_document
---

```{r libraries, message=FALSE}
library(here) # similar to set working directory
library(tidyverse) # data wrangling - includes ggplot2, dplyr, tidyr, readr, purr, tibble, stringr, forcats
library(rpart) # for classification and regression trees
```

```{r load data}
extracted_load <- read.csv("~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022/chaparraldegradation_2022/Extracted_Elevation_Slope_Aspect_Hillshade_20m.csv", header = TRUE, na.strings=c("","NA")) 

extracted <- extracted_load%>% 
  dplyr::select(!c(system.index, ID, .geo)) %>% 
  rename(transect_ID = system_ind, 
         aspect_degrees = aspect, 
         elevation_m = elevation, 
         slope_degrees = slope
         ) %>% 
# “southwest-ness” is calculated using: cos (aspect - 225°)
      mutate(transect_ID = case_when(transect_ID == "1.1" ~ "1.10", # old ~ new
                                   transect_ID == "2.1" ~ "2.10",
                                   transect_ID == "2.2" ~ "2.20",
                                   TRUE ~ transect_ID)) %>% 
  
   mutate(
      aspect_degrees = as.numeric(aspect_degrees),
      elevation_m = as.integer(elevation_m),
      slope_degrees = as.numeric(slope_degrees),
      group_15 = as.character(group_15),
      group_9 = as.character(group_9),
      group_6 = as.character(group_6),
      group_4 = as.character(group_4),
      group_2 = as.character(group_2),
      transect_ID = as.character(transect_ID)
      )
  
```

_Notes on Classification Trees (from ChatGPT)_

Purpose: Classification trees are used to classify an outcome into a discrete set of classes (categories). For example, classifying if an email is spam or not spam.

How It Works:

Splitting: The algorithm splits the dataset into subsets based on the value of input features. The goal is to create branches that result in the most homogenous subsets.
Decision Nodes and Leaves: Each internal node represents a "test" on an attribute (e.g., whether a feature is above or below a certain value), each branch represents the outcome of the test, and each leaf node represents a class label (decision taken after computing all attributes).
Gini Impurity or Entropy: To decide the best splits, the algorithm uses metrics like Gini impurity or entropy to measure the homogeneity of the nodes.
Example:

A classification tree for predicting whether a patient has a disease based on symptoms might start with a root node testing if the patient has a fever, and then branch into other symptoms like cough and sore throat.

```{r classification tree - group_15}
model_15 <- rpart(group_15 ~ aspect_degrees + elevation_m + hillshade + slope_degrees + southwestness,  # tree can't handle interacting terms
                  data = extracted)
summary(model_15)
model_15 # data output

# Plotting the tree
par(mfrow = c(1,2), xpd = NA)
plot(model_15)
text(model_15, use.n = TRUE)
```

```{r classification tree - group_9}
model_9 <- rpart(group_9 ~ aspect_degrees + elevation_m + hillshade + slope_degrees + southwestness,  # tree can't handle interacting terms
                  data = extracted)

summary(model_9) # data output

# Plotting the tree
par(mfrow = c(1,2), xpd = NA)
plot(model_9)
text(model_9, use.n = TRUE)
```

```{r classification tree - group_6}
model_6 <- rpart(group_6 ~ aspect_degrees + elevation_m + hillshade + slope_degrees + southwestness,  # tree can't handle interacting terms
                  data = extracted)

model_6 # data output

# Plotting the tree
par(mfrow = c(1,2), xpd = NA)
plot(model_6)
text(model_6, use.n = TRUE)
```

```{r classification tree - group_4}
model_4 <- rpart(group_4 ~ aspect_degrees + elevation_m + hillshade + slope_degrees + southwestness,  # tree can't handle interacting terms
                  data = extracted)

model_4 # data output

# Plotting the tree
par(mfrow = c(1,2), xpd = NA)
plot(model_4)
text(model_4, use.n = TRUE)
```
```{r classification tree - group_2}
model_2 <- rpart(group_2 ~ aspect_degrees + elevation_m + hillshade + slope_degrees + southwestness,  # tree can't handle interacting terms
                  data = extracted)

model_2 # data output

# model_2 is not a tree, just a root - because there are already two groups so nothing predicts the split.

```

```{r classification tree - aspect}
model <- rpart(community_number ~ aspect_degrees, data = extracted)
# Plotting the tree
par(mfrow = c(1,2), xpd = NA)
plot(model)
text(model, use.n = TRUE)
```

```{r classification tree - elevation}
model <- rpart(community_number ~ elevation_m, data = extracted)
# Plotting the tree
par(mfrow = c(1,2), xpd = NA)
plot(model)
text(model, use.n = TRUE)
```

```{r classification tree - slope}
model <- rpart(community_number ~ slope_degrees, data = extracted)
# Plotting the tree
par(mfrow = c(1,2), xpd = NA)
plot(model)
text(model, use.n = TRUE)
```
