---
title: "Ch1_Degradation_seedbankNMDS"
author: "Stephanie Ma Lucero"
date: "2023-03-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install libraries, message=FALSE}

library(tidyverse) # includes ggplot2, dplyr, tidyr, readr, purr, tibble, stringr, forcats
library(ggplot2) # data visualization
library(readr)
library(dplyr)
library(ggpubr) # customize ggplot2 for publication
library(broom)
library(AICcmodavg)
library(vegan)
library(readxl) # for .xls and .xlsx sheets
library(janitor)
library(calecopal) 
library(multcompView)
library(plotly)

# dune data: https://www.rpubs.com/RGrieger/545184

```

```{r load deg_seedbank csv}
# 
seedbank_load_raw <- read.csv("~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022b/deg_seedbank_20230308.csv", header = TRUE, na.strings=c("","NA"))
head(seedbank_load_raw)
str(seedbank_load_raw)

seedbank_load <- seedbank_load_raw %>% 
  select(c("stand", "site", "soilcore", "depthname", "treatment", "species", "lifeform", "status", "lifecycle", "family", "totalcount"))
  
seedbank_load$totalcount <- as.numeric(seedbank_load$totalcount)
str(seedbank_load)

# load without unknowns and NAs <-- how do I omit UNKNOWNs?
seedbank_load %>%
  filter(!str_detect(species, "unk")) 

# rename treatments to be all lowercase 
seedbank_load <- seedbank_load %>% 
  mutate(treatment = recode(treatment,
                            Oven = 'oven',
                            Ovenchar = 'ovenchar',
                            Char = 'char',
                            Control = 'control', 
                            Native = 'native')
                            ) 
  
         
print(seedbank_load)
unique(seedbank_load$stand)
unique(seedbank_load$site)
unique(seedbank_load$soilcore)
unique(seedbank_load$depthname)
unique(seedbank_load$treatment)
unique(seedbank_load$species)
```

(1) Organize the seed bank data by converting the df into a tibble.
```{r preparing df, include = TRUE}

dim(seedbank_load)
# [1] 2117   19
head(seedbank_load)
tail(seedbank_load)

# convert dataframe to tibble
seedbank_load <- as_tibble(seedbank_load) 

class(seedbank_load) # now a tibble
```

(2) Organize the seedbank data by reordering the classes: Intact, Matrix, Degraded.
```{r reorder stand classes}
seedbank_load_v <- seedbank_load %>%
  mutate(stand = fct_relevel(stand, 
            "intact", "enhanced", "degraded")) %>% 
  mutate(treatment = fct_relevel(treatment, 
                                 "control", "charate", "oven", "ovenchar"))
```

(3) Create new columns for SITE_CORE, SITE_CORE_DEPTH, SITE_CORE_DEPTH_TREATMENT, SITE_TREATMENT, SITE_DEPTH for seed bank data. 
```{r combine columns}

# combine SITE and CORE to one column
seedbank_load_v$stand_core <- paste(seedbank_load_v$site, seedbank_load_v$soilcore, sep="_") 

# combine SITE_CORE and DEPT to one column
seedbank_load_v$stand_core_depth <- paste(seedbank_load_v$stand_core, seedbank_load_v$depthname, sep="_") 

# combine SITE_CORE_DEPTH and TREATMENT to one column
seedbank_load_v$stand_core_depth_treat <- paste(seedbank_load_v$stand_core_depth, seedbank_load_v$treatment, sep="_") 
```

(4) Omit "pre-treatment" as a treatment since I'm only interested in Control, Char, Oven, and CharOven treatments
```{r omit pretreatment data}

seedbank_load_v_omitpretreatment <- seedbank_load_v %>% 
  filter(!str_detect(treatment, "pretreat")) 
head(seedbank_load_v_omitpretreatment)

unique(seedbank_load_v_omitpretreatment$treatment) # list unique treatments --> ordered: control, charate, oven, ovenchar
unique(seedbank_load_v_omitpretreatment$status) # list unique treatments --> native, nonnative
unique(seedbank_load_v_omitpretreatment$species) # list unique species --> 42 species
unique(seedbank_load_v_omitpretreatment$stand_core) # unique stand_core --> 26 = 7 + 10 + 9

```

(5.1) sum of germinates at each stand_core (e.g., present in one core = sum (char, control, oven, charoven)
```{r organize data for nmds, inclue = TRUE}

# species in columns
# sites or observations in rows

# this is the df you'll call to set GROUPS: seedbank_bysitecore
seedbank_bysitecore <- seedbank_load_v_omitpretreatment %>% 
  group_by(stand, site, stand_core, species) %>% # by stand_core and species 
  # This will give a data point for every species based on germination numbers per species
  #group_by(stand_core_depth, species) %>% # by stand_core_depth - DOES NOT CONSIDER TREATMENT
  #group_by(stand_core_detph_treat, species) %>% # by stand_core_depth_treat
  filter(!is.na(species)) %>% # omit blank cells in species 
  dplyr::summarize(totalgerm = sum(totalcount)) 

view(seedbank_bysitecore) # df before transpose - use for ellipses
```

(6) Prepare data for an NMDS
```{r transpose data by species for nmds, include=TRUE}

# species in columns down
# observations or sites in rows across

seedbank_bysitecore_nmdsprep <- seedbank_bysitecore %>% 
 tidyr::spread(species, totalgerm) %>% # transpose data so columns are stand_core_depth
  #tidyr::spread(stand_core_depth, totalgerm) %>% # transpose data so columns are stand_core_depth
  # tidyr::spread(stand_core_depth_treat, totalgerm) %>% # transpose data so columns are stand_core_depth_treat
  replace(is.na(.), 0) # replace NA with zeros --> isn't compatible with character columns

view(seedbank_bysitecore_nmdsprep)
# 26 x 45
```

```{r replace NAs with zeros}

# You don't need to run this right now

seedbank_bysitecore_nmdsprep_num <- seedbank_bysitecore_nmdsprep %>% 
  # use the ungroup() function to take out grouping
  ungroup() %>% 
  select(!c(stand_core, stand, site)) %>% # de-select these columns
  replace(is.na(.), 0)

view(seedbank_bysitecore_nmdsprep_num) # only numbers
# 26 x 42 
```

```{r write nmds matrix to csv}

write.csv(seedbank_bysitecore_nmdsprep_num, "~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022b/processed/seedbank_NMDS_202303.csv")

```

(7) NMDS 
```{r stress tests}
# USE THIS ONE!

# str(seedbank_bysitecore_nmdsprep_num)
# stand_core_depth -->  tibble [89 × 52]

# stress test, will run 20 times
seedbank_bysitecore_nmds <- metaMDS(seedbank_bysitecore_nmdsprep_num, distance = "raup")
# stress tests are really low ==> I may have insufficient data. :/

# stress test, will run 20 times
seedbank_bysitecore_nmds <- metaMDS(seedbank_bysitecore_nmdsprep_num, autotransform = F)
```

```{r scores nmds}

scores(seedbank_bysitecore_nmds)
```

```{r stress test}

stressplot(seedbank_bysitecore_nmds)
# if stress tests are really low ==> I may have insufficient data. :/

#plot_df <- scores(seedbank_bysitecore_nmds, display = "sites") %>% # from An's code
#  as.data.frame() %>% 
#  rownames_to_column("stand_core") %>% 
#  full_join(seedbank_bysitecore, by = "stand_core")

nmds_df <- data.frame(seedbank_bysitecore_nmds$points, seedbank_bysitecore_nmdsprep) # from chatgpt


```

```{r ggplot color}
ggplot(nmds_df, aes(x = MDS1, y = MDS2, color = stand, shape = stand)) +
  geom_point(size = 3, alpha = 0.8) +
  scale_color_manual(values = c("gray0", "gray30", "gray60")) +
  theme_bw() +
  stat_ellipse(linetype = 2, size = 1) +
  labs(title = "NMDS: by seed bank germination counts", color = "Stand type", shape = "Stand type")
```

```{r ggplot bw}
ggplot(nmds_df, aes(x = MDS1, y = MDS2, color = stand, shape = stand)) +
  geom_point(size = 3, alpha = 0.8) +
  scale_color_manual(values = c("darkgreen", "darkmagenta", "darkgoldenrod1")) +
  theme_bw() +
  #stat_ellipse(level = .95, linetype = 2, size = 1) +
  stat_ellipse(linetype = 2, size = 1) +
  labs(title = "NMDS: by seed bank germination counts", color = "Stand type", shape = "Stand type")


```

```{r plot nmds scores}
# code from An Bui: https://rpubs.com/an-bui/vegan-cheat-sheet 



```


```{r plot nmds basics}
#basic normal plot
ordiplot(seedbank_bysitecore_nmds) # without point labels
ordiplot(seedbank_bysitecore_nmds, type = "t") # with point labels

#ggvegan to make a quick nice plot
autoplot(seedbank_bysitecore_nmds) # NMDS1 x NMDS2, Key = "Score": sites, species
# sites == total germ per species
# species == stand_core


#this makes a dataframe that gives a lot of control graphically 
seedbank_bysitecore_nmds_fort <- fortify(seedbank_bysitecore_nmds)

# vectors AND measurments
seedbank_bysitecore_nmds_fortplot <- ggplot() + 
  geom_point(data = subset(seedbank_bysitecore_nmds_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0.5) +
  geom_segment(data = subset(seedbank_bysitecore_nmds_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               linewidth = 0.8) +
  geom_text(data = subset(seedbank_bysitecore_nmds_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1))+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", linewidth = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", linewidth = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

#only vectors shown
seedbank_bysitecore_nmds_fortplot1 <- ggplot() + 
  geom_point(data = subset(seedbank_bysitecore_nmds_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0) +
  geom_segment(data = subset(seedbank_bysitecore_nmds_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               linewidth = 0.8) +
  geom_text(data = subset(seedbank_bysitecore_nmds_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1))+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", linewidth = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
seedbank_bysitecore_nmds_fortplot1

#only measurements shown
seedbank_bysitecore_nmds_fortplot2 <- ggplot() + 
  geom_point(data = subset(seedbank_bysitecore_nmds_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0.5) +
  geom_segment(data = subset(seedbank_bysitecore_nmds_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               size = 0.8,
               alpha = 0) +
  geom_text(data = subset(seedbank_bysitecore_nmds_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1),
            alpha = 0)+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", size = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

#interactive plots that I like
ggplotly(seedbank_bysitecore_nmds_fortplot)

#show both plots 
hi <- ggarrange(seedbank_bysitecore_nmds_fortplot1, seedbank_bysitecore_nmds_fortplot2, ncol = 1)
hi

```

```{r NMDS ellipses}

# from https://www.rpubs.com/RGrieger/545184 - dune data - adding ellipses
data(dune)
data("dune.env")
view(dune)
view(dune.env)

plot(seedbank_bysitecore_nmds, type = "n")

#adding ellipses
ordiplot(seedbank_bysitecore_nmds, type = "n", main = "ellipses")
orditorp(seedbank_bysitecore_nmds, display = "sites", labels = T, pch = c(16, 8, 12) [as.numeric(seedbank_bysitecore$stand)], col = c("green", "blue", "orange") [as.numeric(seedbank_bysitecore$stand)], cex = 1)
ordiellipse(seedbank_bysitecore_nmds, groups = seedbank_bysitecore$stand, draw = "polygon", lty = 1, col = "grey90")

ordiplot(seedbank_bysitecore_nmds,type="n")
orditorp(seedbank_bysitecore_nmds,display="species",col="red",air=0.01)
orditorp(seedbank_bysitecore_nmds,display="sites",cex=1.25,air=0.01)

str(seedbank_bysitecore_nmdsprep_num)
#tibble [90 × 27] (S3: tbl_df/tbl/data.frame)
treat=c(rep("treatment1",7),rep("treatment2",7), rep("treatment3",7)) 
ordiplot(abcover_byspecies_NMDS,type="n")
ordihull(abcover_byspecies_NMDS,groups=treat,draw="polygon",col="grey90",
         label=FALSE)
orditorp(abcover_byspecies_NMDS,display="species",col="red",air=0.01)
orditorp(abcover_byspecies_NMDS,display="sites",col=c(rep("green",50),rep("blue",50),re("orange",25)),
         air=0.01,cex=1.25)


```



(8) sum of germinates by species (e.g., present in one core = sum (char, control, oven, charoven)
```{r seedbank sum germinants by spp, inclue = TRUE}


seedbank_byspp <- seedbank_load_v_omitpretreatment %>% 
  group_by(species, stand_core) %>% # by stand_core
  #group_by(stand_core_depth, species) %>% # by stand_core_depth - DOES NOT CONSIDER TREATMENT
  #group_by(stand_core_detph_treat, species) %>% # by stand_core_depth_treat
  filter(!is.na(species)) %>% # omit blank cells in species 
  dplyr::summarize(totalgerm = sum(totalcount)) 

head(seedbank_byspp) 
```

(9) Prepare data for an NMDS
```{r transpose seedbank data by spp, include=TRUE}

seedbank_byspp_nmdsprep <- seedbank_byspp %>% 
  tidyr::spread(species, totalgerm) %>% # transpose data so columns are stand_core
  replace(is.na(.), 0) # replace NA with zeros
head(seedbank_byspp_nmdsprep)

seedbank_byspp_nmdsprep_num <- seedbank_byspp_nmdsprep %>% 
  # use the ungroup() function to take out grouping
  ungroup() %>% 
  select(!c(stand_core))
         
view(seedbank_byspp_nmdsprep_num) # only numbers
write.csv(seedbank_byspp_nmdsprep_num, "~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022b/processed/seedbank_NMDS_byspp_202303.csv")
```

(10) NMDS 
```{r FROM KIT - nmds by species - use this code from KIT}
# USE THIS ONE!

str(seedbank_byspp_nmdsprep_num)
#tibble [27 × 90] (S3: tbl_df/tbl/data.frame)

# stress test, will run 20 times
seedbank_byspp_nmds <- metaMDS(seedbank_byspp_nmdsprep_num, distance = "raup")
# some data have empty spp

# stress test, will run 20 times
seedbank_byspp_nmds <- metaMDS(seedbank_byspp_nmdsprep_num, autotransform = F)
# stress tests are really low ==> I may have insufficient data. :/

stressplot(seedbank_byspp_nmds)
plot(seedbank_byspp_nmds)

#basic normal plot
ordiplot(seedbank_byspp_nmds) # without point labels
ordiplot(seedbank_byspp_nmds, type = "t") # with point labels

#ggvegan to make a quick nice plot
autoplot(seedbank_byspp_nmds) # NMDS1 x NMDS2, Key = "Score": sites, species
# sites == total germ per species
# species == stand_core


#this makes a dataframe that gives a lot of control graphically 
seedbank_byspp_nmds_fort <- fortify(seedbank_byspp_nmds)

seedbank_byspp_nmds_fortplot <- ggplot() + 
  geom_point(data = subset(seedbank_byspp_nmds_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0.5) +
  geom_segment(data = subset(seedbank_byspp_nmds_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               linewidth = 0.8) +
  geom_text(data = subset(seedbank_byspp_nmds_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1))+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", linewidth = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", linewidth = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
seedbank_byspp_nmds_fort

#only vectors shown
seedbank_byspp_nmds_fortplot1 <- ggplot() + 
  geom_point(data = subset(seedbank_byspp_nmds_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0) +
  geom_segment(data = subset(seedbank_byspp_nmds_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               linewidth = 0.8) +
  geom_text(data = subset(seedbank_byspp_nmds_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1))+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", linewidth = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
seedbank_byspp_nmds_fortplot1

#only measurements shown
seedbank_byspp_nmds_fortplot2 <- ggplot() + 
  geom_point(data = subset(seedbank_byspp_nmds_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0.5) +
  geom_segment(data = subset(seedbank_byspp_nmds_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               size = 0.8,
               alpha = 0) +
  geom_text(data = subset(seedbank_byspp_nmds_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1),
            alpha = 0)+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", size = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
seedbank_byspp_nmds_fortplot2

#interactive plots that I like
ggplotly(seedbank_byspp_nmds_fortplot)

#show both plots 
hi3 <- ggarrange(seedbank_byspp_nmds_fortplot1, seedbank_byspp_nmds_fortplot2, ncol = 1)
hi3


ordiplot(seedbank_byspp_nmds,type="n")
orditorp(seedbank_byspp_nmds,display="species",col="red",air=0.01)
orditorp(seedbank_byspp_nmds,display="sites",cex=1.25,air=0.01)

str(seedbank_byspp_nmdsprep_num)
#tibble [27 × 90] (S3: tbl_df/tbl/data.frame)
# I don't think I can make this becuase my soil cores are diff number for each stand.
treat=c(rep("treatment1",7),rep("treatment2",7), rep("treatment3",7)) 
ordiplot(seedbank_byspp_nmds,type="n")
ordihull(seedbank_byspp_nmds,groups=treat,draw="polygon",col="grey90",
         label=FALSE)
orditorp(seedbank_byspp_nmds,display="species",col="red",air=0.01)
orditorp(seedbank_byspp_nmds,display="sites",col=c(rep("green",7),rep("blue",7),rep("orange",7)),
         air=0.01,cex=1.25)


```




