---
title: "Ch1_Degradation_abcoverNMDS"
author: "Stephanie Ma Lucero"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install libraries, message=FALSE}

library(tidyverse) # includes ggplot2, dplyr, tidyr, readr, purr, tibble, stringr, forcats
library(ggplot2) # data visualization
library(readr)
library(dplyr)
library(ggpubr) # customize ggplot2 for publication
library(broom)
library(AICcmodavg)
library(vegan)
library(readxl) # for .xls and .xlsx sheets
library(janitor)
library(calecopal) 
library(multcompView)
library(plotly)
if (!require(devtools)) {
  install.packages("devtools")
}
devtools::install_github("gavinsimpson/ggvegan")
```

# Load csv

```{r load csv}
# percent cover
abcover_full_raw <- read.csv("~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022b/deg_percentcover_byspecies_bylifeform.csv", header = TRUE, na.strings=c("","NA"))
```

### (1) Organize the percent cover data by converting the dataframe into a tibble.
```{r ab percent cover by status, include = TRUE}

str(abcover_full_raw)

head(abcover_full_raw)
tail(abcover_full_raw)

# convert dataframe to tibble
abcover_full_raw <- as_tibble(abcover_full_raw) 

class(abcover_full_raw) # now a tibble
```

### (2) Organize the percent cover data by reordering the classes: Intact, Matrix, Degraded.
```{r ab percent cover reorder classes to Intact Matrix Degraded}
abcover_full_v <- abcover_full_raw %>%
  mutate(standtype = fct_relevel(standtype, 
            "intact", "matrix", "degraded")) 
unique(abcover_full_v$standtype)

# renaming observations
abcover_full_v$standtype <- factor(abcover_full_v$standtype, levels=c("intact", "matrix", "degraded"), labels=c("Intact", "Matrix", "Degraded"))
unique(abcover_full_v$standtype)

```

### (3) Create new columns for SITE_REP, SITE_REP_DISTANCE, and STAND_STATUS for the percent cover data. 
```{r ab percent cover create site_rep column and site_rep_dist column}

# combine site and rep to one column
abcover_full_v$site_rep <- paste(abcover_full_v$site, abcover_full_v$rep, sep="_") 

# combine site_rep and distance to one column
abcover_full_v$site_rep_dist <- paste(abcover_full_v$site_rep, abcover_full_v$distance, sep="_") 

# combine standtype and native/nonnative status to one column
abcover_full_v$stand_status <- paste(abcover_full_v$standtype, abcover_full_v$status, sep="_") 

# combine site_rep and species to one column
abcover_full_v$site_rep_species <- paste(abcover_full_v$site_rep, abcover_full_v$species, sep="_") 

head(abcover_full_v) # look at new columns
glimpse( abcover_full_v)  
unique(abcover_full_v$site)

```

### (4) Omit "ground" as a status since I'm interested in only Native and Non-native species.
```{r ab percent cover filter for native and non-native only}
abcover_full_v_omitground <- filter(abcover_full_v, status !="ground")
head(abcover_full_v_omitground)
```

### (5) Identify the unique species at each site_rep_dist to get rid of duplicate species recordings. (e.g., Present at one distance = 1/41 to get percent cover) 
```{r ab percent cover unby species, inclue = TRUE}

abcover_byspecies <- abcover_full_v_omitground %>% 
  group_by(standtype, site, site_rep, species) %>% 
  filter(site_rep != "Intact4_4") %>% #omit Intact 4 from data
  filter(!is.na(species)) %>% # omit blank cells in species 
  dplyr::summarize(count_per_dist = n_distinct(site_rep_dist), percentcover = n_distinct(site_rep_dist)/41) %>% 
   select(!c(count_per_dist))
  

head(abcover_byspecies) # prints to new tab


unique(abcover_byspecies$standtype)
unique(abcover_byspecies$site)
unique(abcover_byspecies$site_rep)
unique(abcover_byspecies$species)

```

### (6) Prepare data for an NMDS (percent cover x species x standtype)
```{r transpose percent cover data by species, include=TRUE}

# use this for GROUPS: abcover_byspecies_nmdsprep

abcover_byspecies_nmdsprep <- abcover_byspecies %>% 
  tidyr::spread(species, percentcover) %>% # transpose data so columns are species
  replace(is.na(.), 0) # replace NA with zeros

view(abcover_byspecies_nmdsprep)
# 15 sites (rows) by 48 columns
str(abcover_byspecies_nmdsprep)


abcover_byspecies_nmdsprep_num <- abcover_byspecies_nmdsprep %>% 
  # use the ungroup() function to take out grouping
  ungroup() %>% 
  select(!c(standtype, site, site_rep))
# 15 sites (rows) by 33 species (columns)
# DEG 1-6, ENH 1-6, INT 1-3 = 15 transects
         
view(abcover_byspecies_nmdsprep_num) # only numbers
write.csv(abcover_byspecies_nmdsprep_num, "~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022b/processed/nmds_aboveground.csv")
```

### (7) Percent Cover
```{r above ground percent cover, include=TRUE}

above_percentcover <- abcover_full_v_omitground %>% 
  group_by(standtype, site_rep, species, lifeform, status, family, lifecycle) %>% 
  filter(site_rep != "Intact4_4") %>% #omit Intact 4 from data
  filter(!is.na(species)) %>% # omit blank cells in species 
  dplyr::summarize(count_per_dist = n_distinct(site_rep_dist), percentcover = n_distinct(site_rep_dist)/41) %>% 
   select(!c(count_per_dist))
  

view(above_percentcover) # prints to new tab


unique(abcover_byspecies$standtype)
unique(abcover_byspecies$site)
unique(abcover_byspecies$site_rep)
unique(abcover_byspecies$species)

```


# Above ground species richness
```{r species richness by stand type}
sppr <- specnumber(abcover_byspecies_nmdsprep_num)

summary(aov(sppr ~ standtype, data = abcover_byspecies_nmdsprep)) # by stand type
#             Df Sum Sq Mean Sq F value Pr(>F)
#standtype    2  17.40   8.700   1.921  0.189
#Residuals   12  54.33   4.528

summary(aov(sppr ~ site_rep, data = abcover_byspecies_nmdsprep)) # by stand type

abcover_byspecies_nmdsprep <- abcover_byspecies_nmdsprep %>% 
  mutate(standtype_ID = 1)

abcover_byspecies_nmdsprep$standtype_ID <- 1:15

sppr_df <- sppr %>% 
  enframe() %>% 
  full_join(abcover_byspecies_nmdsprep, by = c("name" = "standtype_ID"))


pal <- c("lightsalmon1", "gold1", "palegreen4")

plot_sppr <- ggplot(sppr_df, aes(x = standtype, y = value, fill = standtype)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  scale_x_discrete(labels = c("Intact \n (n = 9)", "Matrix \n (n = 10)", "Degraded \n (n = 7)")) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = "Stand Type",
       y = "Number of species per site",
       title = "Above ground species richness")
plot_sppr

```



```{r shannon diversity}
# this doesn't make sense for % cover, but I could run it on shrub density. I could also run it on the seedbank data. 
shannondiv <- diversity(abcover_byspecies_nmdsprep_num)
head(shannondiv)

sppdiv_aov <- aov(shannondiv ~ standtype, data = abcover_byspecies_nmdsprep)
summary(sppdiv_aov)

#     Df Sum Sq Mean Sq F value Pr(>F)  
#standtype    2 0.2899 0.14493   3.924 0.0489 *
#Residuals   12 0.4433 0.03694        

```

NMDS 
```{r nmds by species percent cover}

# the AUTOTRANSFORM = F skips the automatic data transformations and uses the input data as is
# - when the raw data contains the information you need (e.g., abundance)
# - when you have a small number of samples or small number of variables
# - when you want to preseve the original scale of the data (e.g., interested in the absolute differences betwen two sets of measurements)
abcover_byspecies_NMDS <- metaMDS(abcover_byspecies_nmdsprep_num, autotransform = F) # this is more appropriate for my data


# uses the RAUP-CRICK dissimilarity index to calculate pairwise dissimilarities
# - adjusts for spp richness
# - downweights importance of rare spp 
# - more useful when interested in community STRUCTURE rather than individual species
# - takes into account differences in species evenness
# abcover_byspecies_NMDS <- metaMDS(abcover_byspecies_nmdsprep_num, distance = "raup") 

#stressplot(abcover_byspecies_NMDS)
#plot(abcover_byspecies_NMDS)
```


```{r scores nmds}

scores(abcover_byspecies_NMDS)
```

```{r stress test}

stressplot(abcover_byspecies_NMDS)
# if stress tests are really low ==> I may have insufficient data. :/

#plot_df <- scores(seedbank_bysitecore_nmds, display = "sites") %>% # from An's code
#  as.data.frame() %>% 
#  rownames_to_column("stand_core") %>% 
#  full_join(seedbank_bysitecore, by = "stand_core")

nmds_above_df <- data.frame(abcover_byspecies_NMDS$points, abcover_byspecies_nmdsprep) # code from chatgpt


```


```{r ggplot above ground}
nmds_above_df_plotcolor <- ggplot(nmds_above_df, aes(x = MDS1, y = MDS2, color = standtype, shape = standtype)) +
  geom_point(size = 3, alpha = 0.8) +
  scale_color_manual(values = c("darkgreen", "darkmagenta", "darkgoldenrod1")) + # in color
  # scale_color_manual(values = c("gray0", "gray30", "gray60")) + # in bw
  theme_bw() +
  theme(legend.position = c(0.1, 0.85)) + # x, y left corner: 0,0
  stat_ellipse(level = .95, linetype = 2, linewidth = 1) +
  labs(title = "NMDS: by percent cover", color = "Stand type", shape = "Stand type")
nmds_above_df_plotcolor

```




```{r nmds color and vectors with envfit function}
# code from An Bui: https://rpubs.com/an-bui/vegan-cheat-sheet 

nmds_above_df <- data.frame(abcover_byspecies_NMDS$points, abcover_byspecies_nmdsprep) # code from chatgpt

abcover_byspecies_nmdsprep <- abcover_byspecies_nmdsprep %>% 
  mutate(standtype_ID = 1)

abcover_byspecies_nmdsprep$standtype_ID <- 1:15

plot_above_df <- scores(abcover_byspecies_NMDS, display = "sites") %>% 
  as.data.frame() %>% 
  rownames_to_column("standtype_ID") %>% 
  mutate_at(c('standtype_ID'), as.numeric) %>% 
  full_join(abcover_byspecies_nmdsprep, by = "standtype_ID") %>% 
  select(!c(standtype_ID))
view(plot_above_df)

# envfit() takes the output of metaMDS() and the species matrix you created
fit_above <- envfit(abcover_byspecies_NMDS, abcover_byspecies_nmdsprep, perm = 999) # pvalues for spp and stand, site, stand_core

# extract p-values for each species
fit_above_pvals <- fit_above$vectors$pvals %>% 
  as.data.frame() %>% 
  rownames_to_column("species") %>% 
  dplyr::rename("pvals" = ".") %>% 
  filter(!c(species == "standtype_ID"))
view(fit_above_pvals)

# extract coordinates for species, only keep species with p-val = 0.05
fit_above_spp <- fit_above %>% 
  scores(., display = "vectors") %>% 
  as.data.frame() %>% 
  rownames_to_column("species") %>% 
  full_join(., fit_above_pvals, by = "species") %>% 
  filter(pvals < 0.01) %>%  # limited number for visual clarity
  filter(!species == 'standtype_ID')
view(fit_above_spp)

# new plot
nmdsabove_plot_new <- ggplot(plot_above_df, aes(x = NMDS1, y = NMDS2)) +
  #coord_fixed() +
  geom_point(aes(color = standtype), size = 3, alpha = 0.8) +
  stat_ellipse(aes(color = standtype)) +
  scale_color_manual(values = c("darkgreen", "darkmagenta", "darkgoldenrod1"),
                     name = "Stand Type",
                     labels = c("Intact" = "Shrub dominated", 
                                "Matrix" = "Matrix", 
                                "Degraded" = "Grass dominated")) + #rename standtypes labels
  geom_segment(data = fit_above_spp, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")),
               col = "black") +
  geom_text(data = fit_above_spp, aes(label = species),  
            position = position_jitter(width = 0.1, height = 0.2)) 


nmds_above_print <- nmdsabove_plot_new + theme_bw() + theme(legend.position = c(0.125, 0.825))

ggsave("Ch1_Degradation_figures/nmds_above_print.png", nmds_above_print, width = 8, height = 6)

```





```{r byplot - ggplot vectors and measurments, include = FALSE}

#this makes a dataframe that gives a lot of control graphically 
abcover_byspecies_NMDS_fort <- fortify(abcover_byspecies_NMDS)

# vectors AND measurments
abcover_byspecies_NMDS_fortplot <- ggplot() + 
  geom_point(data = subset(abcover_byspecies_NMDS_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2, group = ""),
             color = 'black',
             size = 4,
             alpha = 1) +
  geom_segment(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               linewidth = 0.8) +
  geom_text(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1), 
            position = position_jitter(width = 0.2, height = 0.2)) +
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", linewidth = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", linewidth = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
abcover_byspecies_NMDS_fortplot

#only vectors shown
abcover_byspecies_NMDS_fortplot1 <- ggplot() + 
  geom_point(data = subset(abcover_byspecies_NMDS_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0) +
  geom_segment(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               linewidth = 0.8) +
  geom_text(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1),
            position = position_jitter(width = 0.9, height = 0.9)) +
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", linewidth = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
abcover_byspecies_NMDS_fortplot1

#only measurements shown
abcover_byspecies_NMDS_fortplot2 <- ggplot() + 
  geom_point(data = subset(abcover_byspecies_NMDS_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0.5) +
  geom_segment(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               size = 0.8,
               alpha = 0) +
  geom_text(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1),
            position = position_jitter(width = 0.2, height = 0.2),
            alpha = 0)+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", size = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
abcover_byspecies_NMDS_fortplot2

#interactive plots that I like
ggplotly(abcover_byspecies_NMDS_fortplot)

#show both plots 
hi_above <- ggarrange(nmds_above_df_plotcolor, abcover_byspecies_NMDS_fortplot, ncol = 1)
hi_above

```



NMDS
NMDS
NMDS by Plots! to determine which species are driving transect disimiliarity. 

(5) Identify the unique species at each site_rep_dist to get rid of duplicate species recordings. (e.g., Present at one distance = 1/41 to get percent cover) 
```{r ab percent cover unby species, inclue = TRUE}

abcover_byplot_groups <- abcover_full_v_omitground %>% 
  select(c(standtype, site, site_rep, species))
head(abcover_byplot_groups)

abcover_byplot_prep <- abcover_full_v_omitground %>% 
  group_by(standtype, site, site_rep, species) %>% 
  filter(site_rep != "Intact4_4") %>% #omit Intact 4 from data
  filter(!is.na(species)) %>% # omit blank cells in species 
  dplyr::summarize(count_per_dist = n_distinct(site_rep_dist), percentcover = n_distinct(site_rep_dist)/41) 
  
view(abcover_byplot_prep) # prints to new tab

abcover_byplot <- abcover_byplot_prep %>% 
  select(c(site_rep, species, percentcover))

write.csv(abcover_byplot, "~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022b/processed/abcover_byplot.csv")

abcover_byplot_load <- read.csv("~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022b/processed/abcover_byplot_onlysite_repsppcover.csv", header = TRUE, na.strings=c("","NA"))
```


(6) Prepare data for an NMDS (percent cover x species x standtype)
```{r transpose percent cover data by species, include=TRUE}

# use this for GROUPS: abcover_byplot_nmdsprep

head(abcover_byplot_load)

abcover_byplot_nmdsprep <- abcover_byplot_load %>% 
  tidyr::spread(site_rep, percentcover) %>% # transpose data so columns are species
  replace(is.na(.), 0) # replace NA with zeros
view(abcover_byplot_nmdsprep)
dim(abcover_byplot_nmdsprep)

abcover_byplot_nmdsprep_num <- abcover_byplot_nmdsprep %>% 
  # use the ungroup() function to take out grouping
  ungroup() %>% 
  select(!c(species))
# 33 species (rows) by 15 sites (columns)
# DEG 1-6, ENH 1-6, INT 1-3 = 15 transects
dim(abcover_byplot_nmdsprep_num)
```

```{r write nmds to csv}
view(abcover_byplot_nmdsprep_num) # only numbers
write.csv(abcover_byplot_nmdsprep_num, "~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022b/processed/nmds_aboveground.csv")
```

(7) NMDS 
```{r stress tests}
# USE THIS ONE!

# str(seedbank_bysitecore_nmdsprep_num)
# stand_core_depth -->  tibble [89 × 52]

# stress test, will run 20 times
seedbank_byplot_nmds <- metaMDS(abcover_byplot_nmdsprep_num, distance = "raup")

# stress test, will run 20 times
seedbank_byplot_nmds <- metaMDS(abcover_byplot_nmdsprep_num, autotransform = F)
```

```{r scores nmds}

scores(seedbank_byplot_nmds)

```


```{r stress test}

stressplot(seedbank_byplot_nmds)
# if stress tests are really low ==> I may have insufficient data. :/

#plot_df <- scores(seedbank_bysitecore_nmds, display = "sites") %>% # from An's code
#  as.data.frame() %>% 
#  rownames_to_column("stand_core") %>% 
#  full_join(seedbank_bysitecore, by = "stand_core")

nmds_df <- data.frame(seedbank_byplot_nmds$points, # unnamed column is species n = 33
                      abcover_byplot_nmdsprep) # from chatgpt
view(nmds_df)
```

```{r ggplot bw}
ggplot(nmds_df, aes(x = MDS1, y = MDS2, color = stand, shape = stand)) +
  geom_point(size = 3, alpha = 0.8) +
  scale_color_manual(values = c("gray0", "gray30", "gray60")) +
  theme_bw() +
  stat_ellipse(linetype = 2, size = 1) +
  labs(title = "NMDS: by seed bank germination counts", color = "Stand type", shape = "Stand type")
```

```{r with vectors}

#this makes a dataframe that gives a lot of control graphically 
seedbank_byplot_nmds_fort <- fortify(seedbank_byplot_nmds)

# vectors AND measurments
seedbank_byplot_nmdsfortplot <- ggplot() + 
  geom_point(data = subset(seedbank_byplot_nmds_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0.5) +
  geom_segment(data = subset(seedbank_byplot_nmds_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               linewidth = 0.8) +
  geom_text(data = subset(seedbank_byplot_nmds_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1))+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", linewidth = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", linewidth = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

#only vectors shown
seedbank_byplot_nmds_fortplot1 <- ggplot() + 
  geom_point(data = subset(seedbank_byplot_nmds_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0) +
  geom_segment(data = subset(seedbank_byplot_nmds_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               linewidth = 0.8) +
  geom_text(data = subset(seedbank_byplot_nmds_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1))+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", linewidth = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
seedbank_byplot_nmds_fortplot1
```
```{r plot measurements only }
#only measurements shown
seedbank_byplot_nmds_fortplot2 <- ggplot() + 
  geom_point(data = subset(seedbank_byplot_nmds_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0.5) +
  geom_segment(data = subset(seedbank_byplot_nmds_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               size = 0.8,
               alpha = 0) +
  geom_text(data = subset(seedbank_byplot_nmds_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1),
            alpha = 0)+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", size = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

#interactive plots that I like
ggplotly(seedbank_bysitecore_nmds_fortplot)

#show both plots 
hi <- ggarrange(seedbank_bysitecore_nmds_fortplot1, seedbank_bysitecore_nmds_fortplot2, ncol = 1)
hi

```


```{r}

ggplot(nmds_df, aes(x = MDS1, y = MDS2)) +
  geom_point()+
  theme_bw() +
  stat_ellipse(linetype = 2, size = 1) +
  labs(title = "NMDS: by seed bank germination counts", color = "Stand type", shape = "Stand type")

#
ggplot(nmds_df, aes(x = MDS1, y = MDS2, color = stand, shape = stand)) +
  geom_point(size = 3, alpha = 0.8) +
  scale_color_manual(values = c("gray0", "gray30", "gray60")) +
  theme_bw() +
  stat_ellipse(linetype = 2, size = 1) +
  labs(title = "NMDS: by seed bank germination counts", color = "Stand type", shape = "Stand type")
```

```{r ggplot color}
ggplot(nmds_df, aes(x = MDS1, y = MDS2, color = stand, shape = stand)) +
  geom_point(size = 3, alpha = 0.8) +
  scale_color_manual(values = c("darkgreen", "darkmagenta", "darkgoldenrod1")) +
  theme_bw() +
  #stat_ellipse(level = .95, linetype = 2, size = 1) +
  stat_ellipse(linetype = 2, size = 1) +
  labs(title = "NMDS: by seed bank germination counts", color = "Stand type", shape = "Stand type")


```



```{r}
# useful code on how to color code NMDS from YouTube: https://www.youtube.com/watch?v=h7OrVmT7Ja8&t=446s
# and how to group NMDS values: https://www.youtube.com/watch?v=Y0GI34S-ZMI

scores(abcover_byspecies_NMDS) # structure of the file, look at the points to get the matrix of numbers and columns (2 columns)
abcover_byspecies_NMDS$points # to get the points/values of MDS1 and MDS2
scores(abcover_byspecies_NMDS) # scores of your variables
plot(abcover_byspecies_NMDS)

abcover_byspecies_NMDS_centroid <- scores(abcover_byspecies_NMDS$species) %>% 
  as_tibble(rownames="species") %>% 
  inner_join(., abcover_byspecies_nmdsprep, by="species") %>% 
  group_by(standtype) %>% 
 summarize(axis1 = mean(MDS1),
            axis2 = mean(MDS2))

scores(abcover_byspecies_NMDS$species) %>% 
  as_tibble(rownames="species") %>% 
  inner_join(., abcover_bysite_nmdsprep, by="species") %>% 
  ggplot(aes(x=MDS1, y=MDS2, color=standtype)) +
  geom_point() +
  geom_point(data=abcover_byspecies_NMDS_centroid, aes(x=axis1, y=axis2, color=standtype),
             shape = 15, size = 5)

# How can I make interactive data points with call outs when I hover over them?
```

```{r FROM KIT - nmds by species percent cover using code from KIT}
# USE THIS ONE!

# stress test, will run 20 times
abcover_byspecies_NMDS <- metaMDS(abcover_byspecies_nmdsprep_num, 
                                  autotransform = F)
# stress tests are really low ==> I may have insufficient data. :/

stressplot(seedbank_byplot_nmds)
plot(seedbank_byplot_nmds)

#basic normal plot
ordiplot(abcover_byspecies_NMDS) # without point labels
ordiplot(abcover_byspecies_NMDS, type = "t") # with point labels

#ggvegan to make a quick nice plot
autoplot(abcover_byspecies_NMDS) # NMDS1 x NMDS2, Key = "Score": sites, species



#this makes a dataframe that gives a lot of control graphically 
abcover_byspecies_NMDS_fort <- fortify(abcover_byspecies_NMDS)

abcover_byspecies_NMDS_fortplot <- ggplot() + 
  geom_point(data = subset(abcover_byspecies_NMDS_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0.5) +
  geom_segment(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               linewidth = 0.8) +
  geom_text(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1))+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", linewidth = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", linewidth = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

#only vectors shown
abcover_byspecies_NMDS_fortplot1 <- ggplot() + 
  geom_point(data = subset(abcover_byspecies_NMDS_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0) +
  geom_segment(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               linewidth = 0.8) +
  geom_text(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1))+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", linewidth = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

#only measurements shown
abcover_byspecies_NMDS_fortplot2 <- ggplot() + 
  geom_point(data = subset(abcover_byspecies_NMDS_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0.5) +
  geom_segment(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               size = 0.8,
               alpha = 0) +
  geom_text(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1),
            alpha = 0)+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", size = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

#interactive plots that I like
ggplotly(abcover_byspecies_NMDS_fortplot)

#show both plots 
hi <- ggarrange(abcover_byspecies_NMDS_fortplot1, abcover_byspecies_NMDS_fortplot2, ncol = 1)
hi


ordiplot(abcover_byspecies_NMDS,type="n")
orditorp(abcover_byspecies_NMDS,display="species",col="red",air=0.01)
orditorp(abcover_byspecies_NMDS,display="sites",cex=1.25,air=0.01)

str(abcover_byspecies_nmdsprep_num)
#tibble [125 × 34] (S3: tbl_df/tbl/data.frame)
treat=c(rep("treatment1",50),rep("treatment2",50), rep("treatment3",25)) 
ordiplot(abcover_byspecies_NMDS,type="n")
ordihull(abcover_byspecies_NMDS,groups=treat,draw="polygon",col="grey90",
         label=FALSE)
orditorp(abcover_byspecies_NMDS,display="species",col="red",air=0.01)
orditorp(abcover_byspecies_NMDS,display="sites",col=c(rep("green",50),rep("blue",50),re("orange",25)),
         air=0.01,cex=1.25)


```


(7) NMDS by site (e.g., Intact_1)
```{r transpose site percent cover data}

abcover_bysite_nmdsprep <- abcover_byspecies %>% 
  tidyr::spread(site_rep, percentcover) %>% # transpose data so columns are site_rep
  replace(is.na(.), 0) # replace NA with zeros
view(abcover_bysite_nmdsprep)

abcover_bysite_nmdsprep_num <- abcover_bysite_nmdsprep %>% 
  # use the ungroup() function to take out grouping
  ungroup() %>% 
  select(!c(standtype, site, species, status, lifeform, lifecycle, count_per_dist))
         
view(abcover_bysite_nmdsprep_num) # only numbers
```


```{r FROM KIT - nmds by site percent cover}

# stress test with DISTANCE, runs 20 times
# abcover_bysite_NMDS <- metaMDS(abcover_bysite_nmdsprep_num, distance = "raup") 
#In metaMDS(abcover_bysite_nmdsprep_num, distance = "raup") :
#  stress is (nearly) zero: you may have insufficient data

# stress test with AUTOTRANSFORM, runs 20 times
abcover_bysite_NMDS <- metaMDS(abcover_bysite_nmdsprep_num, autotransform = F) 
#In metaMDS(abcover_bysite_nmdsprep_num, autotransform = F) :
#  stress is (nearly) zero: you may have insufficient data

stressplot(abcover_bysite_NMDS)

plot(abcover_bysite_NMDS) 

#basic normal plot
ordiplot(abcover_bysite_NMDS) # without point labels
ordiplot(abcover_bysite_NMDS, type = "t") # with point labels

#ggvegan to make a quick nice plot
autoplot(abcover_bysite_NMDS) # NMDS1 x NMDS2, Key = "Score": sites, species



#this makes a dataframe that gives a lot of control graphically 
abcover_bysite_NMDS_fort <- fortify(abcover_bysite_NMDS)

abcover_bysite_NMDS_fortplot <- ggplot() + 
  geom_point(data = subset(abcover_bysite_NMDS_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0.5) +
  geom_segment(data = subset(abcover_bysite_NMDS_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               linewidth = 0.8) +
  geom_text(data = subset(abcover_bysite_NMDS_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1))+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", linewidth = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", linewidth = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

#only vectors shown
abcover_bysite_NMDS_fortplot1 <- ggplot() + 
  geom_point(data = subset(abcover_bysite_NMDS_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0) +
  geom_segment(data = subset(abcover_bysite_NMDS_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               linewidth = 0.8) +
  geom_text(data = subset(abcover_bysite_NMDS_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1))+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", linewidth = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

#only measurements shown
abcover_bysite_NMDS_fortplot2 <- ggplot() + 
  geom_point(data = subset(abcover_bysite_NMDS_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0.5) +
  geom_segment(data = subset(abcover_bysite_NMDS_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               size = 0.8,
               alpha = 0) +
  geom_text(data = subset(abcover_bysite_NMDS_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1),
            alpha = 0)+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", size = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

#interactive plots that I like
ggplotly(abcover_bysite_NMDS_fortplot)

#show both plots 
hi_site <- ggarrange(abcover_bysite_NMDS_fortplot1, abcover_bysite_NMDS_fortplot2, ncol = 1)
hi_site


ordiplot(abcover_bysite_NMDS,type="n")
orditorp(abcover_bysite_NMDS,display="species",col="red",air=0.01)
orditorp(abcover_bysite_NMDS,display="sites",cex=1.25,air=0.01)

str()
#tibble [125 × 34] (S3: tbl_df/tbl/data.frame)
treat=c(rep("treatment1",3),rep("treatment2",3), rep("treatment3",3)) 
ordiplot(abcover_bysite_NMDS,type="n")
ordihull(abcover_bysite_NMDS,groups=treat,draw="polygon",col="grey90",
         label=FALSE)
orditorp(abcover_bysite_NMDS,display="species",col="red",air=0.01)
orditorp(abcover_bysite_NMDS,display="sites",col=c(rep("green",6),rep("blue",6),re("orange",3)),
         air=0.01,cex=1.25)




```





```{r set seed }
# *** No convergence -- monoMDS stopping criteria:
#    15: no. of iterations >= maxit
#     1: stress < smin
#     4: scale factor of the gradient < sfgrmin

# set.seed(100)
abcover_bysite_NMDS <- metaMDS(abcover_bysite_nmdsprep_num, distance = "raup") # Runs with stress and a number

str(abcover_bysite_NMDS) # structure of the file, look at the points to get the matrix of numbers and columns (2 columns)
abcover_bysite_NMDS$points # to get the points/values of MDS1 and MDS2
scores(abcover_bysite_NMDS) # scores of your variables
plot(abcover_bysite_NMDS)
```




```{r sandbox nmds}

plot(abcover_bysite_NMDS, type = "n") #displays empty ordination space

points(abcover_bysite_NMDS, display = "sites", pch = c(16, 8, 17, 11) [as.numeric(abcover_bysite_nmdsprep$status)], col = c("blue", "orange", "black") [as.numeric(abcover_bysite_nmdsprep$count_per_dist)]) # displays site points where symbols (pch) are different management options and colour (col) are different land uses

# points(abcover_bysite_NMDS, display = "site_rep", 
#       pch = c(16, 8, 17, 11) [as.numeric(abcover_bysite_nmdsprep_num$site_rep)], 
#       col = c("blue", "orange", "black", "green", "pink", "red", "purple", "teal", #"magenta", "cyan", "yellow", "brown", "light blue", "peach", "lime"),
#       [as.numeric(abcover_bysite_nmdsprep_num$Use)]) # displays site points where #symbols (pch) are different management options and colour (col) are different land uses

legend("topright", legend = c(levels(abcover_bysite_nmdsprep_num$site_rep),
                              levels(abcover_bysite_nmdsprep_num$percent)), 
       pch = c(16, 8, 17, 11, 16, 16, 16), 
       col = c("black","black","black","black","blue", "orange", "black"), 
       bty = "n", cex = 1) # displays symbol and colour legend

legend("topleft", "stress = 0.118", bty = "n", cex = 1) # displays legend text of stress value 

nmds <- ordiplot(abcover_bysite_NMDS, type = "none")
ordipointlabel(abcover_bysite_NMDS,display = "sites",add=TRUE)
```




