---
title: "Ch1_Degradation_abcoverNMDS"
author: "Stephanie Ma Lucero"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install libraries, message=FALSE}

library(tidyverse) # includes ggplot2, dplyr, tidyr, readr, purr, tibble, stringr, forcats
library(ggplot2) # data visualization
library(readr)
library(dplyr)
library(ggpubr) # customize ggplot2 for publication
library(broom)
library(AICcmodavg)
library(vegan)
library(readxl) # for .xls and .xlsx sheets
library(janitor)
library(calecopal) 
library(multcompView)
library(plotly)
if (!require(devtools)) {
  install.packages("devtools")
}
devtools::install_github("gavinsimpson/ggvegan")
library(here)
library(ggpmisc)
```

# Point-intercept transect data
### load csv for percent cover

```{r load csv}
# Stand colors: INTACT, MATRIX, DEGRADED 
cal_palette(name = "chaparral3", n = 3, type = "continuous") 
deg <- "#D3E3CA"
mat <- "#92A587"
int <- "#2F3525"
standcolors <- c(int, mat, deg)

# Lifeform colors: shrub, grass, forb
cal_palette(name = "chaparral3", n = 3, type = "continuous") 
shrub <- "#D3E3CA"
grass <- "#92A587"
forb <- "#2F3525"
lifeformcolors <- c(int, mat, deg)

# Status colors: native, exotic 
cal_palette(name = "chaparral1", n = 6, type = "continuous") 
native <- "#AEBFA8"
exotic <- "#F19B34"
statuscolors <- c(native, exotic)

# percent cover
abcover_full_raw <- read.csv("deg_percentcover_byspecies_bylifeform.csv", header = TRUE, na.strings=c("","NA"))

abcover_full_raw <- abcover_full_raw %>% dplyr::select(!c(species.1))
glimpse(abcover_full_raw) # point-intercept transect data

```

```{r species diversity from point-line intercept transect, include = TRUE}
names(abcover_full_raw)
sort(unique(abcover_full_raw$species))
glimpse(abcover_full_raw %>%  na.omit())
# Rows: 2,063
# Columns: 10

# Renaming observations
ab_diversity_load <- abcover_full_raw %>%
  #dplyr::select(!c(species == "bare_ground")) 
   mutate(species = case_when(
    species == "Avena barbata dead" ~ "Avena barbata", # old name ~ new name
        species == "Bromus diandrus dead" ~ "Bromus diandrus", 
           species == "Bromus rubens dead" ~ "Bromus rubens", 
       species == "Bromus spp." ~ "Bromus rubens", 
    species == "Bromus spp. dead" ~ "Bromus rubens",
          species == "Bromus madritensis" ~ "Bromus rubens",
     species == "Centaurea melitensis dead" ~ "Centaurea melitensis",
    species == "Erodium cicutarium" ~ "Erodium spp.", 
 #   species == "Elymus condensatus dead" ~ "Elymus condensatus",
 #   species == "Heteromeles arbutifolia dead" ~ "Heteromeles arbutifolia",
    species == "Hirschfeldia incana dead" ~ "Hirschfeldia incana",
 #   species == "Malacothamnus fasciculatus dead" ~ "Malacothamnus fasciculatus",
      #     species == "Salvia leucophylla dead" ~ "Salvia leucophylla",
     # species == "Stipa lepida dead" ~ "Stipa lepida",
    TRUE ~ as.character(species)  # Keep other observations unchanged
  )) %>% 
  mutate(site = case_when(
    site == "Intact_1" ~ "Intact1",
      site == "Intact_2" ~ "Intact2",
      site == "Intact_3" ~ "Intact3",
    TRUE ~ as.character(site)  # Keep other observations unchanged
  )) %>% 
  mutate(standtype = fct_relevel(standtype, 
            "intact", "matrix", "degraded")) %>% 
   mutate(site = fct_relevel(site, 
            "Intact1", "Intact2", "Intact3", "ENH1",    "ENH2", "DEG1",    "DEG2")) %>% 
  filter(species !="native_litter" ) %>% 
  filter(species != "nonnative_thatch") %>% 
  filter(species != "bare_ground") 

sort(unique(ab_diversity_load$species))
sort(unique(ab_diversity_load$site))
glimpse(ab_diversity_load)
#Rows: 1,450
#Columns: 10

# combine site and rep to one column
ab_diversity_load$site_rep <- paste(ab_diversity_load$site, ab_diversity_load$rep, sep="_") 
ab_diversity_load$site_rep_species <- paste(ab_diversity_load$site_rep, ab_diversity_load$species, sep="_") 
sort(unique(ab_diversity_load$site_rep)) 
#"DEG1_1"    "DEG1_2"    "DEG1_3"    "DEG2_4"    "DEG2_5"    "DEG2_6"    
#"ENH1_1"    "ENH1_2"    "ENH1_3"    "ENH2_4"    "ENH2_5"    "ENH2_6"    
#"Intact1_1" "Intact2_2" "Intact3_3"

glimpse(ab_diversity_load)  # species diversity from point-line intercept transect
# Rows: 1,445
#Columns: 10

ab_diversity <- ab_diversity_load %>% 
  dplyr::select(!c(distance, vertical))

#write.csv(ab_diversity, "processed/ab_diversity.csv")
```

```{r join point intercept sppr with native and non-native percent cover, include=TRUE}

ab_diversity # species richness from point intercept transects
sort(unique(ab_diversity$site_rep))
# "DEG1_1"    "DEG1_2"    "DEG1_3"    "DEG2_4"    "DEG2_5"    "DEG2_6"    "ENH1_1"    "ENH1_2"    "ENH1_3"    "ENH2_4"    "ENH2_5"    "ENH2_6"   
#"Intact1_1" "Intact2_2" "Intact3_3"
glimpse(ab_diversity)
# Rows: 1,450
# Columns: 10

nativecover <- read.csv("processed/abcover_nativecover_nngrasscover.csv") # native percent cover from ch1_standdeg_canopy.Rmd
sort(unique(nativecover$site_rep))
# "DEG1_1"     "DEG1_2"     "DEG1_3"     "DEG2_4"     "DEG2_5"     "DEG2_6"     "ENH1_1"     "ENH1_2"     "ENH1_3"     "ENH2_4"     "ENH2_5"    
# "ENH2_6"     "Intact_1_1" "Intact_2_2" "Intact_3_3"
glimpse(nativecover)

nativecover2 <- nativecover %>% 
  mutate(site_rep = case_when(
    site_rep == "Intact_1_1" ~ "Intact1_1", # old name ~ new name
      site_rep == "Intact_2_2" ~ "Intact2_2",
      site_rep == "Intact_3_3" ~ "Intact3_3",
    TRUE ~ as.character(site_rep)  # Keep other observations unchanged
  )) 


nativecover3 <- nativecover2 %>% 
  dplyr::select(!c(X, standtype.x, standtype.y ))

ab_diversity_native_nn <- ab_diversity %>% # join above ground species diversity with native transect percent cover
  left_join(nativecover3, by = "site_rep") 

glimpse(ab_diversity_native_nn)
# Rows: 1,450
# Columns: 20

sort(unique(ab_diversity_native_nn$species))
write.csv(ab_diversity_native_nn, "processed/ab_diversity_native_nn.csv") # <-- export data to EXCEL and make graphs there

```

```{r summary tables for point intercept transect figure, include=TRUE}

summary_table_pttrans_sppr_tot <- ab_diversity_native_nn %>%
  group_by(standtype, site_rep, nonnative_grass, native_shrub) %>%
    mutate(species = case_when(
    species == "Elymus condensatus dead" ~ "Elymus condensatus", # old name ~ new name
        species == "Heteromeles arbutifolia dead" ~ "Heteromeles arbutifolia", 
    species == "Malacothamnus fasciculatus dead" ~ "Malacothamnus fasciculatus",
    species == "Salvia leucophylla dead" ~ "Salvia leucophylla", 
    species == "Stipa lepida dead" ~ "Stipa lepida",
    TRUE ~ as.character(species)  # Keep other observations unchanged
  )) %>% 
  summarize(total_sppr = n_distinct(species))

#  standtype site_rep  nonnative_grass native_shrub total_sppr
#    <fct>     <chr>               <dbl>        <dbl>      <int>
#  1 intact    Intact1_1          0.244         0.707          9
#  2 intact    Intact2_2          0.0732        0.707         12
#  3 intact    Intact3_3          0.122         0.951          6
#  4 matrix    ENH1_1             0.756         0.537         12
#  5 matrix    ENH1_2             0.659         0.585         10
#  6 matrix    ENH1_3             0.463         0.805         14
#  7 matrix    ENH2_4             0.878         0.415         10
#  8 matrix    ENH2_5             0.829         0.293          8
#  9 matrix    ENH2_6             0.829         0.463          8
# 10 degraded  DEG1_1             1             0.390          9
# 11 degraded  DEG1_2             0.902         0.390          7
# 12 degraded  DEG1_3             0.976         0.195          7
# 13 degraded  DEG2_4             0.976         0.268          7
# 14 degraded  DEG2_5             0.976         0.220          6
# 15 degraded  DEG2_6             0.902         0.171         10

summary_table_pttrans_nat <- ab_diversity_native_nn %>%
  filter(status == "native") %>% 
 group_by(standtype, site_rep, nonnative_grass, native_shrub) %>%
  summarize(native_sppr = n_distinct(species))

summary_table_pttrans_nn <- ab_diversity_native_nn %>%
  filter(status == "nonnative") %>% 
 group_by(standtype, site_rep, nonnative_grass, native_shrub) %>%
  summarize(nn_sppr = n_distinct(species))

#checking the data
ab_diversity_native_nn %>% 
  filter(site_rep == "DEG1_1") %>% 
   # filter(status == "native") %>%
   filter(status == "nonnative") %>%  
  distinct(species)


```
### point-intercept sppr figures
```{r figure point intercept transect - total sppr by nn cover}

ab_diversity_totalsppr_regression <- ggplot(data = summary_table_pttrans_sppr_tot, 
                                            aes(x = nonnative_grass*100, y = total_sppr, shape = standtype, color = standtype)) + #color = standtype
  geom_point(size = 3, alpha = 0.8) +
  scale_shape_manual(values = c(16, 17, 18)) +
  ylim(0, 20)+
  xlim(0,100) +
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", color = "black") +  # lm, loess, glm
  scale_color_manual(values = c(standcolors),
                     name = "Stand Type",
                     labels = c("intact" = "shrub dominated", 
                                "matrix" = "matrix", 
                                "degraded" = "grass dominated")) + #rename standtypes labels
  labs(x = "Non-native grass cover (%)", 
       y = "Total species richness (#)", 
       title = "Total species richness by Non-native grass cover") +
  theme_bw() +  
  stat_poly_line() +
 stat_poly_eq(use_label(c("eq", "R2")))

print(ab_diversity_totalsppr_regression)

# I made another figure in excel sheet: sppr_pointintercept.xlsx

```
```{r figure point intercept transect - native sppr by nn cover}

ab_diversity_nativesppr_regression <- ggplot(data = summary_table_pttrans_nat, 
                                            aes(x = nonnative_grass*100, y = native_sppr, shape = standtype, color = standtype)) + #color = standtype
  geom_point(size = 3, alpha = 0.8) +
  scale_shape_manual(values = c(16, 17, 18)) +
  ylim(0, 20)+
  xlim(0,100) +
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", color = "black") +  # lm, loess, glm
  scale_color_manual(values = c(standcolors),
                     name = "Stand Type",
                     labels = c("intact" = "shrub dominated", 
                                "matrix" = "matrix", 
                                "degraded" = "grass dominated")) + #rename standtypes labels
  labs(x = "Non-native grass cover (%)", 
       y = "Native species richness (#)", 
       title = "Native richness by Non-native grass cover") +
  theme_bw() +  
  stat_poly_line() +
 stat_poly_eq(use_label(c("eq", "R2")))
print(ab_diversity_nativesppr_regression)

```

```{r figure point intercept transect - nonnative sppr by nn cover}

ab_diversity_nnsppr_regression <- ggplot(data = summary_table_pttrans_nn, 
                                            aes(x = nonnative_grass*100, y = nn_sppr, shape = standtype, color = standtype)) + #color = standtype
  geom_point(size = 3, alpha = 0.8) +
  scale_shape_manual(values = c(16, 17, 18)) +
  ylim(0, 20)+
  xlim(0,100) +
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", color = "black") +  # lm, loess, glm
  scale_color_manual(values = c(standcolors),
                     name = "Stand Type",
                     labels = c("intact" = "shrub dominated", 
                                "matrix" = "matrix", 
                                "degraded" = "grass dominated")) + #rename standtypes labels
  labs(x = "Non-native grass cover (%)", 
       y = "Native species richness (#)", 
       title = "Non-native richness by Non-native grass cover") +
  theme_bw() +  
  stat_poly_line() +
 stat_poly_eq(use_label(c("eq", "R2")))
print(ab_diversity_nnsppr_regression)

```

# Belt transect data
### load data
```{r SHRUB species richness by belt transect, include = TRUE}
belttrans_byspecies <- read.csv("deg_beltdensity_byspecies.csv")
sort(unique(belttrans_byspecies$species))
sort(unique(belttrans_byspecies$site))
sort(unique(belttrans_byspecies$origin))
sort(unique(belttrans_byspecies$lifespan))

belttrans_sppr <- belttrans_byspecies %>%  
   mutate(species = case_when(
    species == "Bromus madritensis" ~ "Bromus rubens", # old name ~ new name
        species == "Erodium cicutarium" ~ "Erodium spp.",      
#    species == "Lupinus spp." ~ "Lupinus bicolor", # lupins are in different transects so leaving as three seperate entries: LUBI, LUSU, and Lupinus spp.
#     species == "Lupinus spp." ~ "Lupinus succulentus",
    TRUE ~ as.character(species)  # Keep other observations unchanged
  )) %>% 
      mutate(origin = case_when(
        origin == "non-native" ~ "nonnative", 
        TRUE ~ as.character(origin))) %>% 
  mutate(lifespan = case_when(
    lifespan == "Perennial" ~ "perennial", 
     TRUE ~ as.character(lifespan))) %>% 
    rename(rep = transectstand, #rename(new_name = old_name) 
           status = origin) %>% 
   filter(site != "ENH2") # why am I omitting ENH2? (feb 7, 2024) <-- STEPHANIE

belttrans_sppr$site <- factor(belttrans_sppr$site, levels=c("Intact1", "Intact2", "Intact3", "ENH1", "DEG1", "DEG2"))
unique(belttrans_sppr$site)
unique(belttrans_sppr$species)
  
# combine site and rep to one column
belttrans_sppr$site_rep <- paste(belttrans_sppr$site, belttrans_sppr$rep, sep="_") 
belttrans_sppr$site_rep_species <- paste(belttrans_sppr$site_rep, belttrans_sppr$species, sep="_") 

belttrans_sppr <- belttrans_sppr %>% 
  dplyr::select(!c("transectsite", "lifestage")) 
sort(unique(belttrans_sppr$site))

glimpse(belttrans_sppr)
sort(unique(belttrans_sppr$site_rep))
sort(unique(belttrans_sppr$site))
sort(unique(belttrans_sppr$species))



write.csv(belttrans_sppr, "processed/belttrans_sppr.csv")
```

```{r belt transect join percent cover, include=TRUE}
belttrans_sppr_native <- belttrans_sppr %>% # join above ground species diversity with native transect percent cover
  left_join(nativecover3, by = "site_rep") 
sort(unique(belttrans_sppr_native$site_rep))

#view(belttrans_sppr_native_nn)
glimpse(belttrans_sppr_native_nn)
# Rows: 904
# Columns: 21 

#write.csv(belttrans_sppr_native_nn, "processed/belttrans_sppr_native_nn.csv") #<-- export to EXCEL to create belt transect sppr figures.

```

```{r join sppr from belt and point-intercept transect}
# nov. 6, 2023
# KIT - Can you create a df of the total species list (point intercept transect + belt transect species list) at INT 1-3, ENH 1-3, and DEG 1-6?

belt_sppr <- belttrans_sppr %>% #select only site_rep_species from belt trans data
  dplyr::select(c(site_rep_species)) 

pointintercept_sppr <- ab_diversity_native_nn %>% #select only site_rep_species from belt trans data
  dplyr::select(c(site_rep_species))

belt_point_join_sppr <- belt_sppr %>% 
  full_join(pointintercept_sppr, by = "site_rep_species")

belt_point_join_sppr %>% 
  summarize(total_sppr = n_distinct(site_rep_species))
# ahhh.... standtype, site_rep, status, etc. needs to be brougt in from somewhere.... 


```


```{r summary tables for belt transect figure, include=TRUE}

belttrans_sppr_native_nn$site_rep <- factor(belttrans_sppr_native_nn$site_rep, levels=c("Intact1_1", "Intact2_2", "Intact3_3", "ENH1_1", "ENH1_2", "ENH1_3", "DEG1_1", "DEG1_2","DEG1_3", "DEG2_4","DEG2_5", "DEG2_6"))

belttrans_sppr_native_nn_sppr_tot <- belttrans_sppr_native_nn %>%
  group_by(site_rep) %>%
  summarize(total_sppr = n_distinct(species))
print(belttrans_sppr_native_nn_sppr_tot)
# site_rep. total_sppr
# Intact1_1	18			
# Intact2_2	21			
# Intact3_3	17			
# ENH1_1	13			
# ENH1_2	16			
# ENH1_3	9			
# DEG1_1	11			
# DEG1_2	8			
# DEG1_3	9			
# DEG2_4	10	
# DEG2_5	7			
# DEG2_6	6	


summary_table_belt_nat <- belttrans_sppr_native_nn %>%
  filter(status == "native") %>% 
 group_by(site_rep) %>%
  summarize(native_sppr = n_distinct(species))
print(summary_table_belt_nat)
# site_rep. native_sppr
# Intact1_1	15			
# Intact2_2	17			
# Intact3_3	13			
# ENH1_1	13			
# ENH1_2	14			
# ENH1_3	8			
# DEG1_1	7			
# DEG1_2	8			
# DEG1_3	6			
# DEG2_4	10	
# DEG2_5	7			
# DEG2_6	6

#checking the data
belttrans_sppr_native_nn %>% 
  filter(site_rep == "DEG2_6") %>% 
    filter(status == "native") %>%
   #filter(status == "nonnative") %>%  
  distinct(species)

```
```{r}




# Intact_1 = 15
# I_2 = 17
# I_3 = 13
# ENH_1 = 13
# E_2 = 14
# E_3 = 8
# DEG_1 = 7 
# D_2 = 8
# D_3 = 6
# D_4 = 10
# D_5 = 7
# D_6 = 6

summary_table_belt_nn <- belttrans_sppr_native_nn %>%
  filter(status == "nonnative") %>% 
 group_by(standtype, site_rep, nonnative_grass, native_shrub, percentcover_native, percentcover_nonnative) %>%
  summarize(nn_sppr = n_distinct(species))

#checking the data
belttrans_sppr_native_nn %>% 
  filter(site_rep == "DEG1_1") %>% 
   # filter(status == "native") %>%
   filter(status == "nonnative") %>%  
  distinct(species)

# SPECIES = 9
# Avena barbata				
# Hirschfeldia incana				
# Lupinus spp.				
# Bromus rubens				
# Salvia leucophylla				
# Erodium spp.				
# Dipterostemon capitatus				
# Artemisia californica				
# Acmispon maritimus

# NATIVE SPECEIS = 5
# # Lupinus spp.				
# Salvia leucophylla				
# Dipterostemon capitatus				
# Artemisia californica				
# Acmispon maritimus

# NONNATIVE SPECIES = 4
# Avena barbata				
# Hirschfeldia incana				
# Bromus rubens				
# Erodium spp.


```

### belt transect sppr - total sppr x nngrass
```{r belt transect sppr figure - total sppr by nngrass, include = TRUE}

belttrans_sppr_totalsppr_regression <- ggplot(data = belttrans_sppr_native_nn_sppr_tot,  #belttrans_sppr_native_nn
                                            aes(x = nonnative_grass*100, 
                                                y = total_sppr, 
                                                shape = standtype, 
                                                color = standtype)) + #color = standtype
  geom_point(size = 3, alpha = 0.8) +
  scale_shape_manual(values = c(16, 17, 18)) +
  ylim(0, 20)+
  xlim(0,100) +
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", color = "black") +  # lm, loess, glm
  scale_color_manual(values = c(standcolors),
                     name = "Stand Type",
                     labels = c("intact" = "shrub dominated", 
                                "matrix" = "matrix", 
                                "degraded" = "grass dominated")) + #rename standtypes labels
  labs(x = "Non-native grass cover (%)", 
       y = "Total species richness (#)", 
       title = "Total species richness by Non-native grass cover") +
  theme_bw() +  
  stat_poly_line() +
 stat_poly_eq(use_label(c("eq", "R2")))
print(belttrans_sppr_totalsppr_regression)
  
```


# Join transect and belt sppr.
```{r species richness above ground, include=TRUE}

# created ab_diversity_2 in EXCEL

ab_diversity_2 <- read.csv("processed/ab_diversity_2.csv", header = TRUE, na.strings=c("","NA")) # combined point-intercept transect and belt transect species diversity
nativecover <- read.csv("processed/abcover_nativecover_nngrasscover.csv") # native percent cover from ch1_standdeg_canopy.Rmd
nncover <- read.csv("processed/abcover_nonnativecover_nngrasscover.csv") # non-native percent cover from ch1_standdeg_canopy.Rmd

ab_diversity_2 <- ab_diversity_2 %>% 
  mutate(site = case_when(
    site == "Intact_1" ~ "Intact1",
      site == "Intact_2" ~ "Intact2",
      site == "Intact_3" ~ "Intact3",
    TRUE ~ as.character(site)  # Keep other observations unchanged
  )) %>% 
  mutate(standtype = fct_relevel(standtype, 
            "intact", "matrix", "degraded")) %>% 
   mutate(site = fct_relevel(site, 
            "Intact1", "Intact2", "Intact3", "ENH1",    "ENH2", "DEG1",    "DEG2")) %>% 
   na.omit() 

view(ab_diversity_2)
glimpse(ab_diversity_2)
# Rows: 2,018
# Columns: 9

nativecover <- nativecover %>% 
  dplyr::select(!c(X, standtype.x, standtype.y ))
nncover <- nncover %>% 
  dplyr::select(c(site_rep, percentcover_nonnative))

ab_diversity_2_native <- ab_diversity_2 %>% # join above ground species diversity with native transect percent cover
  left_join(nativecover, by = "site_rep")
ab_diversity_2_native_nn <- ab_diversity_2_native %>% # join above ground species diversity with non-native transect percent cover
  left_join(nncover, by = "site_rep")

view(ab_diversity_2_native_nn)
glimpse(ab_diversity_2_native_nn)
# Rows: 2,018
# Columns: 20

```

### (1) Organize the percent cover data by converting the dataframe into a tibble.
```{r ab percent cover by status, include = TRUE}

str(abcover_full_raw)

head(abcover_full_raw)
tail(abcover_full_raw)

# convert dataframe to tibble
abcover_full_raw <- as_tibble(abcover_full_raw) 

class(abcover_full_raw) # now a tibble
```

### (2) Organize the percent cover data by reordering the classes: Intact, Matrix, Degraded.
```{r ab percent cover reorder classes to Intact Matrix Degraded}
abcover_full_v <- abcover_full_raw %>%
  mutate(standtype = fct_relevel(standtype, 
            "intact", "matrix", "degraded")) %>% 
  na.omit()
unique(abcover_full_v$standtype)

# renaming observations
abcover_full_v$standtype <- factor(abcover_full_v$standtype, levels=c("intact", "matrix", "degraded"), labels=c("Intact", "Matrix", "Degraded"))
unique(abcover_full_v$standtype)

```

### (3) Create new columns for SITE_REP, SITE_REP_DISTANCE, and STAND_STATUS for the percent cover data. 
```{r ab percent cover create site_rep column and site_rep_dist column}

# combine site and rep to one column
abcover_full_v$site_rep <- paste(abcover_full_v$site, abcover_full_v$rep, sep="_") 

# combine site_rep and distance to one column
abcover_full_v$site_rep_dist <- paste(abcover_full_v$site_rep, abcover_full_v$distance, sep="_") 

# combine standtype and native/nonnative status to one column
abcover_full_v$stand_status <- paste(abcover_full_v$standtype, abcover_full_v$status, sep="_") 

# combine site_rep and species to one column
abcover_full_v$site_rep_species <- paste(abcover_full_v$site_rep, abcover_full_v$species, sep="_") 

head(abcover_full_v) # look at new columns
glimpse( abcover_full_v)  
unique(abcover_full_v$site)

```

### (4) Omit "ground" as a status since I'm interested in only Native and Non-native species.
```{r ab percent cover filter for native and non-native only}
abcover_full_v_omitground <- filter(abcover_full_v, status !="ground")
head(abcover_full_v_omitground)
sort(unique(abcover_full_v_omitground$site))
# "DEG1" "DEG2" "ENH1" "ENH2" "INT1" "INT2" "INT3"
glimpse(abcover_full_v_omitground)
```

### (5) Identify the unique species at each site_rep_dist to get rid of duplicate species recordings. (e.g., Present at one distance = 1/41 to get percent cover) 
```{r ab percent cover unby species, inclue = TRUE}

abcover_byspecies <- abcover_full_v_omitground %>% 
  group_by(standtype, site, site_rep, species) %>% 
  filter(site_rep != "Intact4_4") %>% #omit Intact 4 from data
  filter(!is.na(species)) %>% # omit blank cells in species 
  dplyr::summarize(
    count_per_dist = n_distinct(site_rep_dist), 
    percentcover = n_distinct(site_rep_dist)/41) %>% 
   dplyr::select(!c(count_per_dist))
  

head(abcover_byspecies) # prints to new tab


unique(abcover_byspecies$standtype)
unique(abcover_byspecies$site)
unique(abcover_byspecies$site_rep)
unique(abcover_byspecies$species)

```

### (6) Prepare data for an NMDS (percent cover x species x standtype)
```{r transpose percent cover data by species, include=TRUE}

# use this for GROUPS: abcover_byspecies_nmdsprep

abcover_byspecies_nmdsprep <- abcover_byspecies %>% 
  tidyr::spread(species, percentcover) %>% # transpose data so columns are species
  replace(is.na(.), 0) # replace NA with zeros

#view(abcover_byspecies_nmdsprep)
# 15 sites (rows) by 48 columns
str(abcover_byspecies_nmdsprep)


abcover_byspecies_nmdsprep_num <- abcover_byspecies_nmdsprep %>% 
  # use the ungroup() function to take out grouping
  ungroup() %>% 
  dplyr::select(!c(standtype, site_rep, site)) # standtype, site_rep
# 15 sites (rows) by 33 species (columns)
# DEG 1-6, ENH 1-6, INT 1-3 = 15 transects
         
view(abcover_byspecies_nmdsprep_num) # only numbers
write.csv(abcover_byspecies_nmdsprep_num, "processed/nmds_aboveground.csv")
```

### (7) Percent Cover
```{r above ground percent cover, include=TRUE}

above_percentcover <- abcover_full_v_omitground %>% 
  group_by(standtype, site_rep, species, lifeform, status, family, lifecycle) %>% 
  filter(site_rep != "Intact4_4") %>% #omit Intact 4 from data
  filter(!is.na(species)) %>% # omit blank cells in species 
  dplyr::summarize(count_per_dist = n_distinct(site_rep_dist), percentcover = n_distinct(site_rep_dist)/41) %>% 
   dplyr::select(!c(count_per_dist))
  

view(above_percentcover) # prints to new tab


unique(abcover_byspecies$standtype)
unique(abcover_byspecies$site)
unique(abcover_byspecies$site_rep)
sort(unique(abcover_byspecies$species))

```

# Above ground species richness

### load 
```{r species richness by stand type}

ab_diversity_2_native_nn %>% 
  filter(status =="native") %>% 
distinct(species) %>%
  pull(species)
#  [1] "Acmispon glaber"            "Acmispon maritimus"        
#  [3] "Apiastrum angustifolium"    "Artemisia californica"     
#  [5] "Baccharis pilularis"        "Calochortus clavatus"      
#  [7] "Cenothus oliganthus"        "Claytonia perfoliata"      
#  [9] "Dipterostemon capitatus"    "Elymus condensatus"        
# [11] "Eriodictyon crassifolium"   "Eriogonum fasciculatum"    
# [13] "Galium porrigens"           "Heteromeles arbutifolia"   
# [15] "Lupinus spp."               "Malacothamnus fasciculatus"
# [17] "Marah macrocarpa"           "Pellaea andromedifolia"    
# [19] "Rhamnus ilicifolia"         "Rhus ovata"                
# [21] "Salvia apiana"              "Salvia leucophylla"        
# [23] "Sambucus mexicana"          "Solanum xanti"             
# [25] "Stephanomeria virgata"      "Stipa lepida"  

ab_diversity_2_native_nn %>% 
  filter(status =="native") %>% 
distinct(family) %>%
  pull(family)
#  [1] "Fabaceae"             "Apiaceae"             "Asteraceae"          
#  [4] "Liliaceae"               "Montiaceae"          
#  [6] "Themidaceae"          "Poaceae"              "Namaceae"            
#  [9] "Rubiaceae"            "Rosaceae"            
# [11] "Malvaceae"            "Cucurbitaceae"        "Pteridaceae"         
# [14] "Rhamnaceae"           "Anacardiaceae"        "Lamiaceae"           
#  [17] "Solanacaeae" 

ab_diversity_2_native_nn %>% 
  filter(status =="nonnative") %>% 
distinct(species) %>%
  pull(species)
# [1] "Avena barbata"               "Bromus diandrus"            
#  [3] "Bromus hordeaceus"           "Bromus madritensis"         
#  [5] "Bromus rubens"               "Carduus pycnocephalus"      
#  [7] "Centaurea melitensis"        "Erodium spp."               
#  [9] "Hirschfeldia incana"         "Pseudognaphalium luteoalbum"

ab_diversity_2_native_nn %>% 
  filter(status =="nonnative") %>% 
distinct(family) %>%
  pull(family
    )
# [1] "Poaceae"      "Asteraceae"   "Geraniaceae"  "Brassicaceae"

ab_diversity_2_native_nn <- ab_diversity_2_native_nn %>% 
  dplyr::filter(species != "moss") %>% 
   dplyr::filter(species != "nonnative_thatch") %>% 
dplyr::filter(species != "native_litter") 

# combine site and site_rep_species to one column
ab_diversity_2_native_nn$site_rep_species <- paste(ab_diversity_2_native_nn$site_rep, ab_diversity_2_native_nn$species, sep="_") 

ab_diversity_2_native_nn %>% 
  distinct(species) %>%
  pull(species)


ab_div_bysiterep <- ab_diversity_2_native_nn %>% # number of unique species per transect
  group_by(standtype, site_rep) %>%  
  dplyr::summarize(num_species = length(unique(species)))
print(ab_div_bysiterep) 
```


```{r check - number of unique species per transect, include=TRUE}

ab_diversity_2_native_nn %>% 
   dplyr::filter(site_rep == "DEG1_1") %>% 
  distinct(species) %>%
  pull(species)
  
```

### normal
```{r}
shapiro.test(ab_div_bysiterep$num_species)
# data:  ab_div_bysiterep$num_species
# W = 0.96712, p-value = 0.742 <-- normally distrubuted
```

### figure regressions by Kit
```{r scatterplot species diversity by nn grass cover, include=TRUE} 

# KIT's code: this section

ab_diversity_2_native_nn_plot <- ab_diversity_2_native_nn %>%
  group_by(site_rep) %>%
  summarize(richness = n_distinct(species))

hold <- rep(c("DEG1_1", "DEG1_2", "DEG1_3", "DEG2_4", "DEG2_5", "DEG2_6", "ENH1_1", "ENH1_2", "ENH1_3", "ENH2_4", "ENH2_5", "ENH2_6", "Intact_1_1", "Intact_2_2", "Intact_3_3"), times = 3)

hold5 <- rep(c("DEG", "DEG", "DEG", "DEG", "DEG", "DEG", "ENH", "ENH", "ENH", "ENH", "ENH", "ENH", "Intact", "Intact", "Intact"), times = 3)

hold2 <- rep(c("total", "native", "nn"), each = 15)

hold3 <- rep(c(90,98,98,98,90,100,83,83,88,46,66,76,12,7,24), times = 3)

hold4 <- c(10,6,7,7,7,9,8,8,10,14,10,12,6,12,9,5,2,2,3,3,5,3,3,5,10,6,6,5,10,6,5,4,5,4,4,4,5,5,5,4,3,6,1,1,3)

abcover_diversity_plot <- data.frame(site_rep = hold, standtype = hold5, richness = hold4, nngrass = hold3, type = hold2)
  
# Stephanie isn't sure what this code is about... ?
# ab_diversity_2_native_nn_2 <- ab_diversity_2_native_nn %>%
#   left_join(ab_diversity_2_native_nn_plot, by = "site_rep")


ab_sppr_nngrass <- ggplot(data = abcover_diversity_plot, 
       aes(x = nngrass, y = richness, shape = type, color = type)) +
  geom_point(size = 3, alpha = 0.8) +
  scale_shape_manual(values = c(16, 17, 18)) +
  ylim(0, 20)+
  xlim(0,100) +
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", color = "black") +  # lm, loess, glm
  scale_color_manual(values = c(standcolors),
                     name = "Stand Type",
                     labels = c("intact" = "shrub dominated", 
                                "matrix" = "matrix", 
                                "degraded" = "grass dominated")) + #rename standtypes labels
  labs(x = "Non-native grass cover (%)", 
       y = "Species richness (#)", 
       title = "Species richness x non-native grass cover") +
  theme_bw() +  
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2")))

print(ab_sppr_nngrass)
```
### stats
```{r species richness df from An code}
sppr <- specnumber(abcover_byspecies_nmdsprep_num)
sppr
# 11 15  7 15 13 16 11 13 13 10  8  9 12 10 12

abcover_byspecies_nmdsprep <- abcover_byspecies_nmdsprep %>% 
  mutate(standtype_ID = 1)

abcover_byspecies_nmdsprep$standtype_ID <- 1:15

sppr_df <- sppr %>% 
  enframe() %>% 
  full_join(abcover_byspecies_nmdsprep, by = c("name" = "standtype_ID")) %>% 
    rename(spp_rich = value)
view(sppr_df)
```

```{r stats spp richness }

shapiro.test(sppr_df$spp_rich)
# data:  sppr_df$spp_rich
# W = 0.97374, p-value = 0.9091 <-- normally distributed

summary(aov(spp_rich ~ standtype, data = sppr_df)) # by stand type

#             Df Sum Sq Mean Sq F value Pr(>F)  
# standtype    2  35.00  17.500   3.481 0.0643 .
# Residuals   12  60.33   5.028  

summary(aov(spp_rich ~ site, data = sppr_df)) # by hillslope (aka. site)
#             Df Sum Sq Mean Sq F value  Pr(>F)   
# site         6  83.33   13.89   9.259 0.00305 **
# Residuals    8  12.00    1.50 

TukeyHSD(aov(spp_rich ~ site, data = sppr_df))
histogram(sppr_df$spp_rich)

boxplot(spp_rich~site,data=sppr_df, main="Species richness by Hillslope",
   xlab="Hillslope", ylab="Species Richness (N)")
```

```{r species richness figure from An code}
#pal <- c("lightsalmon1", "gold1", "palegreen4")

plot_sppr <- ggplot(sppr_df, aes(x = site, y = spp_rich, fill = site)) +
  geom_boxplot() +
  #scale_fill_manual(spp_rich = pal) +
#  scale_x_discrete(labels = c("Intact \n (n = 9)", "Matrix \n (n = 6)", "Degraded \n (n = 6)")) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = "Hillslope",
       y = "Number of species per hillslope",
       title = "Above ground species richness")
plot_sppr

```

# Shannon Diversity

```{r shannon diversity}
# this doesn't make sense for % cover, but I could run it on shrub density. I could also run it on the seedbank data. 
shannondiv <- diversity(abcover_byspecies_nmdsprep_num)
head(shannondiv)

sppdiv_aov <- aov(shannondiv ~ standtype, data = abcover_byspecies_nmdsprep)
summary(sppdiv_aov)

#     Df Sum Sq Mean Sq F value Pr(>F)  
#standtype    2 0.2899 0.14493   3.924 0.0489 *
#Residuals   12 0.4433 0.03694        

```

# NMDS 
### load
```{r nmds by species percent cover}

# the AUTOTRANSFORM = F skips the automatic data transformations and uses the input data as is
# - when the raw data contains the information you need (e.g., abundance)
# - when you have a small number of samples or small number of variables
# - when you want to preseve the original scale of the data (e.g., interested in the absolute differences betwen two sets of measurements)

abcover_byspecies_NMDS <- metaMDS(abcover_byspecies_nmdsprep_num,
                                  autotransform = F, # or T
                                  distance = "bray",   # bray or raup
                  k = 2, #dimentions
                 try = 100, #runs
                 trymax = 100) # this is more appropriate for my data


# uses the RAUP-CRICK dissimilarity index to calculate pairwise dissimilarities
# - adjusts for spp richness
# - downweights importance of rare spp 
# - more useful when interested in community STRUCTURE rather than individual species
# - takes into account differences in species evenness
# abcover_byspecies_NMDS <- metaMDS(abcover_byspecies_nmdsprep_num, distance = "raup") 

```

### stress
```{r stress tests nmds}
stressplot(abcover_byspecies_NMDS)
# if stress tests are really low ==> I may have insufficient data. :/

plot(abcover_byspecies_NMDS)
ordiplot(abcover_byspecies_NMDS, type = "p")
ordiplot(abcover_byspecies_NMDS, type = "t")
ab_nmds.fort <- fortify(abcover_byspecies_NMDS)
```

```{r scores nmds}

scores(abcover_byspecies_NMDS)

# the below code is from seedbank nmds

#plot_df <- scores(abcover_byspecies_NMDS, display = "sites") %>% # from An's code
#  as.data.frame() %>% 
#  rownames_to_column("stand_core") %>% 
#  full_join(seedbank_bysitecore, by = "stand_core")

nmds_above_df <- data.frame(abcover_byspecies_NMDS$points, abcover_byspecies_nmdsprep) # code from chatgpt

```

### figure 
```{r ggplot above ground}
head(nmds_above_df)

nmds_above_df_plotcolor <- ggplot(nmds_above_df, aes(x = MDS1, y = MDS2, color = standtype, shape = standtype)) +
  geom_point(size = 3, alpha = 0.8) +
  scale_color_manual(values = c("darkgreen", "darkmagenta", "darkgoldenrod1")) + # in color
  # scale_color_manual(values = c("gray0", "gray30", "gray60")) + # in bw
  theme_bw() +
  theme(legend.position = c(0.1, 0.85)) + # x, y left corner: 0,0
  stat_ellipse(level = .95, linetype = 2, linewidth = 1) +
  labs(title = "NMDS: by percent cover", color = "Stand type", shape = "Stand type")
nmds_above_df_plotcolor

```


### p values
```{r nmds color and vectors with envfit function}
# code from An Bui: https://rpubs.com/an-bui/vegan-cheat-sheet 

abcover_byspecies_nmdsprep <- abcover_byspecies_nmdsprep %>% 
  mutate(standtype_ID = 1)
abcover_byspecies_nmdsprep$standtype_ID <- 1:15

plot_above_df <- scores(abcover_byspecies_NMDS, display = "sites") %>% 
  as.data.frame() %>% 
  rownames_to_column("standtype_ID") %>% 
  mutate_at(c('standtype_ID'), as.numeric) %>% 
  full_join(abcover_byspecies_nmdsprep, by = "standtype_ID") %>% 
  dplyr::select(!c(standtype_ID)) %>% 
  mutate(site = case_when(
    site == "INT1" ~ "SD1",
      site == "INT2" ~ "SD2",
      site == "INT3" ~ "SD3",
    site == "ENH1" ~ "MX1",
      site == "ENH2" ~ "MX2",
      site == "DEG1" ~ "GD1",
      site == "DEG2" ~ "GD2",
    TRUE ~ as.character(site)  # Keep other observations unchanged
  )) %>% 
  rename(Hillslope = site) %>% # old name = new name
  mutate(Hillslope = fct_relevel(Hillslope, 
            "SD1", "SD2", "SD3", "MX1", "MX2", "GD1", "GD2"))
  
view(plot_above_df)

# envfit() takes the output of metaMDS() and the species matrix you created
fit_above <- envfit(abcover_byspecies_NMDS, abcover_byspecies_nmdsprep, perm = 999) # pvalues for spp and stand, site, stand_core

# extract p-values for each species
fit_above_pvals <- fit_above$vectors$pvals %>% 
  as.data.frame() %>% 
  rownames_to_column("species") %>% 
  dplyr::rename("pvals" = ".") %>% 
  filter(!c(species == "standtype_ID"))
view(fit_above_pvals)

# extract coordinates for species, only keep species with p-val = 0.05
fit_above_spp <- fit_above %>% 
  scores(., display = "vectors") %>% 
  as.data.frame() %>% 
  rownames_to_column("species") %>% 
  full_join(., fit_above_pvals, by = "species") %>% 
  filter(pvals <= 0.005) %>%  # <-- p-value = 0.005
  filter(!species == 'standtype_ID')
length(fit_above_spp$species) #  <- limited number for visual clarity
#7

#write.csv(fit_above_spp, "~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022b/processed/percentcover_nmds_pvalues.csv")
# P = 1 --> n = 45 species
```

```{r}
# new plot - stand type
nmdsabove_plot_new <- ggplot(plot_above_df, aes(x = NMDS1, y = NMDS2)) +
  #coord_fixed() +
  geom_point(aes(color = standtype), size = 3, alpha = 0.8) +
  stat_ellipse(aes(color = standtype)) +
  scale_color_manual(values = c("darkgreen", "darkmagenta", "darkgoldenrod1"),
                     name = "Stand Type",
                     labels = c("Intact" = "Shrub dominated", 
                                "Matrix" = "Matrix", 
                                "Degraded" = "Grass dominated")) + #rename standtypes labels
  geom_segment(data = fit_above_spp, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")),
               col = "black") +
  geom_text(data = fit_above_spp, aes(label = species),  
            position = position_jitter(width = 0.1, height = 0.2)) 


nmds_above_print <- nmdsabove_plot_new + theme_bw() + theme(legend.position = c(0.125, 0.825))

ggsave("Ch1_Degradation_figures/nmds_above_print.png", nmds_above_print, width = 8, height = 6)

# nmds - by site_rep (aka. hillslope)
nmdsabove_plot_site_rep <- ggplot(plot_above_df, 
                                  aes(x = NMDS1, y = NMDS2)) +
  coord_fixed() +
  #xlim(-1.5,1.5) +
 # ylim(-1.5,1.5) +
  geom_point(aes(color = Hillslope, shape = Hillslope), size = 3, alpha = 0.8) +
  #stat_ellipse(aes(color = site_rep)) +
  scale_color_manual(values = c( "lightblue", "cornflowerblue", "blue","green", "darkgreen", "orange", "brown")) + #"gray0",
   scale_shape_manual(values = c(17, 17, 17, 15, 15, 16,16)) + 
  # labs(name = "Hillslope",
  #                    labels = c("DEG1" = "Shrub dominated", 
  #                               "DEG2" = "Matrix", 
  #                               "ENH1" = "Grass dominated", 
  #                               "ENH2" = "Shrub dominated", 
  #                               "INT1" = "Matrix", 
  #                               "INT2" = "Grass dominated", 
  #                               "INT3" = "SH3")) + #rename standtypes labels
  geom_segment(data = fit_above_spp, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")),
               col = "black") +
  geom_text(data = fit_above_spp, aes(label = species),  
            position = position_jitter(width = 0.0, height = 0.0)) +
  theme_bw()
nmdsabove_plot_site_rep
# nmdsabove_plot_site_rep
# nmdsabove_plot_site_rep
# nmdsabove_plot_site_rep
# nmdsabove_plot_site_rep
# nmdsabove_plot_site_rep
# nmdsabove_plot_site_rep
# nmdsabove_plot_site_rep
# nmdsabove_plot_site_rep
ggsave("Ch1_Degradation_figures/nmds_above_vectors.png", nmdsabove_plot_site_rep, width = 8, height = 6)
```





```{r byplot - ggplot vectors and measurments, include = FALSE}

#this makes a dataframe that gives a lot of control graphically 
abcover_byspecies_NMDS_fort <- fortify(abcover_byspecies_NMDS)

# vectors AND measurments
abcover_byspecies_NMDS_fortplot <- ggplot() + 
  geom_point(data = subset(abcover_byspecies_NMDS_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2, group = ""),
             color = 'black',
             size = 4,
             alpha = 1) +
  geom_segment(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               linewidth = 0.8) +
  geom_text(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1), 
            position = position_jitter(width = 0.2, height = 0.2)) +
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", linewidth = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", linewidth = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
abcover_byspecies_NMDS_fortplot

#only vectors shown
abcover_byspecies_NMDS_fortplot1 <- ggplot() + 
  geom_point(data = subset(abcover_byspecies_NMDS_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0) +
  geom_segment(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               linewidth = 0.8) +
  geom_text(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1),
            position = position_jitter(width = 0.9, height = 0.9)) +
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", linewidth = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
abcover_byspecies_NMDS_fortplot1

#only measurements shown
abcover_byspecies_NMDS_fortplot2 <- ggplot() + 
  geom_point(data = subset(abcover_byspecies_NMDS_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0.5) +
  geom_segment(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               size = 0.8,
               alpha = 0) +
  geom_text(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1),
            position = position_jitter(width = 0.2, height = 0.2),
            alpha = 0)+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", size = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
abcover_byspecies_NMDS_fortplot2

#interactive plots that I like
ggplotly(abcover_byspecies_NMDS_fortplot)

#show both plots 
hi_above <- ggarrange(nmds_above_df_plotcolor, abcover_byspecies_NMDS_fortplot, ncol = 1)
hi_above

```



NMDS
NMDS
NMDS by Plots! to determine which species are driving transect disimiliarity. 

(5) Identify the unique species at each site_rep_dist to get rid of duplicate species recordings. (e.g., Present at one distance = 1/41 to get percent cover) 
```{r ab percent cover unby species, inclue = TRUE}

abcover_byplot_groups <- abcover_full_v_omitground %>% 
  select(c(standtype, site, site_rep, species))
head(abcover_byplot_groups)

abcover_byplot_prep <- abcover_full_v_omitground %>% 
  group_by(standtype, site, site_rep, species) %>% 
  filter(site_rep != "Intact4_4") %>% #omit Intact 4 from data
  filter(!is.na(species)) %>% # omit blank cells in species 
  dplyr::summarize(count_per_dist = n_distinct(site_rep_dist), percentcover = n_distinct(site_rep_dist)/41) 
  
view(abcover_byplot_prep) # prints to new tab

abcover_byplot <- abcover_byplot_prep %>% 
  select(c(site_rep, species, percentcover))

write.csv(abcover_byplot, "~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022b/processed/abcover_byplot.csv")

abcover_byplot_load <- read.csv("~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022b/processed/abcover_byplot_onlysite_repsppcover.csv", header = TRUE, na.strings=c("","NA"))
```


(6) Prepare data for an NMDS (percent cover x species x standtype)
```{r transpose percent cover data by species, include=TRUE}

# use this for GROUPS: abcover_byplot_nmdsprep

head(abcover_byplot_load)

abcover_byplot_nmdsprep <- abcover_byplot_load %>% 
  tidyr::spread(site_rep, percentcover) %>% # transpose data so columns are species
  replace(is.na(.), 0) # replace NA with zeros
view(abcover_byplot_nmdsprep)
dim(abcover_byplot_nmdsprep)

abcover_byplot_nmdsprep_num <- abcover_byplot_nmdsprep %>% 
  # use the ungroup() function to take out grouping
  ungroup() %>% 
  select(!c(species))
# 33 species (rows) by 15 sites (columns)
# DEG 1-6, ENH 1-6, INT 1-3 = 15 transects
dim(abcover_byplot_nmdsprep_num)
```

```{r write nmds to csv}
view(abcover_byplot_nmdsprep_num) # only numbers
write.csv(abcover_byplot_nmdsprep_num, "~/Dropbox/GRADSCHOOL/Dissertation/R_dissertation/chaparraldegradation_2022b/processed/nmds_aboveground.csv")
```

(7) NMDS 
```{r stress tests}
# USE THIS ONE!

# str(seedbank_bysitecore_nmdsprep_num)
# stand_core_depth -->  tibble [89 × 52]

# stress test, will run 20 times
seedbank_byplot_nmds <- metaMDS(abcover_byplot_nmdsprep_num, distance = "raup")

# stress test, will run 20 times
seedbank_byplot_nmds <- metaMDS(abcover_byplot_nmdsprep_num, autotransform = F)
```

```{r scores nmds}

scores(seedbank_byplot_nmds)

```


```{r stress test}

stressplot(seedbank_byplot_nmds)
# if stress tests are really low ==> I may have insufficient data. :/

#plot_df <- scores(seedbank_bysitecore_nmds, display = "sites") %>% # from An's code
#  as.data.frame() %>% 
#  rownames_to_column("stand_core") %>% 
#  full_join(seedbank_bysitecore, by = "stand_core")

nmds_df <- data.frame(seedbank_byplot_nmds$points, # unnamed column is species n = 33
                      abcover_byplot_nmdsprep) # from chatgpt
view(nmds_df)
```

```{r ggplot bw}
ggplot(nmds_df, aes(x = MDS1, y = MDS2, color = stand, shape = stand)) +
  geom_point(size = 3, alpha = 0.8) +
  scale_color_manual(values = c("gray0", "gray30", "gray60")) +
  theme_bw() +
  stat_ellipse(linetype = 2, size = 1) +
  labs(title = "NMDS: by seed bank germination counts", color = "Stand type", shape = "Stand type")
```

```{r with vectors}

#this makes a dataframe that gives a lot of control graphically 
seedbank_byplot_nmds_fort <- fortify(seedbank_byplot_nmds)

# vectors AND measurments
seedbank_byplot_nmdsfortplot <- ggplot() + 
  geom_point(data = subset(seedbank_byplot_nmds_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0.5) +
  geom_segment(data = subset(seedbank_byplot_nmds_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               linewidth = 0.8) +
  geom_text(data = subset(seedbank_byplot_nmds_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1))+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", linewidth = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", linewidth = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

#only vectors shown
seedbank_byplot_nmds_fortplot1 <- ggplot() + 
  geom_point(data = subset(seedbank_byplot_nmds_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0) +
  geom_segment(data = subset(seedbank_byplot_nmds_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               linewidth = 0.8) +
  geom_text(data = subset(seedbank_byplot_nmds_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1))+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", linewidth = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
seedbank_byplot_nmds_fortplot1
```
```{r plot measurements only }
#only measurements shown
seedbank_byplot_nmds_fortplot2 <- ggplot() + 
  geom_point(data = subset(seedbank_byplot_nmds_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0.5) +
  geom_segment(data = subset(seedbank_byplot_nmds_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               size = 0.8,
               alpha = 0) +
  geom_text(data = subset(seedbank_byplot_nmds_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1),
            alpha = 0)+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", size = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

#interactive plots that I like
ggplotly(seedbank_bysitecore_nmds_fortplot)

#show both plots 
hi <- ggarrange(seedbank_bysitecore_nmds_fortplot1, seedbank_bysitecore_nmds_fortplot2, ncol = 1)
hi

```


```{r}

ggplot(nmds_df, aes(x = MDS1, y = MDS2)) +
  geom_point()+
  theme_bw() +
  stat_ellipse(linetype = 2, size = 1) +
  labs(title = "NMDS: by seed bank germination counts", color = "Stand type", shape = "Stand type")

#
ggplot(nmds_df, aes(x = MDS1, y = MDS2, color = stand, shape = stand)) +
  geom_point(size = 3, alpha = 0.8) +
  scale_color_manual(values = c("gray0", "gray30", "gray60")) +
  theme_bw() +
  stat_ellipse(linetype = 2, size = 1) +
  labs(title = "NMDS: by seed bank germination counts", color = "Stand type", shape = "Stand type")
```

```{r ggplot color}
ggplot(nmds_df, aes(x = MDS1, y = MDS2, color = stand, shape = stand)) +
  geom_point(size = 3, alpha = 0.8) +
  scale_color_manual(values = c("darkgreen", "darkmagenta", "darkgoldenrod1")) +
  theme_bw() +
  #stat_ellipse(level = .95, linetype = 2, size = 1) +
  stat_ellipse(linetype = 2, size = 1) +
  labs(title = "NMDS: by seed bank germination counts", color = "Stand type", shape = "Stand type")


```



```{r}
# useful code on how to color code NMDS from YouTube: https://www.youtube.com/watch?v=h7OrVmT7Ja8&t=446s
# and how to group NMDS values: https://www.youtube.com/watch?v=Y0GI34S-ZMI

scores(abcover_byspecies_NMDS) # structure of the file, look at the points to get the matrix of numbers and columns (2 columns)
abcover_byspecies_NMDS$points # to get the points/values of MDS1 and MDS2
scores(abcover_byspecies_NMDS) # scores of your variables
plot(abcover_byspecies_NMDS)

abcover_byspecies_NMDS_centroid <- scores(abcover_byspecies_NMDS$species) %>% 
  as_tibble(rownames="species") %>% 
  inner_join(., abcover_byspecies_nmdsprep, by="species") %>% 
  group_by(standtype) %>% 
 summarize(axis1 = mean(MDS1),
            axis2 = mean(MDS2))

scores(abcover_byspecies_NMDS$species) %>% 
  as_tibble(rownames="species") %>% 
  inner_join(., abcover_bysite_nmdsprep, by="species") %>% 
  ggplot(aes(x=MDS1, y=MDS2, color=standtype)) +
  geom_point() +
  geom_point(data=abcover_byspecies_NMDS_centroid, aes(x=axis1, y=axis2, color=standtype),
             shape = 15, size = 5)

# How can I make interactive data points with call outs when I hover over them?
```

```{r FROM KIT - nmds by species percent cover using code from KIT}
# USE THIS ONE!

# stress test, will run 20 times
abcover_byspecies_NMDS <- metaMDS(abcover_byspecies_nmdsprep_num, 
                                  autotransform = F)
# stress tests are really low ==> I may have insufficient data. :/

stressplot(seedbank_byplot_nmds)
plot(seedbank_byplot_nmds)

#basic normal plot
ordiplot(abcover_byspecies_NMDS) # without point labels
ordiplot(abcover_byspecies_NMDS, type = "t") # with point labels

#ggvegan to make a quick nice plot
autoplot(abcover_byspecies_NMDS) # NMDS1 x NMDS2, Key = "Score": sites, species



#this makes a dataframe that gives a lot of control graphically 
abcover_byspecies_NMDS_fort <- fortify(abcover_byspecies_NMDS)

abcover_byspecies_NMDS_fortplot <- ggplot() + 
  geom_point(data = subset(abcover_byspecies_NMDS_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0.5) +
  geom_segment(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               linewidth = 0.8) +
  geom_text(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1))+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", linewidth = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", linewidth = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

#only vectors shown
abcover_byspecies_NMDS_fortplot1 <- ggplot() + 
  geom_point(data = subset(abcover_byspecies_NMDS_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0) +
  geom_segment(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               linewidth = 0.8) +
  geom_text(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1))+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", linewidth = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

#only measurements shown
abcover_byspecies_NMDS_fortplot2 <- ggplot() + 
  geom_point(data = subset(abcover_byspecies_NMDS_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0.5) +
  geom_segment(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               size = 0.8,
               alpha = 0) +
  geom_text(data = subset(abcover_byspecies_NMDS_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1),
            alpha = 0)+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", size = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

#interactive plots that I like
ggplotly(abcover_byspecies_NMDS_fortplot)

#show both plots 
hi <- ggarrange(abcover_byspecies_NMDS_fortplot1, abcover_byspecies_NMDS_fortplot2, ncol = 1)
hi


ordiplot(abcover_byspecies_NMDS,type="n")
orditorp(abcover_byspecies_NMDS,display="species",col="red",air=0.01)
orditorp(abcover_byspecies_NMDS,display="sites",cex=1.25,air=0.01)

str(abcover_byspecies_nmdsprep_num)
#tibble [125 × 34] (S3: tbl_df/tbl/data.frame)
treat=c(rep("treatment1",50),rep("treatment2",50), rep("treatment3",25)) 
ordiplot(abcover_byspecies_NMDS,type="n")
ordihull(abcover_byspecies_NMDS,groups=treat,draw="polygon",col="grey90",
         label=FALSE)
orditorp(abcover_byspecies_NMDS,display="species",col="red",air=0.01)
orditorp(abcover_byspecies_NMDS,display="sites",col=c(rep("green",50),rep("blue",50),re("orange",25)),
         air=0.01,cex=1.25)


```


(7) NMDS by site (e.g., Intact_1)
```{r transpose site percent cover data}

abcover_bysite_nmdsprep <- abcover_byspecies %>% 
  tidyr::spread(site_rep, percentcover) %>% # transpose data so columns are site_rep
  replace(is.na(.), 0) # replace NA with zeros
view(abcover_bysite_nmdsprep)

abcover_bysite_nmdsprep_num <- abcover_bysite_nmdsprep %>% 
  # use the ungroup() function to take out grouping
  ungroup() %>% 
  select(!c(standtype, site, species, status, lifeform, lifecycle, count_per_dist))
         
view(abcover_bysite_nmdsprep_num) # only numbers
```


```{r FROM KIT - nmds by site percent cover}

# stress test with DISTANCE, runs 20 times
# abcover_bysite_NMDS <- metaMDS(abcover_bysite_nmdsprep_num, distance = "raup") 
#In metaMDS(abcover_bysite_nmdsprep_num, distance = "raup") :
#  stress is (nearly) zero: you may have insufficient data

# stress test with AUTOTRANSFORM, runs 20 times
abcover_bysite_NMDS <- metaMDS(abcover_bysite_nmdsprep_num, autotransform = F) 
#In metaMDS(abcover_bysite_nmdsprep_num, autotransform = F) :
#  stress is (nearly) zero: you may have insufficient data

stressplot(abcover_bysite_NMDS)

plot(abcover_bysite_NMDS) 

#basic normal plot
ordiplot(abcover_bysite_NMDS) # without point labels
ordiplot(abcover_bysite_NMDS, type = "t") # with point labels

#ggvegan to make a quick nice plot
autoplot(abcover_bysite_NMDS) # NMDS1 x NMDS2, Key = "Score": sites, species



#this makes a dataframe that gives a lot of control graphically 
abcover_bysite_NMDS_fort <- fortify(abcover_bysite_NMDS)

abcover_bysite_NMDS_fortplot <- ggplot() + 
  geom_point(data = subset(abcover_bysite_NMDS_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0.5) +
  geom_segment(data = subset(abcover_bysite_NMDS_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               linewidth = 0.8) +
  geom_text(data = subset(abcover_bysite_NMDS_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1))+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", linewidth = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", linewidth = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

#only vectors shown
abcover_bysite_NMDS_fortplot1 <- ggplot() + 
  geom_point(data = subset(abcover_bysite_NMDS_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0) +
  geom_segment(data = subset(abcover_bysite_NMDS_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               linewidth = 0.8) +
  geom_text(data = subset(abcover_bysite_NMDS_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1))+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", linewidth = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

#only measurements shown
abcover_bysite_NMDS_fortplot2 <- ggplot() + 
  geom_point(data = subset(abcover_bysite_NMDS_fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0.5) +
  geom_segment(data = subset(abcover_bysite_NMDS_fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               size = 0.8,
               alpha = 0) +
  geom_text(data = subset(abcover_bysite_NMDS_fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1),
            alpha = 0)+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", size = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

#interactive plots that I like
ggplotly(abcover_bysite_NMDS_fortplot)

#show both plots 
hi_site <- ggarrange(abcover_bysite_NMDS_fortplot1, abcover_bysite_NMDS_fortplot2, ncol = 1)
hi_site


ordiplot(abcover_bysite_NMDS,type="n")
orditorp(abcover_bysite_NMDS,display="species",col="red",air=0.01)
orditorp(abcover_bysite_NMDS,display="sites",cex=1.25,air=0.01)

str()
#tibble [125 × 34] (S3: tbl_df/tbl/data.frame)
treat=c(rep("treatment1",3),rep("treatment2",3), rep("treatment3",3)) 
ordiplot(abcover_bysite_NMDS,type="n")
ordihull(abcover_bysite_NMDS,groups=treat,draw="polygon",col="grey90",
         label=FALSE)
orditorp(abcover_bysite_NMDS,display="species",col="red",air=0.01)
orditorp(abcover_bysite_NMDS,display="sites",col=c(rep("green",6),rep("blue",6),re("orange",3)),
         air=0.01,cex=1.25)




```





```{r set seed }
# *** No convergence -- monoMDS stopping criteria:
#    15: no. of iterations >= maxit
#     1: stress < smin
#     4: scale factor of the gradient < sfgrmin

# set.seed(100)
abcover_bysite_NMDS <- metaMDS(abcover_bysite_nmdsprep_num, distance = "raup") # Runs with stress and a number

str(abcover_bysite_NMDS) # structure of the file, look at the points to get the matrix of numbers and columns (2 columns)
abcover_bysite_NMDS$points # to get the points/values of MDS1 and MDS2
scores(abcover_bysite_NMDS) # scores of your variables
plot(abcover_bysite_NMDS)
```




```{r sandbox nmds}

plot(abcover_bysite_NMDS, type = "n") #displays empty ordination space

points(abcover_bysite_NMDS, display = "sites", pch = c(16, 8, 17, 11) [as.numeric(abcover_bysite_nmdsprep$status)], col = c("blue", "orange", "black") [as.numeric(abcover_bysite_nmdsprep$count_per_dist)]) # displays site points where symbols (pch) are different management options and colour (col) are different land uses

# points(abcover_bysite_NMDS, display = "site_rep", 
#       pch = c(16, 8, 17, 11) [as.numeric(abcover_bysite_nmdsprep_num$site_rep)], 
#       col = c("blue", "orange", "black", "green", "pink", "red", "purple", "teal", #"magenta", "cyan", "yellow", "brown", "light blue", "peach", "lime"),
#       [as.numeric(abcover_bysite_nmdsprep_num$Use)]) # displays site points where #symbols (pch) are different management options and colour (col) are different land uses

legend("topright", legend = c(levels(abcover_bysite_nmdsprep_num$site_rep),
                              levels(abcover_bysite_nmdsprep_num$percent)), 
       pch = c(16, 8, 17, 11, 16, 16, 16), 
       col = c("black","black","black","black","blue", "orange", "black"), 
       bty = "n", cex = 1) # displays symbol and colour legend

legend("topleft", "stress = 0.118", bty = "n", cex = 1) # displays legend text of stress value 

nmds <- ordiplot(abcover_bysite_NMDS, type = "none")
ordipointlabel(abcover_bysite_NMDS,display = "sites",add=TRUE)
```




