---
title: "NMDS Test"
author: "Kit Swift"
date: "2023-03-07"
output: html_document
---

```{r setup, include=FALSE}
remotes::install_github("gavinsimpson/ggvegan")
remotes::install_github("jfq3/ggordiplots")

#loading libraries
library(ggvegan)
library(tidyverse)
library(vegan)
library(ggpubr)
library(plotly)
library(ggordiplots)
```

Pulled from: https://jonlefcheck.net/2012/10/24/nmds-tutorial-in-r/

The goal of NMDS is to represent the original position of communities in
multidimensional space as accurately as possible using a reduced number
of dimensions that can be easily plotted and visualized

The NMDS procedure is iterative and takes place over several steps:
(1) Define the original positions of communities in multidimensional
space
(2) Specify the number m of reduced dimensions (typically 2)
(3) Construct an initial configuration of the samples in 2-dimensions
(4) Regress distances in this initial configuration against the observed
(measured) distances
(5) Determine the stress (disagreement between 2-D configuration and
predicted values from regression)
If the 2-D configuration perfectly preserves the original rank
orders, then a plot ofone against the other must be monotonically
increasing. The extent to which the points on the 2-D configuration
differ from this monotonically increasing line determines the
degree of stress (see Shepard plot)
(6) If stress is high, reposition the points in m dimensions in the
direction of decreasing stress, and repeat until stress is below
some threshold

Generally, stress < 0.05 provides an excellent represention in reduced
dimensions, < 0.1 is great, < 0.2 is good, and stress > 0.3 provides a
poor representation

NOTE: The final configuration may differ depending on the initial
configuration (which is often random) and the number of iterations, so
it is advisable to run the NMDS multiple times and compare the
interpretation from the lowest stress solutions

To begin, NMDS requires a distance matrix, or a matrix of
dissimilarities
Raw Euclidean distances are not ideal for this purpose: they are
sensitive to totalabundances, so may treat sites with a similar number
of species as more similar, even though the identities of the species
are different
They are also sensitive to species absences, so may treat sites with
the same number of absent species as more similar

Consequently, ecologists use the Bray-Curtis dissimilarity calculation,
which has many ideal properties:
It is invariant to changes in units
It is unaffected by additions/removals of species that are not
present in two communities
It is unaffected by the addition of a new community
It can recognize differences in total abudnances when relative
abundances are the same


```{r thieftime}

#this chunk is all code that i have taken from other sources to look at, personally very helpful, but i need to make it known this is not my code.


#code from video (https://www.youtube.com/watch?v=OMrtxobDhrM) (nice guy, kinda boring but gave good code)
data(mite)
data(mite.env)

#transforming data (i dont know all the transformations yet, but its easy to find them; https://ordnews.colostate.narkive.com/lMWF502c/1593-log-sqrt-and-other-transformation-with-bray-curtis-dissimilarity)
mite.hel <- decostand(mite, method = "hellinger")
nmds1 <- metaMDS(mite.hel, 
                 autotransform = F)

#basic normal plot
ordiplot(nmds1)
ordiplot(nmds1, type = "t") 

#ggvegan to make a quick nice plot
autoplot(nmds1) 

#this makes a dataframe that gives a lot of control graphically 
fort <- fortify(nmds1)

fortplot <- ggplot() + 
  geom_point(data = subset(fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0.5) +
  geom_segment(data = subset(fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               size = 0.8) +
  geom_text(data = subset(fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1))+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", size = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

#only vectors shown
fortplot1 <- ggplot() + 
  geom_point(data = subset(fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0) +
  geom_segment(data = subset(fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               size = 0.8) +
  geom_text(data = subset(fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1))+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", size = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

#only measurements shown
fortplot2 <- ggplot() + 
  geom_point(data = subset(fort, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2),
             color = 'black',
             alpha = 0.5) +
  geom_segment(data = subset(fort, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               size = 0.8,
               alpha = 0) +
  geom_text(data = subset(fort, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1),
            alpha = 0)+
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", size = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

#interactive plots that I like
ggplotly(fortplot)

#show both plots 
hi <- ggarrange(fortplot1, fortplot2, ncol = 1)
hi




#More code stealing, from https://jonlefcheck.net/2012/10/24/nmds-tutorial-in-r/ (edited to fit the nmds from above, this is graphical code I wanted)
example_NMDS <- nmds1

#Now we can plot the NMDS
plot(example_NMDS)
# It shows us both the communities ("sites", open circles) and species 
# (red crosses), but we  don't know which are which!

# We can use the functions `ordiplot` and `orditorp` to add text to the 
# plot in place of points
ordiplot(example_NMDS,type="n")
orditorp(example_NMDS,display="species",col="red",air=0.01)
orditorp(example_NMDS,display="sites",cex=1.25,air=0.01)

# There are some additional functions that might of interest
# Let's suppose that communities 1-5 had some treatment applied, and 
# communities 6-10 a different treatment
# We can draw convex hulls connecting the vertices of the points made by
# these communities on the plot
# First, let's create a vector of treatment values:
treat=c(rep("Treatment1",35),rep("Treatment2",35))
ordiplot(example_NMDS,type="n")
ordihull(example_NMDS,groups=treat,draw="polygon",col="grey90",
         label=FALSE)
orditorp(example_NMDS,display="species",col="red",air=0.01)
orditorp(example_NMDS,display="sites",col=c(rep("green",35),rep("blue",35)),
         air=0.01,cex=1.25)
# I find this an intuitive way to understand how communities and species 
# cluster based on treatments
# One can also plot ellipses and "spider graphs" using the functions 
# `ordiellipse` and `orderspider` which emphasize the centroid of the 
# communities in each treatment

# Another alternative is to plot a minimum spanning tree (from the 
# function `hclust`), which clusters communities based on their original 
# dissimilarities and projects the dendrogram onto the 2-D plot
ordiplot(example_NMDS,type="n")
orditorp(example_NMDS,display="species",col="red",air=0.01)
orditorp(example_NMDS,display="sites",col=c(rep("green",35),rep("blue",35)),
         air=0.01,cex=1.25)
ordicluster(example_NMDS,hclust(vegdist(mite.hel,"bray"))) 
# Note that clustering is based on Bray-Curtis distances
# This is one method suggested to check the 2-D plot for accuracy

# You could also plot the convex hulls, ellipses, spider plots, etc. colored based on the treatments
# First, create a vector of color values corresponding of the same length as the vector of treatment values
colors=c(rep("red",35),rep("blue",35))
ordiplot(example_NMDS,type="n")
#Plot convex hulls with colors baesd on treatment
for(i in unique(treat)) {
  ordihull(example_NMDS$point[grep(i,treat),],draw="polygon",groups=treat[treat==i],col=colors[grep(i,treat)],label=F) } 
orditorp(example_NMDS,display="species",col="green",air=0.01)
orditorp(example_NMDS,display="sites",col=c(rep("red",35),rep("blue",35)),
         air=0.01,cex=1.25)

# If the treatment is a continuous variable, consider mapping contour 
# lines onto the plot
# For this example, consider the treatments were applied along an 
# elevational gradient
# We can define random elevations for previous example
elevation=runif(70,0.5,1.5)
# And use the function ordisurf to plot contour lines
ordisurf(example_NMDS,elevation,main="",col="forestgreen")
# Finally, we want to display species on plot
orditorp(example_NMDS,display="species",col="grey30",air=0.1,cex=1)

```

Notes for Stephanie: 
Okay so what im learning/have learned is that the actual NMDS itself is easy to code and find, its the interpretation/graph making where the difficulty comes in. My goals in the following code were to play around with the graphs and find ways to visualize them in ways I like. + a small amount of work looking at statistical analysis beyond NMDS because why not :)

```{r kitrun}
#im choosing the dune data cause the dune books are my favorite fantasy series (in a fun nuanced queer way not in a warmongering incel way)
data(package = "vegan")
data(dune)
data(BCI)

#transforming data
dune.hel <- decostand(dune, method = "hellinger")

#preforming the actual nmds, Im having it run 100 times to prevent it reaching local minimums for stress. (stress being the measure of fit)
dune.nmds <- metaMDS(dune.hel, 
                 autotransform = T,
                 distance = "bray",
                 k = 2,
                 try = 100,
                 trymax = 100)

plot(dune.nmds)
ordiplot(dune.nmds,type="n")
orditorp(dune.nmds,display="species",col="red",air=0.01)
orditorp(dune.nmds,display="sites",cex=1.25,air=0.01)

#how do i do this in ggplot (for better control + cleaner graphics)
treat=c(rep("treatment1",10),rep("treatment2",10))
ordiplot(dune.nmds,type="n")
ordihull(dune.nmds,groups=treat,draw="polygon",col="grey90",
         label=FALSE)
orditorp(dune.nmds,display="species",col="red",air=0.01)
orditorp(dune.nmds,display="sites",col=c(rep("green",10),rep("blue",10)),
         air=0.01,cex=1.25)



#making the nmds usable for ggplot
treat=c(rep("treatment1",10),rep("treatment2",10), rep("species", 30))
dune.nmdsf <- fortify(dune.nmds)
dune.nmdsf$treatment <- treat

#found a package names ggordihulls, maybe try later?

#making the required data for the ordihull (this is stolen, changed a lot though to match how i like things)
grp.a <- dune.nmdsf[dune.nmdsf$treatment == "treatment1", ][chull(dune.nmdsf[dune.nmdsf$treatment == 
    "treatment1", c("NMDS1", "NMDS2")]), ]  # hull values for treatment 1
grp.b <- dune.nmdsf[dune.nmdsf$treatment == "treatment2", ][chull(dune.nmdsf[dune.nmdsf$treatment == 
    "treatment2", c("NMDS1", "NMDS2")]), ]  # hull values for treatment 2
hull.data <- rbind(grp.a, grp.b)  #combine grp.a and grp.b


#making the actual plot, split up to see how to do different visualizations
duneplot <- ggplot() + 
  
  #putting all the "site" data on the graph, these are the different observations
  geom_point(data = subset(dune.nmdsf, Score == 'sites'),
             mapping = aes(x = NMDS1, y = NMDS2, color = treatment),
             alpha = 0.5,  #transparency (to hide/show)
             show.legend = F) +
  
  #while this is the species data, or abundance based on different observations. 
  geom_point(data = subset(dune.nmdsf, Score == 'species'), 
             mapping = aes(x = NMDS1, y = NMDS2),
             alpha = 0.5, #transparency (to hide/show)
             color = "blue",
             show.legend = F) +
  
  #this is an ordihull, grouping data based on treatments to help see the gradient
  geom_polygon(data=hull.data,
               aes(x=NMDS1,y=NMDS2,fill=treatment,group=treatment),
               alpha=0.30,
               show.legend = F) +
  
  #these are for an envfit, graphing vectors for environmental variables
  geom_segment(data = subset(dune.nmdsf, Score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type = 'closed'),
               color = 'darkgrey',
               size = 0.8,
               alpha = 0) +
  
  #making it pretty with colors
  scale_color_manual(values = c(
    "treatment1" = "green",
    "treatment2" = "salmon",
    "species" = "blue"
  )) +
  scale_fill_manual(values = c(
    "treatment1" = "green",
    "treatment2" = "salmon",
    "species" = "blue"
  )) +
  
  
  #this labels each species with its location withing the nmds
  geom_text(data = subset(dune.nmdsf, Score == 'species'),
            mapping = aes(label = Label, x = NMDS1*1.1, y = NMDS2*1.1),
            alpha = 0.5) +  
  
  #lines at 0,0 
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", size = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "gray")+
  
  #theme to make it come together better
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

duneplot
ggplotly(duneplot) #gotta figure out how to make species be toggleable
```


```{r ggordihulls}

#these are easier ways to do what I did above but I dont understand the "dune" data to be able to play with them signficantly, need dataset with known context (and also a low use unknown package with like 0 documentation)

gg_envfit(dune.nmds, dune, groups = dune.env$Management)
gg_ordisurf(dune.nmds, dune)

gg <- gg_ordiplot(dune.nmds, 
            groups = dune.env$Use, 
            spiders=TRUE, ellipse=TRUE, plot=TRUE) 
  
#have to make the ggordiplot a plot then add all the goodies I want
gg$plot +
  geom_abline(intercept = 0,slope = 0,linetype = "dashed", size = 0.8,colour = "gray")+
  geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, colour = "black")
  

```










