---
title: "transect_continuous"
author: "Stephanie Ma Lucero"
date: "2024-04-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r code to hang on to, include=FALSE}

# df<- df_load %>% 
#  mutate(site_rep = paste(site, transect, sep = "_")) %>% # create new column
# unite("lifeform_total", 2:3, sep = "_") # consolidate columns

```

```{r libraries, message=FALSE}

library(here) # similar to set working directory
library(tidyverse) # data wrangling - includes ggplot2, dplyr, tidyr, readr, purr, tibble, stringr, forcats
# library(lubridate) # dates and times
library(ggplot2) # data visualization
library(gridExtra)
# library(readr)
library(dplyr)
library(ggthemes)
library(plotly)
library(ggpubr) # customize ggplot2 for publication
#library(broom)
#library(AICcmodavg)
library(vegan) #community ecology package - includes, MASS, cluster, mgcv
library(readxl) # for .xls and .xlsx sheets
#library(janitor)
library(multcompView) #Visualizations of Paired Comparisons, functions: TukeyHSD, dist{stats}, simint, simtest, csimint, csimtest{multcomp}, friedmanmc, kruskalmc{pgirmess}
#library(multcomp) # functions: multcompLetters2, multcompLetters3, multcompLetters4 using aov object
#library(DHARMa) #The ‘DHARMa’ package uses a simulation-based approach to create readily interpretable scaled (quantile) residuals for fitted (generalized) linear mixed models.
#citation("DHARMa")
library(ggpmisc)
library(segmented)
library(strucchange)
library(ggpattern)
library(lme4)
library(lmerTest) # load again or update if you're not getting p-values from environmental models
```

```{r colors, include=TRUE}
devtools::install_github("an-bui/calecopal")

library(calecopal)
# all palettes
names(cal_palettes)

# Stand colors: INTACT, MATRIX, DEGRADED 
cal_palette(name = "chaparral3", n = 3, type = "continuous") 
#deg <- "#D3E3CA"
#mat <- "#92A587"
#int <- "#2F3525"
deg <- "gray60"
mat <- "gray40"
int <- "gray20"
line <- "red"
degshape <- 16
matshape <- 17
intshape <- 18


standcolors <- c(int, mat, deg)
standcolors_line <- c(line, int, mat, deg)
standshape <- c(intshape, matshape, degshape)

# Lifeform colors: shrub, grass, forb
cal_palette(name = "chaparral3", n = 3, type = "continuous") 
shrub <- "green4"
grass <- "yellow2"
forb <- "purple2"
fern <- "black"
lifeformcolors <- c(shrub, grass, forb, fern)

# Status colors: native, exotic 
cal_palette(name = "chaparral1", n = 6, type = "continuous") 
native <- "gray50"
nonnative <- "gray0"
statuscolors <- c(native, nonnative)


# "orange", "lightblue", "limegreen", "deeppink","mediumpurple2", "darkgreen"
```

# load data

Here I'm loading the point-intercept transect data.
Years collected: 2019, 2024
Sites: 1 and 2
Comparing to intact sites: INT1, INT2, INT3
```{r load transect data, include=TRUE}
# percent cover - belt transects, point intercept transects
pointintercept_load <- read.csv("deg_percentcover_byspecies_bylifeform_2024.csv", header = TRUE, na.strings=c("","NA")) %>% 
  #rename(speciescode = species.1 ) %>%
  mutate(standtype = fct_relevel(standtype, 
            "degraded", "matrix", "intact")) %>% 
  dplyr::select(!c(species_code, notes)) # %>% 
    # mutate(site = case_when(site == "Intact_1" ~ "INT1", 
    #                             site == "Intact_2" ~ "INT2",
    #                             site == "Intact_3" ~ "INT3", 
    #       TRUE ~ site))

glimpse(pointintercept_load)
str(pointintercept_load)
```

```{r load schlero leaf data, echo=FALSE, eval=F}

shrub_leaftype <- read_excel("shrub_leaftype.xlsx")
print(shrub_leaftype)
```

```{r}
# convert dataframe to tibble
abcover_full_tib <- as_tibble(pointintercept_load) 
class(abcover_full_tib)

abcover_full_v <- abcover_full_tib %>% 
   mutate(site_rep = paste(site, transect, sep = "_")) %>% # combine site and rep to one column
  mutate(site_rep_dist = paste(site_rep, distance, sep = "_")) %>% # combine site_rep and distance to one column
  mutate(stand_status = paste(standtype, status, sep = "_")) # combine standtype and native/nonnative status to one column
# %>%   mutate(site_rep = fct_relevel
         # (site_rep, 
         #                        "INT1_1", "INT2_2", "INT3_3", "ENH1_1", "ENH1_2",    "ENH1_3"  ,   "ENH2_4"  ,   "ENH2_5"    , "ENH2_6"  , "DEG1_1"   ,  "DEG1_2"  ,   "DEG1_3"     ,"DEG2_4" ,    "DEG2_5"    , "DEG2_6" ))

glimpse(abcover_full_v) # look at new columns
```

```{r}
sort(unique(abcover_full_v$species))
```

```{r}
abcover_full_v %>%  
   # na.omit() %>%  
  group_by(site, hillslope, transect) %>% 
  summarize(data_points = length(unique(distance)))
```

```{r}
# omit "ground" as a status since interested in Native and Non-native species 
abcover_full_v_omitground <- filter(abcover_full_v, status =="native" |status == "nonnative")
print(abcover_full_v_omitground)
```

```{r preprbindleaftype, include=TRUE}

shrubs <- abcover_full_v_omitground %>% 
  filter(lifeform == "shrub") %>% 
  left_join(shrub_leaftype, by = "species") %>% 
  dplyr::select(!c(lifeform)) %>% 
    rename(lifeform = leaftype ) # newname = oldname

herbsgrassesfern <-  abcover_full_v_omitground %>% 
  filter(lifeform =="herb" | lifeform == "grass" | lifeform == "fern") 
  
lifeformleaftype <- rbind(herbsgrassesfern, shrubs) %>%  
  group_by(site_rep_dist) %>%
   mutate(lifeform_status = paste(lifeform, status, sep = "_")) %>% 
  group_by(standtype, site, transect, site_rep, lifeform, status, lifeform_status) %>% 
  dplyr::summarize(
    count_per_transect = n_distinct(site_rep_dist), 
    percentcover = n_distinct(site_rep_dist)/41) %>% # <-- calculate percent cover
  ungroup() #%>% 
  #  mutate(standtype = fct_relevel(standtype,"intact", "matrix", "degraded")) %>% 
  #  mutate(site = fct_relevel(site, "INT1", "INT2", "INT3", "ENH1", "ENH2", "DEG1", "DEG2")) %>%
  # # mutate(site_rep = fct_relevel(site_rep, "INT1_1", "INT2_2", "INT3_3", "ENH1_1" ,    "ENH1_2"   ,  "ENH1_3" ,    "ENH2_4" ,    "ENH2_5"    ,"ENH2_6", "DEG1_1",     "DEG1_2"  ,   "DEG1_3"   ,  "DEG2_4" ,    "DEG2_5"   ,  "DEG2_6")) %>% 
  # mutate(lifeform_status = fct_relevel(lifeform_status, "grass_nonnative", "herb_nonnative",
  #                                 "fern_native",     "grass_native", "herb_native", "nonsclero_native", "sclerophyllous_native")
  #)

print(lifeformleaftype)
```

### Percent cover by lifeform by native/nonnative status - there shouldn't be zeros in your df

```{r cover x status x lifeform - data prep, include = TRUE}
# line 933 from Ch1_Degradation_original.Rmd

# data prep
abcover_bylf_bystatus <- abcover_full_v_omitground %>% 
  group_by(standtype, lifeform, status, site, transect, site_rep, year) %>% 
  dplyr::summarize(
    count_per_transect = n_distinct(site_rep_dist), 
    percentcover = n_distinct(site_rep_dist)/41 # <-- calculate percent cover by transect, lifeform, and status
    ) %>% 
  ungroup() %>% 
  unite("lifeform_total", 2:3, sep = "_") #combines characters in two columns to one column
# view(abcover_bylf_bystatus) # prints to new tab
#write.csv(abcover_bylf_bystatus,"processed/abcover_bylf_bystatus.csv")
# visualizing data

## histogram 
hist(abcover_bylf_bystatus$percentcover) 
```
This is the spread of my percent cover data by standtype wrapped by lifeform_status
```{r}
## jitter - by standtype
ggplot(data = abcover_bylf_bystatus, aes(y = percentcover*100, x = standtype, color = year)) + 
  geom_jitter()+
  facet_wrap(~lifeform_total)
```
Here I'm transposing the data from long to wide by lifeform_status becuase I want to visualize native shrub cover by non-native grass cover.
```{r}
abcover_bylf_bystatus_wide <- abcover_bylf_bystatus %>% 
  dplyr::select(!c(count_per_transect)) %>% 
  pivot_wider(names_from = lifeform_total, values_from = percentcover) %>% 
  replace_na(replace = list(grass_native = 0, grass_nonnative = 0, 
                            herb_native = 0, herb_nonnative = 0, 
                            shrub_native = 0, shrub_nonnative = 0,
                            fern_native = 0, fern_nonnative = 0))  # Replace NA with 0 for each value column
  
```

This is the spread of native shrub cover over non-native grass cover...
```{r}

ggplot(data = abcover_bylf_bystatus_wide, 
       aes(y = shrub_native*100, x = grass_nonnative*100, color = year)) + 
  geom_jitter()
```
... and here is the spread of non-native grass cover overnative shrub cover.
```{r}
ggplot(data = abcover_bylf_bystatus_wide, 
       aes(y = grass_nonnative*100, x = shrub_native*100, color = year)) + 
  geom_jitter()
```
